### 2024-01-26 paper/paper.tex
add preliminary paper

\documentclass{scrartcl}

\usepackage[utf8]{inputenc}
\usepackage[usenames,dvipsnames,svgnames]{xcolor}
\usepackage[english]{babel}
\usepackage{mathtools}
\usepackage{bm,amsmath,amssymb}
\usepackage{microtype}
\usepackage[paperwidth=17cm, paperheight=24cm, top=2.5cm, left=2.0cm, bottom=1.5cm, right=.1cm]{geometry}
\usepackage[exponent-product=\cdot]{siunitx}
\usepackage{helvet}
\usepackage[pagecontinue=true]{pageslts}
\pagenumbering{arabic}
\usepackage{scrlayer-scrpage}
\usepackage{eurosym}
\usepackage{fp}
\usepackage{xspace}
\usepackage{verbatim}
\usepackage{enumitem}

\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{standalone}
\pgfplotsset{compat=newest}
\usetikzlibrary{positioning,calc}

\usepackage{lipsum}
% TODO use separate metadata.tex ?
\usepackage[colorlinks=true, linkcolor=MidnightBlue, citecolor=MidnightBlue, urlcolor=MidnightBlue, pdfauthor={Philipp Diercks}, pdfpagelayout=SinglePage, bookmarks, bookmarksopen]{hyperref}
\usepackage{cleveref}% must be loaded after hyperref
\usepackage{bookmark}

\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{wrapfig}

\usepackage{textcomp}
\usepackage{csquotes}
\usepackage{relsize}
\usepackage{bamcolors}

\usepackage[font=small,labelfont={bf,it},textfont=it]{caption}
\captionsetup{format=plain}

\usepackage{setspace}
\setstretch{1.2}

\usepackage[backend=biber,
    style=numeric, %alphabetic, numeric
    giveninits=true,
    natbib=true,
    url=false,
    doi=true,
    eprint=false,
    isbn=false,
    defernumbers=true,
    labelnumber,
    hyperref=true,
    maxbibnames=3,
    sorting=none,%remove this to have things sorted, e.g. use style=alphabetic
    ]{biblatex}

% add mapping to use shortjournal instead of journal field
\DeclareSourcemap{%
  \maps[datatype=bibtex]{%
    \map[overwrite]{% Notice the overwrite: replace one field with another
      \step[fieldsource=shortjournal,fieldtarget=journal]%
    }
  }  
}

% TODO add literature.bib
% \graphicspath{{img/}}
% \addbibresource{literature.bib}

% --------------------------------------------------------------------------------
% KOMA options
\KOMAoptions{paper=a4}
\KOMAoption{fontsize}{11pt}
\pagestyle{scrheadings}

% Maybe customize fonts
% \setkomafont{section}{\normalsize}
% \setkomafont{subsection}{\normalsize}
% \setkomafont{subparagraph}{\normalfont\itshape}

% \renewcommand{\familydefault}{\sfdefault}
% \renewcommand{\headfont}{\sffamily\footnotesize}

% Maybe redeclare section commands
% \RedeclareSectionCommands[
% beforeskip=.1em,
% runin=false,
% afterskip=-0.5\parskip
% ]{section,subsection,subsubsection}

% --------------------------------------------------------------------------------
% custom commands
\newcommand{\ie}{i.\,e.\@\xspace}
\newcommand{\eg}{e.\,g.\@\xspace}

\begin{document}

% TODO Titlepage
% TODO frontmatter

\section{Introduction}

\section{Methods}

\section{Results}

\subsection{Basis Construction}
For the beam problem, there are three oversampling problems to consider (left, inner, right).
For each of the associated parametric transfer operators, $n_{train}$ parameter samples are chosen, and
the range for each of these (fixed) transfer operators is approximated via random sampling. In
the sampling \textit{normal} or \textit{multivariate normal} distribution is used.
The range approximation of the $n_{train}$ transfer operators yields $n_{train}$ sets of
basis functions, which are further compressed via POD, to obtain the final parameter
independent basis functions (POD modes).

\subsubsection{Singular values of POD modes}

\includegraphics[width=0.6\textwidth]{../work/beam/fig_loc_svals_left.pdf}
\includegraphics[width=0.6\textwidth]{../work/beam/fig_loc_svals_inner.pdf}
\includegraphics[width=0.6\textwidth]{../work/beam/fig_loc_svals_right.pdf}

\subsection{Projection Error}
The projection error is computed for a test set for each configuration that was computed
using the FOM. Each test set has size $n_{test}$.

\includegraphics[width=0.6\textwidth]{../work/beam/fig_proj_error_left.pdf}
\includegraphics[width=0.6\textwidth]{../work/beam/fig_proj_error_inner.pdf}
\includegraphics[width=0.6\textwidth]{../work/beam/fig_proj_error_right.pdf}

\section{Conclusions}

% TODO Bibliography
% TODO backmatter

\end{document}

### 2024-01-26 src/getting_started/plot_projerr.py
update plots

@@ -1,57 +1,59 @@
+import sys
 import numpy as np
 from multi.plotting_context import PlottingContext
 from multi.postprocessing import read_bam_colors
 
 
-def main():
+def main(args):
     from .tasks import beam
     bamcd = read_bam_colors()
 
     markers = {"multivariate_normal": "x", "normal": "o"}  # use different markers per distribution
     colors = {"inner": bamcd["red"][0], "left": bamcd["blue"][0], "right": bamcd["green"][0]}  # use different colors per configuration
+    # FIXME: this could be done better
+    # issue is that PlottingContext is designed to get argv if len(argv) == 2, but
+    # I need len(argv) > 2
+    config = args.pop(2)
 
-    for config in beam.configurations:
-        outfile = beam.fig_proj_error(config)
+    with PlottingContext(args, ["paper_onecol"]) as fig:
+        # figure.figsize: 4.773, 2.950
+        width = 4.773
+        height = 2.95
+        factor = 1.0
+        fig.set_size_inches(factor * width, factor * height)
+        ax = fig.subplots()
 
-        with PlottingContext([__file__, outfile.as_posix()], "fast") as fig:
-            # figure.figsize: 4.773, 2.950
-            width = 4.773
-            height = 2.95
-            factor = 1.0
-            fig.set_size_inches(factor * width, factor * height)
-            ax = fig.subplots()
+        for distr in beam.distributions:
+            label = ""
+            if distr == "multivariate_normal":
+                label = "correlated, "
+            if distr == "normal":
+                label = "uncorrelated, "
 
-            for distr in beam.distributions:
-                label = ""
-                if distr == "multivariate_normal":
-                    label = "correlated, "
-                if distr == "normal":
-                    label = "uncorrelated, "
+            label += config
+            infile = beam.proj_error(distr, config)
+            errors = np.load(infile)
+            num_modes = np.arange(errors.size)
 
-                label += config
-                infile = beam.proj_error(distr, config)
-                errors = np.load(infile)
-                num_modes = np.arange(errors.size)
+            color = colors[config]
+            marker = markers[distr]
+            ax.semilogy(num_modes, errors, color=color, marker=marker, label=label)
+            # ax.semilogy(
+            #         data["num_dofs"], data["err"],
+            #         color=clr, marker=mark, label=label
[…]

### 2024-01-26 src/getting_started/plot_svals.py

@@ -1,9 +1,10 @@
+import sys
 import numpy as np
 from multi.plotting_context import PlottingContext
 from multi.postprocessing import read_bam_colors
 
 
-def main():
+def main(args):
     from .tasks import beam
 
     bamcd = read_bam_colors()
@@ -13,41 +14,39 @@ def main():
         "left": bamcd["blue"][0],
         "right": bamcd["green"][0],
     }
+    config = args.pop(2)
 
-    for config in beam.configurations:
-        outfile = beam.fig_loc_svals(config)
+    with PlottingContext(args, ["paper_onecol"]) as fig:
+        width = 4.773
+        height = 2.95
+        factor = 1.0
+        fig.set_size_inches(factor * width, factor * height)
+        ax = fig.subplots()
 
-        with PlottingContext([__file__, outfile.as_posix()], "fast") as fig:
-            width = 4.773
-            height = 2.95
-            factor = 1.0
-            fig.set_size_inches(factor * width, factor * height)
-            ax = fig.subplots()
+        for distr in beam.distributions:
+            label = ""
+            if distr == "multivariate_normal":
+                label = "correlated, "
+            if distr == "normal":
+                label = "uncorrelated, "
 
-            for distr in beam.distributions:
-                label = ""
-                if distr == "multivariate_normal":
-                    label = "correlated, "
-                if distr == "normal":
-                    label = "uncorrelated, "
+            svals = np.load(beam.loc_singular_values(distr, config))
 
-                svals = np.load(beam.loc_singular_values(distr, config))
+            color = colors[config]
+            marker = markers[distr]
+            label += config
+            ax.semilogy(
+                np.arange(svals.size),
+                svals / svals[0],
+                color=color,
+                marker=marker,
+                label=label,
+            )
 
-                color = colors[config]
-                marker = markers[distr]
-                label += config
-                ax.semilogy(
-                    np.arange(svals.size),
-                    svals / svals[0],
-                    color=color,
-                    marker=marker,
-                    label=label,
-                )
-
-            ax.set_xlabel("Number of basis functions")
-            ax.set_ylabel("Singular values")  # TODO: add formula
-            ax.legend(loc="best")
+        ax.set_xlabel("Number of basis functions")
+        ax.set_ylabel("Singular values")  # TODO: add formula
+        ax.legend(loc="best")
 
 
 if __name__ == "__main__":
-    main()
+    main(sys.argv)

### 2024-01-26 src/getting_started/tasks.py

@@ -90,7 +90,7 @@ def task_plot_loc_svals():
         yield {
             "basename": f"fig_loc_svals_{beam.name}_{config}",
             "file_dep": deps,
-            "actions": ["python3 -m {}".format(module)],
+            "actions": ["python3 -m {} %(targets)s {}".format(module, config)],
             "targets": [beam.fig_loc_svals(config)],
             "clean": True,
         }
@@ -148,7 +148,7 @@ def task_plot_proj_error():
         yield {
             "basename": f"fig_proj_err_{beam.name}_{config}",
             "file_dep": deps,
-            "actions": ["python3 -m {}".format(module)],
+            "actions": ["python3 -m {} %(targets)s {}".format(module, config)],
             "targets": [beam.fig_proj_error(config)],
             "clean": True,
         }
@@ -185,5 +185,5 @@ def task_optimize():
         "file_dep": [optpy, beam.reduced_model],
         "actions": ["python3 -m src.getting_started.optimization"],
         "verbosity": 2,
-        "uptodate": [False],
+        "uptodate": [True],
     }

### 2024-02-07 paper/eccomas-2024-paper-instructions.tex

\documentclass[a4paper]{eccomas_paper-2024}
%\usepackage{pslatex}
\usepackage{graphicx}

\usepackage{hyperref}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}

\title{INSTRUCTIONS TO PREPARE A FULL PAPER FOR THE 9th EUROPEAN CONGRESS ON COMPUTATIONAL METHODS IN APPLIED SCIENCES AND ENGINEERING\break ECCOMAS CONGRESS 2024}

\author{FIRST A. AUTHOR$^1$, SECOND B. AUTHOR$^2$ AND THIRD C. AUTHOR$^1$}

\heading{First A. Author, Second B. Author and Third C. Author}

\address{$^{1}$ International Center for Numerical Methods in Engineering (CIMNE)\\
Universidad Polit\'{e}cnica de Catalu\~{n}a\\
Campus Norte UPC, 08034 Barcelona, Spain\\
e-mail: congress@cimne.upc.edu, www.cimne.com
\and
$^{2}$ Spanish Association for Numerical Methods in Engineering (SEMNI)\\
Edificio C1, Campus Norte UPC\\
Gran Capit\'{a}n s/n, 08034 Barcelona, Spain\\
email: semni@cimne.upc.edu, www.semni.org
}

\keywords{Computational Mechanics, FEM, Contact Problems}

\abstract{This document provides information and instructions for
preparing a Full Paper to be included in the Proceedings
of {\it ECCOMAS CONGRESS 2024}. }

\begin{document}
\thispagestyle{empty}

\section{INTRODUCTION}

All the participants whose Abstract has been accepted for presentation at the ECCOMAS 2022 Congress are kindly requested to submit the Full Paper electronically via the web page of the \href{https://eccomas2024.org/lisbon}{Congress}, before \textbf{July 1st, 2024}. Congress Proceedings (including full papers only) will be published on Scipedia with DOI number after the congress and will be submitted for indexation to SCOPUS database. Submission of the full paper is not mandatory. The Full Paper should be written following the format of macros for submission. The file has to be translated into Portable Document Format (PDF) before submission via the Conference site. The organizers do not commit themselves to include in the Proceedings any Full Paper received later than the above-mentioned deadline. The corresponding author should be the speaker, and is expected to register and pay his registration fee during the advance period (before \textbf{April 1st, 2024}) for the paper to be included in the technical programme of the ECCOMAS 2024.

\section{GENERAL ESPECIFICATIONS}

The Full Paper must be written in English within a printing box of 16cm x 21cm, centered in the page. The Full Paper including figures, tables and references must have a minimum length of 6 pages and must not exceed 12 pages. Maximum file size is 4 MB.

\section{TITLE, AUTHORS, AFFILIATION, KEY WORDS}

The first page must contain the Title, Author(s), Affiliation(s),
Key words and the Summary. The Introduction must begin immediately
below, following the format of this template.

\subsection{Title}

The title should be written centered, in 14pt, boldface Roman, all
capital letters. It should be single spaced if the title is more
than one line long.

\subsection{Author}

The author's name should include first name, middle initial and
surname. It should be written centered, in 12pt boldface Roman,
12pt below the title.

\subsection{Affiliation}

Author's affiliation should be written centered, in 11pt Roman,
12pt below the list of authors. A 12pt space should separate two
different affiliations.

\subsection{Key words}

Please, write no more than six key words. They should be written
left aligned, in 12pt Roman, and the line must begin with the
words {\bf Key words}: boldfaced. A 12pt space should separate the
key words from the affiliations.

\subsection {Summary (optional)}

Use 12pt Italic Roman for the summary. The word {\bf Summary} must
be set in boldface, not italicized, at the beginning of the first
line. The text should be justified and separated 12pt from the key
words, as shown in the first page of these instructions.

\section{HEADINGS}

\subsection{Main headings}

The main headings should be written left aligned, in 12pt,
boldface and all capital Roman letters. There should be a 12pt
space before and 6pt after the main headings.

\subsection {Secondary headings}

Secondary headings should be written left aligned, 12 pt, boldface
Roman, with an initial capital for first word only. There should
be a 12pt space before and 6pt after the secondary headings.

\section{EDITORIAL HEADING}

The first page has to include the Editorial Heading, as shown in
the first page of these instructions. Successive pages will
include the name of the authors.

\section{TEXT}

The normal text should be written single-spaced, justified, using
12pt (Times New) Roman in one column. The first line of each
paragraph must be indented 0.5cm. There is not inter-paragraph
spacing.

\section{PAGE NUMBERS}

In order to organize the Full Paper, it is better to number
the pages. Page numbers are not included in the printing box.

\section{FIGURES}

All figures should be numbered consecutively and captioned. The
caption title should be written centered, in 10pt Roman, with
upper and lower case letters.

\begin{figure}[t]
\centering
\includegraphics[width=11cm]{Lisbon}
\caption{Lisbon city}
\end{figure}

A 6pt space should separate the figure from the caption, and a
12pt space should separate the upper part of the figure and the
bottom of the caption from the surrounding text.

Figures may be included in the text or added at the bottom of the paper.

\section{EQUATIONS}

A displayed equation is numbered, using Arabic numbers in
parentheses. It should be centered, leaving a 6pt space above and
below to separate it from the surrounding text.

The following example is a single line equation:
\vskip-.6cm
\begin{eqnarray}
Ax = b
\end{eqnarray}

The next example is a multi-line equation:
\vskip-.6cm
\begin{eqnarray}
Ax = b \\
Ax = b \nonumber
\end{eqnarray}

\section{TABLES}

All tables should be numbered consecutively and captioned, the
caption should be 10pt Roman, upper and lower case letters.

\begin{table}[h!]
\caption{Example of the construction of one table}
\begin{center}
\begin{tabular}{*{3}{c}}
\hline
C11 & C12 & C13 \\
\hline
C21 & C22 & C23 \\
\hline
C31 & C32 & C33 \\
\hline
C41 & C42 & C43 \\
\hline
C51 & C52 & C53 \\
\hline
\end{tabular}
\end{center}
\end{table}

A 6pt space should separate the table from the caption, and a 12pt
space should separate the table from the surrounding text.

\section{FORMAT OF REFERENCES}

References should be quoted in the text by superscript
numbers \cite{McQuerry18,nssga,Oden18} and grouped together at the end of the Full Paper in numerical order as shown in these instructions.

\section{CONCLUSIONS}

\begin{itemize}
\item[-] Full Papers in format for publication should be submitted electronically via the web page of the Conference, before \textbf{July 1st, 2024}. They must be converted to  Portable Document Format (PDF) before submission. The maximum size of the file is 4 Mb.

\end{itemize}

\begin{thebibliography}{99}
\bibitem{McQuerry18}  McQuerry, M. 2018. “Validation of a Clothing Heat Transfer Model in Nonisothermal Test Conditions.” J Test Eval 46, no. 1 (January/February): 1–7. http://doi.org/10.1520/JTE20170073

\bibitem{nssga}   National Stone, Sand, and Gravel Association (NSSGA). 2013. The Aggregates Handbook, 2nd ed. Englewood, CO: Society of Mining, Metallurgy, and Exploration.

\bibitem{Oden18}  Oden, C. P., C. L. Ho, and H. F. Kashani. 2018. “Man-Portable Real-Time Ballast Inspection Device Using Ground-Penetrating Radar.” In Railroad Ballast Testing and Properties, edited by T. Stark, R. Swan, and R. Szecsy, 77–104. West Conshohocken, PA: ASTM International. http://doi.org/10.1520/STP160520170023

\end{thebibliography}

\end{document}

WCCM-ECCOMAS2020@cimne.upc.edu

\begin{thebibliography}{99}
\bibitem{Zienkiewicz}  Zienkiewicz, O.C. and  Taylor, R.L. \textit{The finite element method}. McGraw Hill,
Vol. I., (1989), Vol. II., (1991).
\bibitem{Idelsohn} Idelsohn, S.R. and O\~{n}ate, E. Finite element and finite volumes. Two good friends.
\textit{Int. J. Num. Meth. Engng.} (1994) \textbf{37}:3323--3341.
\end{thebibliography}

\end{document}

### 2024-02-07 paper/paper.tex

@@
-\documentclass{scrartcl}
+\documentclass[a4paper]{eccomas_paper-2024}
 
-\usepackage[utf8]{inputenc}
-\usepackage[usenames,dvipsnames,svgnames]{xcolor}
-\usepackage[english]{babel}
-\usepackage{mathtools}
-\usepackage{bm,amsmath,amssymb}
-\usepackage{microtype}
-\usepackage[paperwidth=17cm, paperheight=24cm, top=2.5cm, left=2.0cm, bottom=1.5cm, right=.1cm]{geometry}
-\usepackage[exponent-product=\cdot]{siunitx}
-\usepackage{helvet}
-\usepackage[pagecontinue=true]{pageslts}
-\pagenumbering{arabic}
-\usepackage{scrlayer-scrpage}
-\usepackage{eurosym}
-\usepackage{fp}
-\usepackage{xspace}
-\usepackage{verbatim}
-\usepackage{enumitem}
+\usepackage{graphicx}
+\usepackage{hyperref}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
 
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{standalone}
-\pgfplotsset{compat=newest}
-\usetikzlibrary{positioning,calc}
+\title{INSTRUCTIONS TO PREPARE A FULL PAPER FOR THE 9th EUROPEAN CONGRESS ON COMPUTATIONAL METHODS IN APPLIED SCIENCES AND ENGINEERING\break ECCOMAS CONGRESS 2024}
 
-\usepackage{lipsum}
-% TODO use separate metadata.tex ?
-\usepackage[colorlinks=true, linkcolor=MidnightBlue, citecolor=MidnightBlue, urlcolor=MidnightBlue, pdfauthor={Philipp Diercks}, pdfpagelayout=SinglePage, bookmarks, bookmarksopen]{hyperref}
-\usepackage{cleveref}% must be loaded after hyperref
-\usepackage{bookmark}
+\author{FIRST A. AUTHOR$^1$, SECOND B. AUTHOR$^2$ AND THIRD C. AUTHOR$^1$}
 
-\usepackage{gensymb}
-\usepackage{graphicx}
-\usepackage{wrapfig}
+\heading{First A. Author, Second B. Author and Third C. Author}
 
-\usepackage{textcomp}
-\usepackage{csquotes}
-\usepackage{relsize}
-\usepackage{bamcolors}
-
-\usepackage[font=small,labelfont={bf,it},textfont=it]{caption}
-\captionsetup{format=plain}
-
-\usepackage{setspace}
-\setstretch{1.2}
-
-\usepackage[backend=biber,
-    style=numeric, %alphabetic, numeric
-    giveninits=true,
-    natbib=true,
-    url=false,
-    doi=true,
-    eprint=false,
-    isbn=false,
-    defernumbers=true,
-    labelnumber,
-    hyperref=true,
-    maxbibnames=3,
-    sorting=none,%remove this to have things sorted, e.g. use style=alphabetic
-    ]{biblatex}
-
-% add mapping to use shortjournal instead of journal field
-\DeclareSourcemap{%
-  \maps[datatype=bibtex]{%
-    \map[overwrite]{% Notice the overwrite: replace one field with another
-      \step[fieldsource=shortjournal,fieldtarget=journal]%
-    }
-  }  
+\address{$^{1}$ International Center for Numerical Methods in Engineering (CIMNE)\\
+Universidad Polit\'{e}cnica de Catalu\~{n}a\\
+Campus Norte UPC, 08034 Barcelona, Spain\\
+e-mail: congress@cimne.upc.edu, www.cimne.com
+\and
+$^{2}$ Spanish Association for Numerical Methods in Engineering (SEMNI)\\
+Edificio C1, Campus Norte UPC\\
+Gran Capit\'{a}n s/n, 08034 Barcelona, Spain\\
+email: semni@cimne.upc.edu, www.semni.org
 }
 
-% TODO add literature.bib
-% \graphicspath{{img/}}
-% \addbibresource{literature.bib}
+\keywords{Computational Mechanics, FEM, Contact Problems}
 
-% --------------------------------------------------------------------------------
-% KOMA options
-\KOMAoptions{paper=a4}
-\KOMAoption{fontsize}{11pt}
-\pagestyle{scrheadings}
-
-% Maybe customize fonts
-% \setkomafont{section}{\normalsize}
-% \setkomafont{subsection}{\normalsize}
-% \setkomafont{subparagraph}{\normalfont\itshape}
-
-% \renewcommand{\familydefault}{\sfdefault}
-% \renewcommand{\headfont}{\sffamily\footnotesize}
-
-% Maybe redeclare section commands
-% \RedeclareSectionCommands[
-% beforeskip=.1em,
-% runin=false,
-% afterskip=-0.5\parskip
-% ]{section,subsection,subsubsection}
-
-% --------------------------------------------------------------------------------
-% custom commands
-\newcommand{\ie}{i.\,e.\@\xspace}
-\newcommand{\eg}{e.\,g.\@\xspace}
+\abstract{This document provides information and instructions for
+preparing a Full Paper to be included in the Proceedings
+of {\it ECCOMAS CONGRESS 2024}. }
 
 \begin{document}
+\thispagestyle{empty}
 
-% TODO Titlepage
-% TODO frontmatter
+\section{INTRODUCTION}
 
-\section{Introduction}
+All the participants whose Abstract has been accepted for presentation at the ECCOMAS 2022 Congress are kindly requested to submit the Full Paper electronically via the web page of the \href{https://eccomas2024.org/lisbon}{Congress}, before \textbf{July 1st, 2024}.
+Congress Proceedings (including full papers only) will be published on Scipedia with DOI number after the congress and will be submitted for indexation to SCOPUS database.
+Submission of the full paper is not mandatory.
+The Full Paper should be written following the format of macros for submission.
+The file has to be translated into Portable Document Format (PDF) before submission via the Conference site.
+The organizers do not commit themselves to include in the Proceedings any Full Paper received later than the above-mentioned deadline.
+The corresponding author should be the speaker, and is expected to register and pay his registration fee during the advance period (before \textbf{April 1st, 2024}) for the paper to be included in the technical programme of the ECCOMAS 2024.
 
-\section{Methods}
+\section{METHODS}
 
-\section{Results}
+\section{RESULTS}
 
@@
 
-\includegraphics[width=0.6\textwidth]{../work/beam/fig_loc_svals_left.pdf}
-\includegraphics[width=0.6\textwidth]{../work/beam/fig_loc_svals_inner.pdf}
-\includegraphics[width=0.6\textwidth]{../work/beam/fig_loc_svals_right.pdf}
+\begin{figure}[!htb]
+\minipage{0.32\textwidth}
+  \includegraphics[width=\linewidth]{../work/beam/fig_loc_svals_left.pdf}
+  \caption{A really Awesome Image}\label{fig:awesome_image1}
+\endminipage\hfill
+\minipage{0.32\textwidth}
+  \includegraphics[width=\linewidth]{../work/beam/fig_loc_svals_inner.pdf}
+  \caption{A really Awesome Image}\label{fig:awesome_image2}
+\endminipage\hfill
+\minipage{0.32\textwidth}%
+  \includegraphics[width=\linewidth]{../work/beam/fig_loc_svals_right.pdf}
+  \caption{A really Awesome Image}\label{fig:awesome_image3}
+\endminipage
+\end{figure}
 
@@
 
-\includegraphics[width=0.6\textwidth]{../work/beam/fig_proj_error_left.pdf}
-\includegraphics[width=0.6\textwidth]{../work/beam/fig_proj_error_inner.pdf}
-\includegraphics[width=0.6\textwidth]{../work/beam/fig_proj_error_right.pdf}
+\begin{figure}[!htb]
+\minipage{0.32\textwidth}
+  \includegraphics[width=\linewidth]{../work/beam/fig_proj_error_left.pdf}
+  \caption{A really Awesome Image}\label{fig:awesome_image1}
+\endminipage\hfill
+\minipage{0.32\textwidth}
+  \includegraphics[width=\linewidth]{../work/beam/fig_proj_error_inner.pdf}
+  \caption{A really Awesome Image}\label{fig:awesome_image2}
+\endminipage\hfill
+\minipage{0.32\textwidth}%
+  \includegraphics[width=\linewidth]{../work/beam/fig_proj_error_right.pdf}
+  \caption{A really Awesome Image}\label{fig:awesome_image3}
+\endminipage
+\end{figure}
 
-\section{Conclusions}
+\section{FORMAT OF REFERENCES}
 
-% TODO Bibliography
-% TODO backmatter
+References should be quoted in the text by superscript
+numbers \cite{McQuerry18,nssga,Oden18} and grouped together at the end of the Full Paper in numerical order as shown in these instructions.
+
+\section{CONCLUSIONS}
+
+\begin{itemize}
+\item[-] Full Papers in format for publication should be submitted electronically via the web page of the Conference, before \textbf{July 1st, 2024}.
+They must be converted to  Portable Document Format (PDF) before submission.
+The maximum size of the file is 4 Mb.
+\end{itemize}
+
+\begin{thebibliography}{99}
+\bibitem{McQuerry18}  McQuerry, M.
+2018.
+“Validation of a Clothing Heat Transfer Model in Nonisothermal Test Conditions.” J Test Eval 46, no.
+1 (January/February): 1–7.
+http://doi.org/10.1520/JTE20170073
+
+\bibitem{nssga}   National Stone, Sand, and Gravel Association (NSSGA). 2013. The Aggregates Handbook, 2nd ed. Englewood, CO: Society of Mining, Metallurgy, and Exploration.
+
+\bibitem{Oden18}  Oden, C.
+P., C.
+L.
+Ho, and H.
+F.
+Kashani.
+2018.
+“Man-Portable Real-Time Ballast Inspection Device Using Ground-Penetrating Radar.” In Railroad Ballast Testing and Properties, edited by T.
+Stark, R.
+Swan, and R.
+Szecsy, 77–104.
+West Conshohocken, PA: ASTM International.
+http://doi.org/10.1520/STP160520170023
+
+\end{thebibliography}
 

### 2024-02-27 src/getting_started/definitions.py
definitions: add realizations; increase mu range

@@ -7,6 +7,7 @@ from multi.problems import MultiscaleProblemDefinition
 from dolfinx import default_scalar_type
 
 ROOT = Path(__file__).parents[2]
+SRC = Path(__file__).parent
 WORK = ROOT / "work"
 
 
@@ -16,6 +17,7 @@ class BeamData:
 
     Args:
         name: The name of the example.
+        num_real: The number of realizations of this example.
         gdim: The geometric dimension of the problem.
         length: The length of the beam.
         height: The height of the beam.
@@ -39,6 +41,7 @@ class BeamData:
     """
 
     name: str = "example"
+    num_real: int = 1
     gdim: int = 2
     length: float = 10.0
     height: float = 1.0
@@ -49,13 +52,13 @@ class BeamData:
     fe_deg: int = 2
     poisson_ratio: float = 0.3
     youngs_modulus: float = 20e3
-    mu_range: tuple[float, float] = (1.0, 2.0)
+    mu_range: tuple[float, float] = (0.1, 10.0)
     rrf_ttol: float = 5e-2
     rrf_ftol: float = 1e-15
     rrf_num_testvecs: int = 20
     pod_rtol: float = 1e-5
     configurations: tuple[str, str, str] = ("inner", "left", "right")
-    distributions: tuple[str, str] = ("normal", "multivariate_normal")
+    distributions: tuple[str, ...] = ("normal",)
     range_product: str = "h1"
     lhs: dict = field(
         default_factory=lambda: {
@@ -215,6 +218,19 @@ class BeamData:
         """loc ROM error relative to FOM"""
         return self.rf / f"loc_rom_error_{distr}.csv"
 
+    @property
+    def realizations(self) -> Path:
+        """Returns realizations that can be used to create ``np.random.SeedSequence``"""
+        file = SRC / "realizations.npy"
+        if not file.exists():
+            self._generate_realizations(file)
+        return file
+
+    def _generate_realizations(self, outpath: Path) -> None:
+        seed = np.random.SeedSequence()
+        realizations = seed.generate_state(self.num_real)
+        np.save(outpath, realizations)
+
 
 class BeamProblem(MultiscaleProblemDefinition):
     def __init__(self, coarse_grid: str, fine_grid: str):
@@ -336,4 +352,6 @@ class BeamProblem(MultiscaleProblemDefinition):
 
 if __name__ == "__main__":
     data = BeamData(name="beam")
[…]

### 2024-02-27 src/getting_started/tasks.py
tasks: fix deps in task_paper

@@ -259,6 +259,7 @@ def task_paper():
     for config in CONFIGS:
         deps.append(beam.fig_loc_svals(config))
         deps.append(beam.fig_proj_error(config))
+        deps.append(beam.fig_loc_rom_error)
     return {
         "file_dep": deps,
         "actions": ["latexmk -cd -pdf %s" % source],

### 2024-02-27 src/getting_started/run_locrom.py
run_locrom: dont use submeshes for reconstruction

@@ -5,20 +5,20 @@ from scipy.sparse import coo_array
 import numpy as np
 
 from mpi4py import MPI
-from dolfinx import mesh, fem, default_scalar_type
+from dolfinx import fem, default_scalar_type
 from dolfinx.io import gmshio
-from dolfinx.io.utils import XDMFFile
 from basix.ufl import element
 
 from pymor.parameters.base import Parameters
 from pymor.core.defaults import set_defaults
 from pymor.core.logger import getLogger
-from pymor.bindings.fenicsx import FenicsxVectorSpace, FenicsxMatrixOperator
+from pymor.bindings.fenicsx import FenicsxVectorSpace, FenicsxMatrixOperator, FenicsxVisualizer
 from pymor.operators.constructions import VectorOperator, LincombOperator
 from pymor.operators.numpy import NumpyMatrixOperator
 from pymor.algorithms.projection import project
 from pymor.models.basic import StationaryModel
 
+from multi.interpolation import make_mapping
 from multi.boundary import point_at
 from multi.domain import RectangularSubdomain
 from multi.dofmap import DofMap
@@ -27,12 +27,61 @@ from multi.materials import LinearElasticMaterial
 from multi.problems import LinElaSubProblem
 
 
+# FIXME: this reconstruction method does not seem to be stable wrt the mesh
+# it did work using `create_rectangle` but not for `create_voided_rectangle`
+# my best guess is that the submesh creation is not reproducible?
+# and therefore you cannot assume the same doflayout for every subspace?
+# def reconstruct(
+#     U_rb: np.ndarray,
+#     dofmap: DofMap,
+#     bases: list[np.ndarray],
+#     subspaces: list[fem.FunctionSpaceBase],
+#     cell_maps: list[np.ndarray],
+#     u_global: fem.Function,
+# ) -> None:
+#     """Reconstructs rom solution on the global domain.
+#
+#     Args:
+#         Urb: ROM solution in the reduced space.
+#         dofmap: The dofmap of the reduced space.
+#         bases: Local basis for each subdomain.
+#         subspaces: The subspace of each subdomain.
+#         cell_maps: Mapping from child (submesh cell) to parent cell.
+#         subdomains: The cells of each subdomain.
+#         u_global: The global solution field to be filled with values.
+#
+#     """
+#     V = u_global.function_space
[…]

### 2024-02-27 paper/paper.tex
paper: add loc rom error plot

@@
 
-\subsection{Projection Error}
+\subsubsection{Projection Error}
 The projection error is computed for a test set for each configuration that was computed
@@
 \endminipage
+\end{figure}
+
+\subsection{Localized ROM}
+
+\begin{figure}[!htb]
+    \includegraphics[width=\linewidth]{../work/beam/fig_loc_rom_error.pdf}
+    \caption{A really Awesome Image}\label{fig:loc_rom_error}
 \end{figure}

### 2024-03-06 src/beam/definitions.py
definitions: write pdfs to figures dir

@@ -7,8 +7,9 @@ from multi.problems import MultiscaleProblemDefinition
 from dolfinx import default_scalar_type
 
 ROOT = Path(__file__).parents[2]
-SRC = Path(__file__).parent
+FIGURES = ROOT / "figures"
 WORK = ROOT / "work"
+SRC = Path(__file__).parent
 
 
 @dataclass
@@ -175,16 +176,16 @@ class BeamData:
 
     def fig_proj_error(self, conf: str) -> Path:
         """figure of projection error plot"""
-        return self.rf / f"fig_proj_error_{conf}.pdf"
+        return FIGURES / f"fig_proj_error_{conf}.pdf"
 
     def fig_loc_svals(self, config: str) -> Path:
         """figure of singular values of POD compression after rrf"""
-        return self.rf / f"fig_loc_svals_{config}.pdf"
+        return FIGURES / f"fig_loc_svals_{config}.pdf"
 
     @property
     def fig_loc_rom_error(self) -> Path:
         """figure of loc rom error"""
-        return self.rf / f"fig_loc_rom_error.pdf"
+        return FIGURES / f"fig_loc_rom_error.pdf"
 
     def config_to_cell(self, config: str) -> int:
         """Maps config to global cell index."""

### 2024-03-06 src/beam/definitions.py
definitions: add figures subfolder

@@ -7,7 +7,7 @@ from multi.problems import MultiscaleProblemDefinition
 from dolfinx import default_scalar_type
 
 ROOT = Path(__file__).parents[2]
-FIGURES = ROOT / "figures"
+FIGURES = ROOT / "figures" / "beam"
 WORK = ROOT / "work"
 SRC = Path(__file__).parent
 

### 2024-03-06 paper/paper.tex
paper: fix figure path and labels

@@
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../work/beam/fig_loc_svals_left.pdf}
-  \caption{A really Awesome Image}\label{fig:awesome_image1}
+  \includegraphics[width=\linewidth]{../figures/beam/fig_loc_svals_left.pdf}
+  \caption{A really Awesome Image}\label{fig:loc_svals_left}
 \endminipage\hfill
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../work/beam/fig_loc_svals_inner.pdf}
-  \caption{A really Awesome Image}\label{fig:awesome_image2}
+  \includegraphics[width=\linewidth]{../figures/beam/fig_loc_svals_inner.pdf}
+  \caption{A really Awesome Image}\label{fig:loc_svals_inner}
 \endminipage\hfill
 \minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{../work/beam/fig_loc_svals_right.pdf}
-  \caption{A really Awesome Image}\label{fig:awesome_image3}
+  \includegraphics[width=\linewidth]{../figures/beam/fig_loc_svals_right.pdf}
+  \caption{A really Awesome Image}\label{fig:loc_svals_right}
 \endminipage
@@
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../work/beam/fig_proj_error_left.pdf}
-  \caption{A really Awesome Image}\label{fig:awesome_image1}
+  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_left.pdf}
+  \caption{A really Awesome Image}\label{fig:proj_error_left}
 \endminipage\hfill
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../work/beam/fig_proj_error_inner.pdf}
-  \caption{A really Awesome Image}\label{fig:awesome_image2}
+  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_inner.pdf}
+  \caption{A really Awesome Image}\label{fig:proj_error_inner}
 \endminipage\hfill
 \minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{../work/beam/fig_proj_error_right.pdf}
-  \caption{A really Awesome Image}\label{fig:awesome_image3}
+  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_right.pdf}
+  \caption{A really Awesome Image}\label{fig:proj_error_right}
 \endminipage
@@
 \begin{figure}[!htb]
-    \includegraphics[width=\linewidth]{../work/beam/fig_loc_rom_error.pdf}
+    \includegraphics[width=\linewidth]{../figures/beam/fig_loc_rom_error.pdf}
     \caption{A really Awesome Image}\label{fig:loc_rom_error}

### 2024-03-06 src/beam/tasks.py
tasks: fix module name

@@ -7,8 +7,8 @@ import shutil
 from doit.tools import run_once
 
 os.environ["PYMOR_COLORS_DISABLE"] = "1"
-SRC = ROOT / "src/getting_started"  # source for this example
 beam = BeamData(name="beam")
+SRC = ROOT / "src" / f"{beam.name}"
 CONFIGS = beam.configurations
 DISTR = beam.distributions
 
@@ -72,7 +72,7 @@ def task_build_rom():
             SRC / "fom.py",
             SRC / "rom.py",
         ],
-        "actions": ["python3 -m src.getting_started.rom"],
+        "actions": [f"python3 -m src.{beam.name}.rom"],
         "targets": [beam.reduced_model, beam.singular_values],
         "clean": True,
     }
@@ -80,7 +80,7 @@ def task_build_rom():
 
 def task_edge_range_approximation():
     """Getting started: Construct POD edge basis"""
-    module = "src.getting_started.range_approx_edge"
+    module = f"src.{beam.name}.range_approx_edge"
     file = SRC / "range_approx_edge.py"
     nworkers = 4  # number of workers in pool
     for distr in DISTR:
@@ -107,7 +107,7 @@ def task_edge_range_approximation():
 
 def task_plot_loc_svals():
     """Getting started: Figure Singular Values"""
-    module = "src.getting_started.plot_svals"
+    module = f"src.{beam.name}.plot_svals"
     code = SRC / "plot_svals.py"
     for config in beam.configurations:
         deps = [code]
@@ -125,7 +125,7 @@ def task_plot_loc_svals():
 
 def task_test_sets():
     """Getting started: Generate FOM test sets"""
-    module = "src.getting_started.fom_test_set"
+    module = f"src.{beam.name}.fom_test_set"
     code = SRC / "fom_test_set.py"
     num_solves = 20
     map = {"inner": 4, "left": 0, "right": 9}
@@ -147,7 +147,7 @@ def task_test_sets():
 
 def task_proj_error():
     """Getting started: Projection error for FOM test sets"""
-    module = "src.getting_started.projerr"
+    module = f"src.{beam.name}.projerr"
     code = SRC / "projerr.py"
     for distr in DISTR:
         for config in CONFIGS:
@@ -170,7 +170,7 @@ def task_proj_error():
 
 def task_plot_proj_error():
     """Getting started: Figure Projection Error"""
-    module = "src.getting_started.plot_projerr"
+    module = f"src.{beam.name}.plot_projerr"
[…]

### 2024-03-06 src/beam/definitions.py
add plot of unit cell

@@ -124,6 +124,10 @@ class BeamData:
     def unit_cell_grid(self) -> Path:
         return self.grids_path / "unit_cell.msh"
 
+    @property
+    def fig_unit_cell(self) -> Path:
+        return FIGURES / "unit_cell.png"
+
     def fine_oversampling_grid(self, configuration: str) -> Path:
         assert configuration in ("inner", "left", "right")
         return self.grids_path / f"fine_oversampling_grid_{configuration}.xdmf"

### 2024-03-06 src/beam/tasks.py

@@ -1,4 +1,4 @@
-"""tasks for the getting started example"""
+"""tasks for the beam example"""
 
 from .definitions import BeamData, ROOT
 from pathlib import Path
@@ -44,7 +44,7 @@ def with_h5(xdmf: Path) -> list[Path]:
 
 
 def task_preprocessing():
-    """Getting started: Preprocessing"""
+    """Beam example: Preprocessing"""
     from .preprocessing import generate_meshes
 
     mesh_files = [beam.coarse_grid, beam.unit_cell_grid, *with_h5(beam.fine_grid)]
@@ -62,7 +62,7 @@ def task_preprocessing():
 
 
 def task_build_rom():
-    """Getting started: Build ROM"""
+    """Beam example: Build ROM"""
     return {
         "basename": f"build_rom_{beam.name}",
         "file_dep": [
@@ -79,7 +79,7 @@ def task_build_rom():
 
 
 def task_edge_range_approximation():
-    """Getting started: Construct POD edge basis"""
+    """Beam example: Construct POD edge basis"""
     module = f"src.{beam.name}.range_approx_edge"
     file = SRC / "range_approx_edge.py"
     nworkers = 4  # number of workers in pool
@@ -106,7 +106,7 @@ def task_edge_range_approximation():
 
 
 def task_plot_loc_svals():
-    """Getting started: Figure Singular Values"""
+    """Beam example: Figure Singular Values"""
     module = f"src.{beam.name}.plot_svals"
     code = SRC / "plot_svals.py"
     for config in beam.configurations:
@@ -124,7 +124,7 @@ def task_plot_loc_svals():
 
 
 def task_test_sets():
-    """Getting started: Generate FOM test sets"""
+    """Beam example: Generate FOM test sets"""
     module = f"src.{beam.name}.fom_test_set"
     code = SRC / "fom_test_set.py"
     num_solves = 20
@@ -146,7 +146,7 @@ def task_test_sets():
 
 
 def task_proj_error():
-    """Getting started: Projection error for FOM test sets"""
+    """Beam example: Projection error for FOM test sets"""
     module = f"src.{beam.name}.projerr"
     code = SRC / "projerr.py"
     for distr in DISTR:
@@ -169,7 +169,7 @@ def task_proj_error():
 
 
 def task_plot_proj_error():
-    """Getting started: Figure Projection Error"""
+    """Beam example: Figure Projection Error"""
     module = f"src.{beam.name}.plot_projerr"
[…]

### 2024-03-07 src/beam/tasks.py
crop unit cell png

@@ -277,7 +277,7 @@ def task_plot_unit_cell_domain():
         plot_domain(domain, cell_tags=None, transparent=False, colormap="bam-RdBu", output=targets[0])
 
     return {
-            "actions": [make_plot],
+            "actions": [make_plot, "convert -trim %(targets)s %(targets)s"],
             "targets": [beam.fig_unit_cell],
             "uptodate": [True],
             "clean": True,

### 2024-03-07 paper/paper.tex
update paper.pdf
crop unit cell png
add global domain png

@@
 \section{RESULTS}
+
+\subsection{Problem setup}
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=0.65\textwidth]{../figures/beam/unit_cell.png}
+    \end{center}
+    \caption{Unit cell domain partition.}\label{fig:unit_cell_domain}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=0.65\textwidth]{../figures/beam/global_domain.png}
+    \end{center}
+    \caption{The partition of the global domain. The colors indicate the variation in Young's modulus for each subdomain.}\label{fig:global_domain}
+\end{figure}
 

### 2024-03-07 src/beam/definitions.py

@@ -121,6 +121,10 @@ class BeamData:
         return self.grids_path / "fine_grid.xdmf"
 
     @property
+    def fig_fine_grid(self) -> Path:
+        return FIGURES / "global_domain.png"
+
+    @property
     def unit_cell_grid(self) -> Path:
         return self.grids_path / "unit_cell.msh"
 

### 2024-03-07 src/beam/tasks.py

@@ -277,8 +277,31 @@ def task_plot_unit_cell_domain():
         plot_domain(domain, cell_tags=None, transparent=False, colormap="bam-RdBu", output=targets[0])
 
     return {
+            "file_dep": [beam.unit_cell_grid],
             "actions": [make_plot, "convert -trim %(targets)s %(targets)s"],
             "targets": [beam.fig_unit_cell],
+            "uptodate": [True],
+            "clean": True,
+            }
+
+
+def task_plot_global_domain():
+    """Beam example: Plot global domain"""
+
+    def make_plot(dependencies, targets):
+        from mpi4py import MPI
+        from dolfinx.io.utils import XDMFFile
+        from multi.postprocessing import plot_domain
+        xdmf_file = Path(dependencies[0]).with_suffix(".xdmf")
+        with XDMFFile(MPI.COMM_WORLD, xdmf_file.as_posix(), "r") as xdmf:
+            domain = xdmf.read_mesh(name="Grid")
+            cell_tags = xdmf.read_meshtags(domain, "subdomains")
+        plot_domain(domain, cell_tags=cell_tags, transparent=False, colormap="bam-RdBu", output=targets[0])
+
+    return {
+            "file_dep": with_h5(beam.fine_grid),
+            "actions": [make_plot, "convert -trim %(targets)s %(targets)s"],
+            "targets": [beam.fig_fine_grid],
             "uptodate": [True],
             "clean": True,
             }

### 2024-03-07 src/beam/tasks.py
add beam sketch

@@ -305,3 +305,14 @@ def task_plot_global_domain():
             "uptodate": [True],
             "clean": True,
             }
+
+
+def task_fig_beam_sketch():
+    """Beam example: Compile beam sketch"""
+    target = ROOT / "figures/beam/beam_sketch.pdf"
+    return {
+            "file_dep": [SRC / "figures/beam_sketch.tex"],
+            "actions": [f"latexmk -cd -pdf -outdir={target.parent} %(dependencies)s"],
+            "targets": [target],
+            "clean": True,
+            }

### 2024-03-07 src/beam/tasks.py
add beam sketch to paper

@@ -229,31 +229,6 @@ def task_plot_loc_rom_error():
             }
 
 
-def task_paper():
-    """Beam example: Compile Paper"""
-    source = ROOT / "paper/paper.tex"
-    deps = [source]
-    for config in CONFIGS:
-        deps.append(beam.fig_loc_svals(config))
-        deps.append(beam.fig_proj_error(config))
-        deps.append(beam.fig_loc_rom_error)
-    return {
-        "file_dep": deps,
-        "actions": ["latexmk -cd -pdf %s" % source],
-        "targets": [source.with_suffix(".pdf")],
-        "clean": ["latexmk -cd -C %s" % source],
-    }
-
-
-def task_show_paper():
-    """Beam example: View Paper"""
-    return {
-        "file_dep": [ROOT / "paper/paper.pdf"],
-        "actions": ["zathura %(dependencies)s"],
-        "uptodate": [False],
-    }
-
-
 def task_optimize():
     """Beam example: Optimization"""
     optpy = ROOT / f"src/{beam.name}/optimization.py"
@@ -316,3 +291,33 @@ def task_fig_beam_sketch():
             "targets": [target],
             "clean": True,
             }
+
+
+def task_paper():
+    """Beam example: Compile Paper"""
+    source = ROOT / "paper/paper.tex"
+    deps = [source]
+    deps.append(beam.fig_loc_rom_error)
+    deps.append(beam.fig_fine_grid)
+    deps.append(beam.fig_unit_cell)
+    deps.append(ROOT / "figures/beam/beam_sketch.pdf")
+    for config in CONFIGS:
+        deps.append(beam.fig_loc_svals(config))
+        deps.append(beam.fig_proj_error(config))
+    return {
+        "file_dep": deps,
+        "actions": ["latexmk -cd -pdf %s" % source],
+        "targets": [source.with_suffix(".pdf")],
+        "clean": ["latexmk -cd -C %s" % source],
+    }
+
+
+def task_show_paper():
+    """Beam example: View Paper"""
+    return {
+        "file_dep": [ROOT / "paper/paper.pdf"],
+        "actions": ["zathura %(dependencies)s"],
+        "uptodate": [False],
+    }
+
+

### 2024-03-07 paper/paper.tex
paper: add comment

@@
     \begin{center}
+        \includegraphics[width=0.95\textwidth]{../figures/beam/beam_sketch.pdf}
+    \end{center}
+    \caption{Schematic representation of the structural problem. A simply supported beam is subjected to a constant load.}\label{fig:beam_sketch}
+\end{figure}
+
+% TODO placement
+\begin{figure}
+    \begin{center}
         \includegraphics[width=0.65\textwidth]{../figures/beam/unit_cell.png}
@@
     \begin{center}
-        \includegraphics[width=0.65\textwidth]{../figures/beam/global_domain.png}
+        \includegraphics[width=0.95\textwidth]{../figures/beam/global_domain.png}
     \end{center}

### 2024-03-12 src/beam/range_approx_edge.py
fix range approximation hapod

@@ -327,7 +327,7 @@ def approximate_range(beam, mu, configuration, distribution="normal"):
     logger = getLogger(
         "range_approximation",
         level="INFO",
-        filename=beam.log_edge_range_approximation(distribution, configuration).as_posix(),
+        filename=beam.log_edge_range_approximation(distribution, configuration, "hapod").as_posix(),
     )
     pid = os.getpid()
     logger.info(f"{pid=},\tApproximating range of T for {mu=} using {distribution=}.\n")

### 2024-03-12 src/beam/tasks.py

@@ -243,7 +243,7 @@ def task_loc_rom():
                     "name": f"locrom_{name}_{distr}",
                     "file_dep": deps,
                     "actions": ["python3 -m {} {} {} --output {}".format(module, distr, num_test, target)],
-                    "targets": [target, beam.log_run_locrom(distr)],
+                    "targets": [target, beam.log_run_locrom(distr, name)],
                     "clean": True,
                     }
 

### 2024-03-12 src/beam/extension.py
extension: fix definitions

@@ -28,11 +28,11 @@ def main(args):
     set_defaults(
         {
             "pymor.core.logger.getLogger.filename": beam.log_extension(
-                args.distribution, args.cell
+                args.distribution, args.name, args.cell
             )
         }
     )
-    logger = getLogger(Path(__file__).stem, level="DEBUG")
+    logger = getLogger(Path(__file__).stem, level="INFO")
 
     # problem definition
     beamproblem = BeamProblem(beam.coarse_grid.as_posix(), beam.fine_grid.as_posix())
@@ -62,7 +62,11 @@ def main(args):
     boundary_entities = np.array([], dtype=np.intc)
     edges = set(["bottom", "left", "right", "top"])
     for edge in edges:
-        edge_entities = mesh.locate_entities_boundary(problem.domain.grid, problem.domain.tdim-1, problem.domain.str_to_marker(edge))
+        edge_entities = mesh.locate_entities_boundary(
+            problem.domain.grid,
+            problem.domain.tdim - 1,
+            problem.domain.str_to_marker(edge),
+        )
         boundary_entities = np.append(boundary_entities, edge_entities)
 
     bc_factory = BoundaryDataFactory(problem.domain.grid, boundary_entities, problem.V)
@@ -87,11 +91,15 @@ def main(args):
         # ### Fine scale edge modes
         configuration = beam.cell_to_config(ci)
         infile = beam.fine_scale_edge_modes_npz(args.distribution, configuration)
-        logger.debug(f"Reading fine scale modes for cell {args.cell} from file: {infile}")
+        logger.debug(
+            f"Reading fine scale modes for cell {args.cell} from file: {infile}"
+        )
         fine_scale_edge_modes = np.load(infile)
 
         modes = fine_scale_edge_modes[loc_edge]
-        logger.debug(f"Number of fine scale modes for local edge {loc_edge}: {len(modes)}")
+        logger.debug(
+            f"Number of fine scale modes for local edge {loc_edge}: {len(modes)}"
+        )
         end += modes.shape[0]
 
         if args.cell == ci:
@@ -102,7 +110,10 @@ def main(args):
             # in this case modes from neighbouring configuration are extended
             # the mapping of DOFs between different edge spaces has to be considered
             boundary = dof_layout.local_edge_index_map[local_ent]
-            map = make_mapping(problem.edge_spaces["fine"][boundary], problem.edge_spaces["fine"][loc_edge])
[…]

### 2024-03-12 src/beam/tasks.py

@@ -223,7 +223,7 @@ def task_extension():
                 yield {
                         "name": f"xi_{name}_{distr}_{cell_index}",
                         "file_dep": [beam.fine_scale_edge_modes_npz(distr, config, name)],
-                        "actions": ["python3 -m {} {} {}".format(module, distr, cell_index)],
+                        "actions": ["python3 -m {} {} {}".format(module, distr, name, cell_index)],
                         "targets": [beam.local_basis_npz(distr, name, cell_index), beam.fine_scale_modes_bp(distr, name, cell_index), beam.log_extension(distr, name, cell_index)],
                         "clean": [rm_rf],
                         }

### 2024-03-12 src/beam/tasks.py
tasks: shorten task names

@@ -79,7 +79,7 @@ def task_build_rom():
     }
 
 
-def task_edge_range_approximation():
+def task_edge_range_approx():
     """Beam example: Construct POD edge basis"""
     module = f"src.{beam.name}.range_approx_edge"
     file = SRC / "range_approx_edge.py"
@@ -106,7 +106,7 @@ def task_edge_range_approximation():
             }
 
 
-def task_heuristic_edge_range_approximation():
+def task_heuristic_edge_range_approx():
     """Beam example: Construct fine scale edge basis via heuristic range finder"""
     module = f"src.{beam.name}.heuristic_range_approx"
     file = SRC / "heuristic_range_approx.py"

### 2024-03-12 src/beam/definitions.py
BeamData: increase number of samples in LHS

@@ -68,19 +68,19 @@ class BeamData:
             "inner": {
                 "name": "E",
                 "ndim": 3,
-                "samples": 30,
+                "samples": 60,
                 "criterion": "center",
             },
             "left": {
                 "name": "E",
                 "ndim": 2,
-                "samples": 20,
+                "samples": 40,
                 "criterion": "center",
             },
             "right": {
                 "name": "E",
                 "ndim": 2,
-                "samples": 20,
+                "samples": 40,
                 "criterion": "center",
             },
         }

### 2024-03-12 src/beam/heuristic_range_approx.py
heuristic_range_approx: numpy conversion for output

@@ -404,10 +404,15 @@ def main(args):
             testing_set,
         )
 
+    # convert to numpy
+    output = {}
+    for k, vector_array in bases.items():
+        output[k] = vector_array.to_numpy()
+
     # write fine scale basis
     np.savez(
         beam.fine_scale_edge_modes_npz(args.distribution, args.configuration, "heuristic"),
-        **bases,
+        **output,
     )
 
 

### 2024-03-12 src/beam/extension.py
extension: make it pass

@@ -90,7 +90,7 @@ def main(args):
 
         # ### Fine scale edge modes
         configuration = beam.cell_to_config(ci)
-        infile = beam.fine_scale_edge_modes_npz(args.distribution, configuration)
+        infile = beam.fine_scale_edge_modes_npz(args.distribution, configuration, args.name)
         logger.debug(
             f"Reading fine scale modes for cell {args.cell} from file: {infile}"
         )

### 2024-03-12 src/beam/tasks.py

@@ -214,7 +214,7 @@ def task_plot_proj_error():
 
 def task_extension():
     """Beam example: Extend fine scale modes and write final basis"""
-    module = "src.{beam.name}.extension"
+    module = f"src.{beam.name}.extension"
     num_cells = beam.nx * beam.ny
     for distr in DISTR:
         for name in NAMES:
@@ -223,7 +223,7 @@ def task_extension():
                 yield {
                         "name": f"xi_{name}_{distr}_{cell_index}",
                         "file_dep": [beam.fine_scale_edge_modes_npz(distr, config, name)],
-                        "actions": ["python3 -m {} {} {}".format(module, distr, name, cell_index)],
+                        "actions": ["python3 -m {} {} {} {}".format(module, distr, name, cell_index)],
                         "targets": [beam.local_basis_npz(distr, name, cell_index), beam.fine_scale_modes_bp(distr, name, cell_index), beam.log_extension(distr, name, cell_index)],
                         "clean": [rm_rf],
                         }

### 2024-03-12 src/beam/plot_projerr.py
tweak proj error plot

@@ -1,4 +1,3 @@
-import sys
 import numpy as np
 from multi.plotting_context import PlottingContext
 from multi.postprocessing import read_bam_colors
@@ -19,12 +18,8 @@ def main(args):
         "right": bamcd["green"][0],
         "top": bamcd["yellow"][0],
     }
-    # FIXME: this could be done better
-    # issue is that PlottingContext is designed to get argv if len(argv) == 2, but
-    # I need len(argv) > 2
-    config = args.pop(2)
 
-    with PlottingContext(args, ["paper_onecol"]) as fig:
+    with PlottingContext([__file__, args.output], ["paper_onecol"]) as fig:
         # figure.figsize: 4.773, 2.950
         width = 4.773
         height = 2.95
@@ -32,29 +27,39 @@ def main(args):
         fig.set_size_inches(factor * width, factor * height)
         ax = fig.subplots()
 
-        for distr in beam.distributions:
-            infile = beam.proj_error(distr, config)
-            npz_err = np.load(infile)
-            marker = markers[distr]
+        config = args.configuration
+        name = args.name
+        distr = beam.distributions[0]
 
-            for edge, color in colors.items():
+        infile = beam.proj_error(distr, config, name)
+        npz_err = np.load(infile)
+        marker = markers[distr]
 
-                if distr == "multivariate_normal":
-                    label = f"correlated, {config}, {edge}"
-                if distr == "normal":
-                    label = f"uncorrelated, {config}, {edge}"
+        ymin = []
 
-                err = npz_err[edge]
-                num_modes = np.arange(err.size)
-                ax.semilogy(num_modes, err, color=color, marker=marker, label=label)
-                # ax.semilogy(
-                #         data["num_dofs"], data["err"],
-                #         color=clr, marker=mark, label=label
-                #         )
-                # ax.fill_between(data["num_dofs"], data["err"]-std, data["err"]+std,
-                #         alpha=0.2, color=clr
-                #         )
+        for edge, color in colors.items():
+            label = ""
+            if distr == "multivariate_normal":
+                label = f"correlated, {config}, {edge}"
+            if distr == "normal":
+                label = f"uncorrelated, {config}, {edge}"
 
+            err = npz_err[edge]
+            num_modes = np.arange(err.size)
+            ax.semilogy(num_modes, err, color=color, marker=marker, label=label)
[…]

### 2024-03-12 src/beam/tasks.py

@@ -206,7 +206,7 @@ def task_plot_proj_error():
             yield {
                 "name": f"fig_proj_err_{name}_{config}",
                 "file_dep": deps,
-                "actions": ["python3 -m {} %(targets)s {}".format(module, config)],
+                "actions": ["python3 -m {} {} {} --output %(targets)s".format(module, config, name)],
                 "targets": [beam.fig_proj_error(config, name)],
                 "clean": True,
             }

### 2024-03-12 src/beam/plot_locrom_error.py
tweak and plot local rom error

@@ -8,7 +8,7 @@ if __name__ == "__main__":
     from .tasks import beam
 
     bamcd = read_bam_colors()
-    color = {'normal': bamcd["red"][0], "multivariate_normal": bamcd["blue"][0]}
+    color = {'hapod': bamcd["red"][0], "heuristic": bamcd["blue"][0]}
 
     with PlottingContext(sys.argv, ["paper_onecol"]) as fig:
         fig.set_size_inches(4.773, 2.95)
@@ -22,7 +22,7 @@ if __name__ == "__main__":
                 data = np.loadtxt(infile.as_posix(), delimiter=",")
                 modes = data[:, 0]
                 error = data[:, 1]
-                ax.semilogy(modes, error, color=color[distribution], marker="o", label=name)
+                ax.semilogy(modes, error, color=color[name], marker="o", label=name)
 
         ax.legend(loc="best")
 

### 2024-03-12 src/beam/tasks.py

@@ -242,7 +242,7 @@ def task_loc_rom():
             yield {
                     "name": f"locrom_{name}_{distr}",
                     "file_dep": deps,
-                    "actions": ["python3 -m {} {} {} --output {}".format(module, distr, num_test, target)],
+                    "actions": ["python3 -m {} {} {} {} --output {}".format(module, distr, name, num_test, target)],
                     "targets": [target, beam.log_run_locrom(distr, name)],
                     "clean": True,
                     }

### 2024-03-12 paper/paper.tex
update all figures and paper

@@
 
-\subsubsection{Singular values of POD modes}
+\subsubsection{Distributed approximate POD}
 
@@
   \includegraphics[width=\linewidth]{../figures/beam/fig_loc_svals_left.pdf}
-  \caption{A really Awesome Image}\label{fig:loc_svals_left}
+  \caption{Singular valus left}\label{fig:loc_svals_left}
 \endminipage\hfill
@@
   \includegraphics[width=\linewidth]{../figures/beam/fig_loc_svals_inner.pdf}
-  \caption{A really Awesome Image}\label{fig:loc_svals_inner}
+  \caption{Singular valus inner}\label{fig:loc_svals_inner}
 \endminipage\hfill
@@
   \includegraphics[width=\linewidth]{../figures/beam/fig_loc_svals_right.pdf}
-  \caption{A really Awesome Image}\label{fig:loc_svals_right}
+  \caption{Singular valus right}\label{fig:loc_svals_right}
 \endminipage
@@
 
-\subsubsection{Projection Error}
-The projection error is computed for a test set for each configuration that was computed
-using the FOM. Each test set has size $n_{test}$.
+% \subsubsection{Projection Error}
+% The projection error is computed for a test set for each configuration that was computed
+% using the FOM. Each test set has size $n_{test}$.
 
@@
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_left.pdf}
-  \caption{A really Awesome Image}\label{fig:proj_error_left}
+  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_left_hapod.pdf}
+  \caption{Projection error hapod}\label{fig:proj_error_left_hapod}
 \endminipage\hfill
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_inner.pdf}
-  \caption{A really Awesome Image}\label{fig:proj_error_inner}
+  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_inner_hapod.pdf}
+  \caption{Projection error hapod}\label{fig:proj_error_inner_hapod}
 \endminipage\hfill
 \minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_right.pdf}
-  \caption{A really Awesome Image}\label{fig:proj_error_right}
+  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_right_hapod.pdf}
+  \caption{Projection error hapod}\label{fig:proj_error_right_hapod}
+\endminipage
+\end{figure}
+
+\subsubsection{Heuristic range finder}
+
+\begin{figure}[!htb]
+\minipage{0.32\textwidth}
+  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_left_heuristic.pdf}
+  \caption{Projection error }\label{fig:proj_error_left_heuristic}
+\endminipage\hfill
+\minipage{0.32\textwidth}
+  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_inner_heuristic.pdf}
+  \caption{Projection error }\label{fig:proj_error_inner_heuristic}
+\endminipage\hfill
+\minipage{0.32\textwidth}%
+  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_right_heuristic.pdf}
+  \caption{Projection error }\label{fig:proj_error_right_heuristic}
 \endminipage

### 2024-03-12 src/beam/decompose_pod_basis.py
remove unused files

@@ -1,105 +1,0 @@
-from pathlib import Path
-from mpi4py import MPI
-from dolfinx import fem, mesh
-from dolfinx.io import gmshio
-from basix.ufl import element
-import numpy as np
-
-from pymor.core.defaults import set_defaults
-from pymor.core.logger import getLogger
-from pymor.bindings.fenicsx import FenicsxVectorSpace
-
-from multi.misc import locate_dofs, x_dofs_vectorspace
-from multi.domain import RectangularSubdomain
-from multi.materials import LinearElasticMaterial
-from multi.problems import LinElaSubProblem
-from multi.basis_construction import compute_phi
-
-
-def main(args):
-    """decompose pod basis into coarse and fine scale parts"""
-    from .tasks import beam
-    set_defaults(
-        {
-            "pymor.core.logger.getLogger.filename": beam.log_decompose_pod_basis(
-                args.distribution, args.configuration
-            ),
-        }
-    )
-    logger = getLogger(Path(__file__).stem, level="INFO")
-
-    gdim = beam.gdim
-    domain, _, _ = gmshio.read_from_msh(
-        beam.unit_cell_grid.as_posix(), MPI.COMM_WORLD, gdim=gdim
-    )
-    omega = RectangularSubdomain(12, domain)
-    omega.create_coarse_grid(1)
-    omega.create_boundary_grids()
-
-    # ### FE spaces
-    degree = beam.fe_deg
-    fe = element("P", domain.basix_cell(), degree, shape=(gdim,))
-    V = fem.functionspace(omega.grid, fe)
-    source = FenicsxVectorSpace(V)
-
-    E = beam.youngs_modulus
-    NU = beam.poisson_ratio
-    mat = LinearElasticMaterial(gdim, E, NU, plane_stress=False)
-
-    # ### Problem on unit cell domain
-    problem = LinElaSubProblem(omega, V, phases=(mat,))
-    problem.setup_coarse_space()
-    problem.setup_edge_spaces()
-    problem.create_map_from_V_to_L()
-
-    # ### Full POD basis
-    pod_basis_data = np.load(beam.loc_pod_modes(args.distribution, args.configuration))
-    pod_basis = source.from_numpy(pod_basis_data)
-    logger.info(f"Size of POD basis: {len(pod_basis)}.")
-
-    # ### Coarse scale basis
-    vertices = omega.coarse_grid.topology.connectivity(2, 0).links(0)
[…]

### 2024-03-12 src/beam/definitions.py

@@ -315,7 +315,6 @@ class BeamProblem(MultiscaleProblemDefinition):
 
     def get_neumann(self, cell_index: Optional[int] = None):
         # is the same for all oversampling problems
-        # see range_approximation.py
         return None
 
     def get_kernel_set(self, cell_index: int) -> tuple[int, ...]:

### 2024-03-13 src/beam/definitions.py
task_hapod: add more targets

@@ -172,6 +172,12 @@ class BeamData:
         """length of each edge basis after rrf algo in hapod training"""
         return self.rf / f"hapod_rrf_bases_length_{distr}_{conf}.npz"
 
+    def hapod_table(self, conf: str) -> Path:
+        return self.rf / f"hapod_table_{conf}.csv"
+
+    def pod_data(self, distr: str, conf: str) -> Path:
+        return self.rf / f"pod_data_{distr}_{conf}.json"
+
     def fine_scale_edge_modes_npz(self, distr: str, conf: str, name: str) -> Path:
         """edge-restricted fine scale part of pod modes"""
         return self.rf / f"fine_scale_edge_modes_{distr}_{conf}_{name}.npz"

### 2024-03-13 src/beam/range_approx_edge.py

@@ -1,6 +1,7 @@
 """approximate the range of transfer operators for fixed parameter values"""
 
 import os
+import json
 from typing import Optional
 # from itertools import repeat
 from pathlib import Path
@@ -555,11 +556,15 @@ def main(args):
     pod_table.append(["Edge", "Number of Snapshots", "Number of POD modes", "Rel. Tolerance"])
     pod_modes = {}
     pod_svals = {}
+    pod_data = {}
+
     for edge, snapshot_data in snapshots.items():
         modes, svals = pod(snapshot_data, product=range_products[edge], modes=None, rtol=beam.pod_rtol)
         pod_modes[edge] = modes.to_numpy()
         pod_svals[edge] = svals
         pod_table.append([edge, len(snapshot_data), len(modes), beam.pod_rtol])
+        pod_data[edge] = (len(modes), len(snapshot_data))
+
     logger.info(format_table(pod_table, title="\nPOD of edge basis functions"))
 
     # write output: fine scale edge modes, singular values, rrf bases length
@@ -568,6 +573,9 @@ def main(args):
     )
     np.savez(beam.loc_singular_values_npz(args.distribution, args.configuration), **pod_svals)
     np.savez(beam.hapod_rrf_bases_length(args.distribution, args.configuration), **rrf_bases_length)
+    with beam.pod_data(args.distribution, args.configuration).open("w") as fh:
+        fh.write(json.dumps(pod_data))
+
 
 
 if __name__ == "__main__":

### 2024-03-13 src/beam/tasks.py

@@ -102,6 +102,7 @@ def task_hapod():
                     beam.fine_scale_edge_modes_npz(distr, config, "hapod"),
                     beam.loc_singular_values_npz(distr, config),
                     beam.hapod_rrf_bases_length(distr, config),
+                    beam.pod_data(distr, config),
                 ],
                 "clean": [rm_rf],
             }

### 2024-03-13 src/beam/generate_hapod_table.py
add generate_hapod_table.py

@@ -1,0 +1,58 @@
+import csv
+import json
+import numpy as np
+from collections import defaultdict
+
+
+def main(args):
+    from .tasks import beam
+
+    train_set_size = beam.lhs_options[args.configuration]["samples"]
+    num_testvecs = beam.rrf_num_testvecs
+    edges = ["bottom", "left", "right", "top"]
+    rrf_bases_length = np.load(beam.hapod_rrf_bases_length("normal", args.configuration))
+    avg_bases_length = {}
+    for edge in edges:
+        avg_bases_length[edge] = np.average(rrf_bases_length[edge])
+
+    with beam.pod_data("normal", args.configuration).open("r") as fh:
+        data = fh.read()
+        pod_data = json.loads(data)
+
+    basis_size = defaultdict(list)
+    num_snapshots = defaultdict(list)
+    for edge, data in pod_data.items():
+        basis_size[edge].append(data[0])
+        num_snapshots[edge].append(data[1])
+
+    table = []
+    top_row = ["Edge", "Basis size", "Snapshots", "Average basis size after RRF"]
+    table.append(top_row)
+    for edge in edges:
+        row = []
+        row.append(edge)
+        row.append(basis_size[edge])
+        row.append(num_snapshots[edge])
+        row.append(avg_bases_length[edge])
+        table.append(row)
+
+    with beam.hapod_table(args.configuration).open("w") as fh:
+
+        fh.write(f"# {train_set_size=}, {num_testvecs=}")
+        csv_writer = csv.writer(fh, delimiter=",")
+        for row in table:
+            csv_writer.writerow(row)
+
+
+if __name__ == "__main__":
+    import sys, argparse
+
+    parser = argparse.ArgumentParser()
+    parser.add_argument(
+        "configuration",
+        type=str,
+        help="The type of oversampling problem.",
+        choices=("inner", "left", "right"),
+    )
+    args = parser.parse_args(sys.argv[1:])
+    main(args)

### 2024-03-13 src/beam/generate_hapod_table.py
add hapod table

@@ -1,7 +1,6 @@
 import csv
 import json
 import numpy as np
-from collections import defaultdict
 
 
 def main(args):
@@ -13,17 +12,17 @@ def main(args):
     rrf_bases_length = np.load(beam.hapod_rrf_bases_length("normal", args.configuration))
     avg_bases_length = {}
     for edge in edges:
-        avg_bases_length[edge] = np.average(rrf_bases_length[edge])
+        avg_bases_length[edge] = np.around(np.average(rrf_bases_length[edge]), decimals=0)
 
     with beam.pod_data("normal", args.configuration).open("r") as fh:
         data = fh.read()
         pod_data = json.loads(data)
 
-    basis_size = defaultdict(list)
-    num_snapshots = defaultdict(list)
+    basis_size = {}
+    num_snapshots = {}
     for edge, data in pod_data.items():
-        basis_size[edge].append(data[0])
-        num_snapshots[edge].append(data[1])
+        basis_size[edge] = data[0]
+        num_snapshots[edge] = data[1]
 
     table = []
     top_row = ["Edge", "Basis size", "Snapshots", "Average basis size after RRF"]
@@ -38,7 +37,7 @@ def main(args):
 
     with beam.hapod_table(args.configuration).open("w") as fh:
 
-        fh.write(f"# {train_set_size=}, {num_testvecs=}")
+        fh.write(f"# {train_set_size=}, {num_testvecs=}\n")
         csv_writer = csv.writer(fh, delimiter=",")
         for row in table:
             csv_writer.writerow(row)

### 2024-03-13 src/beam/tasks.py

@@ -325,6 +325,37 @@ def task_fig_beam_sketch():
             }
 
 
+def task_hapod_table_csv():
+    """Beam example: Generate HAPOD table in csv format"""
+    module = f"src.{beam.name}.generate_hapod_table"
+    distr = "normal"
+    deps = []
+    for config in CONFIGS:
+        deps.append(beam.hapod_rrf_bases_length(distr, config))
+        deps.append(beam.pod_data(distr, config))
+        yield {
+                "name": f"{config}",
+                "file_dep": deps,
+                "actions": ["python3 -m {} {}".format(module, config)],
+                "targets": [beam.hapod_table(config)],
+                "clean": True,
+                }
+
+
+def task_compile_tables():
+    """Beam example: Compile standalone tables"""
+    sources = list((SRC / "tables").glob("*.tex"))
+    for src in sources:
+        yield {
+                "name": f"{src.stem}",
+                "file_dep": [src],
+                "actions": [f"latexmk -cd -pdf -outdir={ROOT / 'tables'} %(dependencies)s"],
+                "targets": [ROOT / "tables" / (src.stem + ".pdf")],
+                "clean": True,
+                }
+
+
+
 def task_paper():
     """Beam example: Compile Paper"""
     source = ROOT / "paper/paper.tex"

### 2024-03-13 src/beam/definitions.py
make table compilation work

@@ -178,6 +178,12 @@ class BeamData:
     def pod_data(self, distr: str, conf: str) -> Path:
         return self.rf / f"pod_data_{distr}_{conf}.json"
 
+    def heuristic_data(self, distr: str, conf: str) -> Path:
+        return self.rf / f"heuristic_data_{distr}_{conf}.json"
+
+    def heuristic_table(self, conf: str) -> Path:
+        return self.rf / f"heuristic_table_{conf}.csv"
+
     def fine_scale_edge_modes_npz(self, distr: str, conf: str, name: str) -> Path:
         """edge-restricted fine scale part of pod modes"""
         return self.rf / f"fine_scale_edge_modes_{distr}_{conf}_{name}.npz"

### 2024-03-13 src/beam/heuristic_range_approx.py

@@ -1,3 +1,4 @@
+import json
 import numpy as np
 from pathlib import Path
 
@@ -40,7 +41,9 @@ def lhs_design(beam, sampling_options):
     return training_set
 
 
-def solve_neumann_problem(beam, beam_problem, transfer_problem, mu, args, neumann_snapshots):
+def solve_neumann_problem(
+    beam, beam_problem, transfer_problem, mu, args, neumann_snapshots
+):
     neumann_problem = transfer_problem.problem
     neumann_problem.clear_bcs()
     omega = neumann_problem.domain
@@ -95,7 +98,10 @@ def solve_neumann_problem(beam, beam_problem, transfer_problem, mu, args, neuman
         U_orth = U_in
     else:
         U_orth = orthogonal_part(
-            U_in, transfer_problem.kernel, product=transfer_problem.range_product, orthonormal=True
+            U_in,
+            transfer_problem.kernel,
+            product=transfer_problem.range_product,
+            orthonormal=True,
         )
 
     # fine scale part of neumann snapshot
@@ -319,7 +325,9 @@ def approximate_range(
             edge_space = range_spaces[edge]
 
             # restrict the training sample to the edge
-            udofs = edge_space.from_numpy(U.dofs(transfer_problem.subproblem.V_to_L[edge]))
+            udofs = edge_space.from_numpy(
+                U.dofs(transfer_problem.subproblem.V_to_L[edge])
+            )
             coeff = udofs.dofs(edge_boundary_dofs[edge])
             ufine = udofs - coarse_basis[edge].lincomb(coeff)
 
@@ -335,10 +343,12 @@ def approximate_range(
 
         logger.info(f"Iteration: {num_solves}:\t{maxnorm=}.")
 
+    rrf_bases_length = {}
     with logger.block("Completed range approximation"):
         logger.info(f"Number of iterations: {num_solves}")
         for edge in edges:
             logger.info(f"Number of modes {edge}: {len(fine_scale_edge_bases[edge])}")
+            rrf_bases_length[edge] = len(fine_scale_edge_bases[edge])
 
     # Multiscale problem definition
     beam_problem = BeamProblem(beam.coarse_grid.as_posix(), beam.fine_grid.as_posix())
@@ -346,17 +356,36 @@ def approximate_range(
     for i in range(num_solves):
         mu = training_set[i]
         # FIXME avoid adding more functions if extension fails for i < num_solves
[…]

### 2024-03-13 src/beam/tasks.py

@@ -128,6 +128,7 @@ def task_heuristic_rrf():
                 "targets": [
                     beam.log_edge_range_approximation(distr, config, "heuristic"),
                     beam.fine_scale_edge_modes_npz(distr, config, "heuristic"),
+                    beam.heuristic_data(distr, config),
                 ],
                 "clean": [rm_rf],
             }
@@ -342,14 +343,37 @@ def task_hapod_table_csv():
                 }
 
 
+def task_heuristic_table_csv():
+    """Beam example: Generate heuristic table in csv format"""
+    module = f"src.{beam.name}.generate_heuristic_table"
+    distr = "normal"
+    deps = []
+    for config in CONFIGS:
+        deps.append(beam.heuristic_data(distr, config))
+        yield {
+                "name": f"{config}",
+                "file_dep": deps,
+                "actions": ["python3 -m {} {}".format(module, config)],
+                "targets": [beam.heuristic_table(config)],
+                "clean": True,
+                }
+
+
 def task_compile_tables():
     """Beam example: Compile standalone tables"""
     sources = list((SRC / "tables").glob("*.tex"))
     for src in sources:
+        data = []
+        name = src.stem
+        train, config = name.split("_")
+        if train.startswith("hapod"):
+            data.append(beam.hapod_table(config))
+        if train.startswith("heuristic"):
+            data.append(beam.heuristic_table(config))
         yield {
                 "name": f"{src.stem}",
-                "file_dep": [src],
-                "actions": [f"latexmk -cd -pdf -outdir={ROOT / 'tables'} %(dependencies)s"],
+                "file_dep": [src] + data,
+                "actions": [f"latexmk -cd -pdf -outdir={ROOT / 'tables'} {src}"],
                 "targets": [ROOT / "tables" / (src.stem + ".pdf")],
                 "clean": True,
                 }

### 2024-03-13 paper/paper.tex
paper: all paths relative to ROOT

@@
 \usepackage{amssymb}
+
+\usepackage{standalone}
+\usepackage{booktabs}
+\usepackage{pgfplotstable}
+\pgfplotsset{compat=1.18}
 
@@
     \begin{center}
-        \includegraphics[width=0.95\textwidth]{../figures/beam/beam_sketch.pdf}
+        \includegraphics[width=0.95\textwidth]{./figures/beam/beam_sketch.pdf}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=0.65\textwidth]{../figures/beam/unit_cell.png}
+        \includegraphics[width=0.65\textwidth]{./figures/beam/unit_cell.png}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=0.95\textwidth]{../figures/beam/global_domain.png}
+        \includegraphics[width=0.95\textwidth]{./figures/beam/global_domain.png}
     \end{center}
@@
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../figures/beam/fig_loc_svals_left.pdf}
+  \includegraphics[width=\linewidth]{./figures/beam/fig_loc_svals_left.pdf}
   \caption{Singular valus left}\label{fig:loc_svals_left}
@@
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../figures/beam/fig_loc_svals_inner.pdf}
+  \includegraphics[width=\linewidth]{./figures/beam/fig_loc_svals_inner.pdf}
   \caption{Singular valus inner}\label{fig:loc_svals_inner}
@@
 \minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{../figures/beam/fig_loc_svals_right.pdf}
+  \includegraphics[width=\linewidth]{./figures/beam/fig_loc_svals_right.pdf}
   \caption{Singular valus right}\label{fig:loc_svals_right}
@@
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_left_hapod.pdf}
+  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_left_hapod.pdf}
   \caption{Projection error hapod}\label{fig:proj_error_left_hapod}
@@
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_inner_hapod.pdf}
+  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_inner_hapod.pdf}
   \caption{Projection error hapod}\label{fig:proj_error_inner_hapod}
@@
 \minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_right_hapod.pdf}
+  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_right_hapod.pdf}
   \caption{Projection error hapod}\label{fig:proj_error_right_hapod}
@@
 \end{figure}
+
+\begin{table}
+    \caption{HAPOD inner}\label{tab:hapod_inner}
+    \input{./src/beam/tables/hapod_inner.tex}
+\end{table}
 
@@
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_left_heuristic.pdf}
+  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_left_heuristic.pdf}
   \caption{Projection error }\label{fig:proj_error_left_heuristic}
@@
 \minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_inner_heuristic.pdf}
+  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_inner_heuristic.pdf}
   \caption{Projection error }\label{fig:proj_error_inner_heuristic}
@@
 \minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{../figures/beam/fig_proj_error_right_heuristic.pdf}
+  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_right_heuristic.pdf}
   \caption{Projection error }\label{fig:proj_error_right_heuristic}
@@
 
+\begin{table}
+    \caption{Heuristic rrf inner}\label{tab:heuristic_inner}
+    \input{./src/beam/tables/heuristic_inner.tex}
+\end{table}
+
 \subsection{Localized ROM}
@@
 \begin{figure}[!htb]
-    \includegraphics[width=\linewidth]{../figures/beam/fig_loc_rom_error.pdf}
+    \includegraphics[width=\linewidth]{./figures/beam/fig_loc_rom_error.pdf}
     \caption{A really Awesome Image}\label{fig:loc_rom_error}

### 2024-03-13 src/beam/tasks.py

@@ -373,7 +373,7 @@ def task_compile_tables():
         yield {
                 "name": f"{src.stem}",
                 "file_dep": [src] + data,
-                "actions": [f"latexmk -cd -pdf -outdir={ROOT / 'tables'} {src}"],
+                "actions": [f"latexmk -pdf -outdir={ROOT / 'tables'} {src}"],
                 "targets": [ROOT / "tables" / (src.stem + ".pdf")],
                 "clean": True,
                 }
@@ -394,9 +394,9 @@ def task_paper():
             deps.append(beam.fig_proj_error(config, name))
     return {
         "file_dep": deps,
-        "actions": ["latexmk -cd -pdf %s" % source],
+        "actions": [f"latexmk -pdf -outdir={ROOT / 'paper'} {source}"],
         "targets": [source.with_suffix(".pdf")],
-        "clean": ["latexmk -cd -C %s" % source],
+        "clean": ["latexmk -C %s" % source],
     }
 
 

### 2024-03-14 paper/paper.tex
adjust figsizes

@@
 \section{INTRODUCTION}
+
+\makeatletter
+orig: \f@size
+\verb+\normalsize+ \normalsize \f@size
+\makeatother
+It seems that the font size is actually set to 11pt, although 12pt should be used in the body of the text.
 
@@
 \begin{figure}[!htb]
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_loc_svals_left.pdf}
+\minipage{0.49\textwidth}
+  \includegraphics[]{./figures/beam/fig_loc_svals_left.pdf}
   \caption{Singular valus left}\label{fig:loc_svals_left}
 \endminipage\hfill
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_loc_svals_inner.pdf}
+\minipage{0.49\textwidth}
+  \includegraphics[]{./figures/beam/fig_loc_svals_right.pdf}
+  \caption{Singular valus right}\label{fig:loc_svals_right}
+\endminipage\hfill\\%
+\begin{center}
+\minipage{0.49\textwidth}
+  \includegraphics[]{./figures/beam/fig_loc_svals_inner.pdf}
   \caption{Singular valus inner}\label{fig:loc_svals_inner}
-\endminipage\hfill
-\minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{./figures/beam/fig_loc_svals_right.pdf}
-  \caption{Singular valus right}\label{fig:loc_svals_right}
 \endminipage
+\end{center}
 \end{figure}
@@
 \begin{figure}[!htb]
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_left_hapod.pdf}
+\minipage{0.49\textwidth}
+  \includegraphics{./figures/beam/fig_proj_error_left_hapod.pdf}
   \caption{Projection error hapod}\label{fig:proj_error_left_hapod}
 \endminipage\hfill
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_inner_hapod.pdf}
+\minipage{0.49\textwidth}
+  \includegraphics{./figures/beam/fig_proj_error_right_hapod.pdf}
+  \caption{Projection error hapod}\label{fig:proj_error_right_hapod}
+\endminipage\hfill\\
+\begin{center}
+\minipage{0.49\textwidth}%
+  \includegraphics{./figures/beam/fig_proj_error_inner_hapod.pdf}
   \caption{Projection error hapod}\label{fig:proj_error_inner_hapod}
-\endminipage\hfill
-\minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_right_hapod.pdf}
-  \caption{Projection error hapod}\label{fig:proj_error_right_hapod}
 \endminipage
+\end{center}
 \end{figure}
@@
 \begin{table}
+    \centering
     \caption{HAPOD inner}\label{tab:hapod_inner}
@@
 \begin{figure}[!htb]
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_left_heuristic.pdf}
+\minipage{0.49\textwidth}
+  \includegraphics{./figures/beam/fig_proj_error_left_heuristic.pdf}
   \caption{Projection error }\label{fig:proj_error_left_heuristic}
 \endminipage\hfill
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_inner_heuristic.pdf}
+\minipage{0.49\textwidth}
+  \includegraphics{./figures/beam/fig_proj_error_right_heuristic.pdf}
+  \caption{Projection error }\label{fig:proj_error_right_heuristic}
+\endminipage\hfill\\
+\begin{center}
+\minipage{0.49\textwidth}%
+  \includegraphics{./figures/beam/fig_proj_error_inner_heuristic.pdf}
   \caption{Projection error }\label{fig:proj_error_inner_heuristic}
-\endminipage\hfill
-\minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_right_heuristic.pdf}
-  \caption{Projection error }\label{fig:proj_error_right_heuristic}
 \endminipage
+\end{center}
 \end{figure}
@@
 \begin{table}
+    \centering
     \caption{Heuristic rrf inner}\label{tab:heuristic_inner}
@@
 \begin{figure}[!htb]
-    \includegraphics[width=\linewidth]{./figures/beam/fig_loc_rom_error.pdf}
-    \caption{A really Awesome Image}\label{fig:loc_rom_error}
+    \centering
+    \includegraphics{./figures/beam/fig_loc_rom_error.pdf}
+    \caption{Local ROM error relative to FOM.}\label{fig:loc_rom_error}
 \end{figure}

### 2024-03-14 src/beam/definitions.py

@@ -97,6 +97,11 @@ class BeamData:
                 self.bases_path(distr, name).mkdir(exist_ok=True, parents=True)
 
     @property
+    def plotting_style(self) -> Path:
+        """eccomas proceedings mplstyle"""
+        return ROOT / "src/proceedings.mplstyle"
+
+    @property
     def rf(self) -> Path:
         """run folder"""
         return WORK / f"{self.name}"

### 2024-03-14 src/beam/plot_locrom_error.py

@@ -10,8 +10,7 @@ if __name__ == "__main__":
     bamcd = read_bam_colors()
     color = {'hapod': bamcd["red"][0], "heuristic": bamcd["blue"][0]}
 
-    with PlottingContext(sys.argv, ["paper_onecol"]) as fig:
-        fig.set_size_inches(4.773, 2.95)
+    with PlottingContext(sys.argv, [beam.plotting_style.as_posix()]) as fig:
         ax = fig.subplots()
         ax.set_xlabel("Number of fine scale basis functions per edge.")
         ax.set_ylabel("locROM error relative to FOM.")
@@ -25,6 +24,3 @@ if __name__ == "__main__":
                 ax.semilogy(modes, error, color=color[name], marker="o", label=name)
 
         ax.legend(loc="best")
-
-
-

### 2024-03-14 src/beam/plot_projerr.py

@@ -19,12 +19,10 @@ def main(args):
         "top": bamcd["yellow"][0],
     }
 
-    with PlottingContext([__file__, args.output], ["paper_onecol"]) as fig:
-        # figure.figsize: 4.773, 2.950
-        width = 4.773
-        height = 2.95
-        factor = 1.0
-        fig.set_size_inches(factor * width, factor * height)
+    with PlottingContext([__file__, args.output], [beam.plotting_style.as_posix()]) as fig:
+        width = 3.149 # 1/2 of the printing box width
+        height = 1.946
+        fig.set_size_inches(width, height)
         ax = fig.subplots()
 
         config = args.configuration
@@ -38,15 +36,10 @@ def main(args):
         ymin = []
 
         for edge, color in colors.items():
-            label = ""
-            if distr == "multivariate_normal":
-                label = f"correlated, {config}, {edge}"
-            if distr == "normal":
-                label = f"uncorrelated, {config}, {edge}"
 
             err = npz_err[edge]
             num_modes = np.arange(err.size)
-            ax.semilogy(num_modes, err, color=color, marker=marker, label=label)
+            ax.semilogy(num_modes, err, color=color, marker=marker, label=edge)
             ymin.append(err[-1])
             # ax.semilogy(
             #         data["num_dofs"], data["err"],

### 2024-03-20 src/beam/tables/minimization_data.tex
add minimization tables

\documentclass{standalone}

\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.18}

% Setup siunitx:
\sisetup{
  round-mode          = places, % Rounds numbers
  round-precision     = 2, % to 2 places
}

\begin{document}

\pgfplotstabletypeset[
  col sep=comma,
  % columns={Model,Iterations,Evaluations,Time,{Output J}},
  display columns/3/.style={column name={Time}, precision=3},
  display columns/4/.style={column name={Output $J$}, sci zerofill, precision=3},
  every head row/.style={
    before row={\toprule}, % have a rule at top
    % set unit for each column
    after row={
        & - & - & \si{\second} & \si{\newton\milli\metre}\\
    \midrule}
    },
  every last row/.style={after row=\bottomrule}, % rule at bottom
  column type=r,
  columns/Model/.style={string type,column type=l},
]{./work/beam/minimization_data.csv}

\end{document}

### 2024-03-20 src/beam/definitions.py

@@ -249,6 +249,14 @@ class BeamData:
         return self.rf / f"rom_minimization_data_{distr}_{name}.out"
 
     @property
+    def minimization_data_table(self) -> Path:
+        return self.rf / "minimization_data.csv"
+
+    @property
+    def minimization_comparison_table(self) -> Path:
+        return self.rf / "minimization_comparison.csv"
+
+    @property
     def fig_fom_opt(self) -> Path:
         return FIGURES / "fig_fom_opt.pdf"
 

### 2024-03-20 src/beam/generate_minimization_comparison_table.py

@@ -1,0 +1,49 @@
+import csv
+from pymor.core.pickle import load
+
+
+def main(args):
+    from .tasks import beam
+
+    with beam.fom_minimization_data.open("rb") as instream:
+        fom_data = load(instream)
+
+    with beam.rom_minimization_data("normal", "heuristic").open("rb") as instream:
+        rom_data = load(instream)
+
+    table = []
+    top_row = ["", "", "Beam example"]
+    table.append(top_row)
+
+    # true output
+    table.append(["True output", r"$J(\mu^{\ast})$", fom_data["J(mu_min)"]])
+
+    # reduced output
+    table.append(["Reduced output", r"$J_N(\mu_N^{\ast})$", rom_data["J_N(mu_N_min)"]])
+
+    # true output at reduced optimal solution
+    table.append([r"True output at $\mu_N^{\ast}$", r"$J(\mu_N^{\ast})$", rom_data["J(mu_N_min)"]])
+
+    # error in optimal solution μ
+    table.append(["Abs. error in optimal solution", r"$\norm{\mu_N^{\ast} - \mu^{\ast}}$", rom_data["abs_err_mu"]])
+
+    # error in output
+    table.append(["Abs. error in output", r"$\vert J_N(\mu_N^{\ast}) - J(\mu^{\ast}) \vert$", rom_data["abs_err_J"]])
+
+    # suboptimality
+    table.append(["Suboptimality", r"$\frac{J(\mu_N^{\ast}) - J(\mu^{\ast})}{J(\mu^{\ast})}$", rom_data["suboptimality"]])
+
+    with beam.minimization_comparison_table.open("w") as fh:
+        header = "# terminology,formula,value\n"
+        fh.write(header)
+        csv_writer = csv.writer(fh, delimiter=",")
+        for row in table:
+            csv_writer.writerow(row)
+
+
+if __name__ == "__main__":
+    import sys, argparse
+
+    parser = argparse.ArgumentParser()
+    args = parser.parse_args(sys.argv[1:])
+    main(args)

### 2024-03-20 src/beam/optimization.py

@@ -1,10 +1,10 @@
 """Solution of an optimization problem"""
 
-# check pymordemos/linear_optimization.py
-# - [x] optimize with the FOM using FD
-# - [x] optimize with the ROM using FD
-# - [ ] optimize with the FOM using pymor gradient computation (if possible)
-# - [ ] optimize with the ROM using pymor gradient computation (if possible)
+# NOTE
+# the below code was written based on the pyMOR Tutorial
+# "Model order reduction for PDE-constrained optimization problems"
+# See https://docs.pymor.org/2023-2-0/tutorial_optimization.html
+
 from mpi4py import MPI
 
 import numpy as np
@@ -16,9 +16,6 @@ def main(args):
     """solve optimization problem for different models"""
     from .tasks import beam
     from .fom import discretize_fom
-
-    # TODO add logger
-    # TODO gather output data
 
     # ### Build FOM
     fom = discretize_fom(beam)
@@ -42,26 +39,26 @@ def main(args):
     }
 
     num_subdomains = beam.nx * beam.ny
-    initial_guess = fom.parameters.parse([5. for _ in range(num_subdomains)])
+    initial_guess = fom.parameters.parse([5.0 for _ in range(num_subdomains)])
     parameter_space = fom.parameters.space(beam.mu_range)
     bounds = [parameter_space.ranges["E"] for _ in range(num_subdomains)]
 
-    mu_min = fom.parameters.parse([0.1 for _ in range(num_subdomains)])
-    mu_max = fom.parameters.parse([10.0 for _ in range(num_subdomains)])
+    # mu_min = fom.parameters.parse([0.1 for _ in range(num_subdomains)])
+    # mu_max = fom.parameters.parse([10.0 for _ in range(num_subdomains)])
 
     # values of compliance using weights ω_i = 0.0!
     # C_mu_min = fom.output(mu_min)[0, 0] # 507.808
     # C_mu_max = fom.output(mu_max)[0, 0] # 5.07808
-    J_mu_min = fom.output(mu_min)[0, 0]
-    J_mu_max = fom.output(mu_max)[0, 0]
-    print(f"{J_mu_min=}")
-    print(f"{J_mu_max=}")
+    # J_mu_min = fom.output(mu_min)[0, 0] # type: ignore
+    # J_mu_max = fom.output(mu_max)[0, 0] # type: ignore
+    # print(f"FOM {J_mu_min=}")
+    # print(f"FOM {J_mu_max=}")
 
[…]

### 2024-03-20 src/beam/tasks.py

@@ -416,19 +416,50 @@ def task_heuristic_table_csv():
         }
 
 
+def task_minimization_data_csv():
+    """Beam example: Generate minimization data table in csv format"""
+    module = f"src.{beam.name}.generate_minimization_data_table"
+    deps = [beam.fom_minimization_data, beam.rom_minimization_data("normal", "heuristic")]
+    targets = [beam.minimization_data_table]
+    return {
+            "file_dep": deps,
+            "actions": ["python3 -m {}".format(module)],
+            "targets": targets,
+            "clean": True,
+            }
+
+
+def task_minimization_comparison_csv():
+    """Beam example: Generate minimization comparison table in csv format"""
+    module = f"src.{beam.name}.generate_minimization_comparison_table"
+    deps = [beam.fom_minimization_data, beam.rom_minimization_data("normal", "heuristic")]
+    targets = [beam.minimization_comparison_table]
+    return {
+            "file_dep": deps,
+            "actions": ["python3 -m {}".format(module)],
+            "targets": targets,
+            "clean": True,
+            }
+
+
 def task_compile_tables():
     """Beam example: Compile standalone tables"""
     sources = list((SRC / "tables").glob("*.tex"))
     for src in sources:
         data = []
         name = src.stem
-        train, config = name.split("_")
-        if train.startswith("hapod"):
+        # FIXME: better way to handle deps?
+        if name.startswith("hapod"):
+            _, config = name.split("_")
             data.append(beam.hapod_table(config))
-        if train.startswith("heuristic"):
+        elif name.startswith("heuristic"):
+            _, config = name.split("_")
             data.append(beam.heuristic_table(config))
+        elif name.startswith("minimization"):
+            data.append(beam.minimization_data_table)
+            data.append(beam.minimization_comparison_table)
         yield {
-            "name": f"{src.stem}",
+            "name": f"{name}",
             "file_dep": [src] + data,
             "actions": [f"latexmk -pdf -outdir={ROOT / 'tables'} {src}"],
             "targets": [ROOT / "tables" / (src.stem + ".pdf")],

### 2024-03-20 paper/paper.tex

@@
 \usepackage{amssymb}
+\usepackage{physics}
+\usepackage{siunitx}
 
@@
 
+\subsection{Optimization problem}
+
+\begin{table}
+    \caption{Result of the optimization with FOM and ROM.}\label{tab:minimization_data}
+    \centering
+    \input{./src/beam/tables/minimization_data.tex}
+\end{table}
+
+\begin{table}
+    \caption{Comparison of reduced optimal solution $\mu_N^{\ast}$ and true optimal solution $\mu^{\ast}$.}\label{tab:minimization_comparison}
+    \centering
+    \input{./src/beam/tables/minimization_comparison.tex}
+\end{table}
+
 \section{FORMAT OF REFERENCES}

### 2024-03-20 src/beam/tables/minimization_comparison.tex

\documentclass{standalone}

\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{pgfplotstable}
\usepackage{amsmath,physics}
\pgfplotsset{compat=1.18}

% Setup siunitx:
\sisetup{
  round-mode          = places, % Rounds numbers
  round-precision     = 2, % to 2 places
}

\begin{document}

\pgfplotstabletypeset[
  % FIXME using column type {S} does not lead to desired result 
  % columns/{Beam example}/.style={
  %     column type={S},
  %     sci zerofill,
  %     precision=3,
  % },
  multicolumn names,
  col sep=comma,
  column type=r,
  % columns={Model,Iterations,Evaluations,Time,{Output J}},
  display columns/0/.style={string type, column name={}, column type=l},
  display columns/1/.style={string type, column name={}, column type=l},
  display columns/2/.style={column name={Beam example}, sci zerofill, precision=4},
  every head row/.style={
    before row={\toprule}, % have a rule at top
    after row={\midrule},
    },
  every last row/.style={after row=\bottomrule}, % rule at bottom
]{./work/beam/minimization_comparison.csv}

\end{document}

### 2024-07-04 paper/paper.tex
[paper][intro] initial commit

--- 2024-03-13 paper/paper.tex
@@
 \usepackage{graphicx}
-\usepackage{hyperref}
+\usepackage{url}
+% \usepackage{hyperref}
 \usepackage{amsmath}
@@
 \usepackage{amssymb}
+\usepackage{bm}
+\usepackage{physics}
+% \usepackage{siunitx}
 
@@
 
-\title{INSTRUCTIONS TO PREPARE A FULL PAPER FOR THE 9th EUROPEAN CONGRESS ON COMPUTATIONAL METHODS IN APPLIED SCIENCES AND ENGINEERING\break ECCOMAS CONGRESS 2024}
+\usepackage[commentmarkup=uwave]{changes}
+\usepackage{xspace}
 
-\author{FIRST A. AUTHOR$^1$, SECOND B. AUTHOR$^2$ AND THIRD C. AUTHOR$^1$}
+\definechangesauthor[color=blue, name={Philipp Diercks}]{pd}
+\definechangesauthor[color=red, name={Joerg F. Unger}]{jfu}
+\definechangesauthor[color=orange, name={Annika Robens-Radermacher}]{arr}
 
-\heading{First A. Author, Second B. Author and Third C. Author}
+% custom commands
+\newcommand{\ie}{i.\,e.\@\xspace}
+\newcommand{\eg}{e.\,g.\@\xspace}
 
-\address{$^{1}$ International Center for Numerical Methods in Engineering (CIMNE)\\
-Universidad Polit\'{e}cnica de Catalu\~{n}a\\
-Campus Norte UPC, 08034 Barcelona, Spain\\
-e-mail: congress@cimne.upc.edu, www.cimne.com
+\title{AN EFFICIENT LOCALIZED MODEL ORDER REDUCTION FRAMEWORK FOR THE SHAPE OPTIMIZATION OF ADDITIVELY MANUFACTURED LATTICE STRUCTURES\break ECCOMAS CONGRESS 2024}
+
+\author{Philipp Diercks$^{1}$, Karen Veroy$^{2}$, Annika Robens-Radermacher$^{1}$ and Jörg F. Unger$^{1}$}
+
+\heading{Philipp Diercks, Karen Veroy, Annika Robens-Radermacher and Jörg F. Unger}
+
+\address{$^{1}$ Bundesanstalt für Materialforschung und -prüfung, Unter den Eichen 87, 12205 Berlin, \{philipp.diercks, annika.robens-radermacher, joerg.unger\}@bam.de, \url{www.bam.de}
 \and
-$^{2}$ Spanish Association for Numerical Methods in Engineering (SEMNI)\\
-Edificio C1, Campus Norte UPC\\
-Gran Capit\'{a}n s/n, 08034 Barcelona, Spain\\
-email: semni@cimne.upc.edu, www.semni.org
-}
+$^{2}$ Centre for Analysis, Scientific Computing and Applications (CASA), Department of Mathematics and Computer Science, TU Eindhoven, P.O.
+Box 513, 5600 MB Eindhoven, The Netherlands, k.p.veroy@tue.nl}
 
-\keywords{Computational Mechanics, FEM, Contact Problems}
+\keywords{Multiscale methods, Domain decomposition methods, Model order reduction, parameterized PDEs}
 
-\abstract{This document provides information and instructions for
-preparing a Full Paper to be included in the Proceedings
-of {\it ECCOMAS CONGRESS 2024}. }
+% TODO abstract
+\abstract{Add abstract.}
 
@@
 
-\section{INTRODUCTION}
+\section{INTRODUCTION}%
+\label{sec:introduction}
 
-All the participants whose Abstract has been accepted for presentation at the ECCOMAS 2022 Congress are kindly requested to submit the Full Paper electronically via the web page of the \href{https://eccomas2024.org/lisbon}{Congress}, before \textbf{July 1st, 2024}.
-Congress Proceedings (including full papers only) will be published on Scipedia with DOI number after the congress and will be submitted for indexation to SCOPUS database.
-Submission of the full paper is not mandatory.
-The Full Paper should be written following the format of macros for submission.
-The file has to be translated into Portable Document Format (PDF) before submission via the Conference site.
-The organizers do not commit themselves to include in the Proceedings any Full Paper received later than the above-mentioned deadline.
-The corresponding author should be the speaker, and is expected to register and pay his registration fee during the advance period (before \textbf{April 1st, 2024}) for the paper to be included in the technical programme of the ECCOMAS 2024.
+Additive manufacturing (AM), commonly known as 3D printing, is a manufacturing technique that allows for the production of a wide range of structures and complex geometries.
+The objects are build successively by adding material layer by layer from three-dimensional models.
+The technology offers numerous advantages over conventional manufacturing, including greater design flexibility, reduced material waste, and the possibility to produce complex structures with tailored material properties.
+It has been used in a wide range of applications, including aerospace, biomechanical, automotive, and construction industries \comment[id=pd]{TODO:  add citations}.%~\cite{duplessis2022properties,wu2016critical}.
 
-\section{METHODS}
+A common engineering practice is to optimize the geometry of a structure by an iterative process, in which an objective function is minimized by systematically choosing parameter values $\bm\mu$ (design variables) and computing the value of the function.
+In structural mechanics, a prominent example of an objective function is mass or compliance which is a function of the parameter-dependent displacement field $\bm{u}(\bm\mu)$.
+In each iteration of the optimization process, the geometry is manually changed in the CAD model and the high fidelity finite element (FE) model, also called full order model (FOM), is simply re-evaluated.
+While this approach is not only an ineffective use of resources, it is also infeasible when solving the FE model is a computationally demanding task, \eg{} in multiscale or large-scale industrial applications.
+Due to the prohibitive cost of even a single FOM solution, the iterative design process cannot be performed using direct numerical simulations.
+\comment[id=pd]{TODO: add suitable refs for localized MOR}
+Therefore, we propose a framework based on localized, also called component-based (CB), parametric model order reduction (pMOR).
+The main idea is to precompute, in a localized manner, empirical basis functions which approximate the solution for some part of the domain without the need to solve the global FOM even once.
+The global approximation is obtained by a suitable coupling of the local reduced spaces spanned by the aforementioned basis functions, in which one naturally relies on domain decomposition (DD) strategies.
+In particular, regarding lattice structures, this approach allows to take advantage of the repetitiveness of the lattice, such that a computation of the local basis is required only for few components, \ie{} unit cells.
 
-\section{RESULTS}
+The computation of the local basis is an essential task in localized pMOR and is often done using the concept of oversampling\comment[id=pd]{cite Hou}.
+In this approach, the target subdomain $\varOmega_{\mathrm{in}}$, \ie{} that part of the domain for which one would like to construct basis functions, is extended and boundary conditions are prescribed on the boundary of the larger so-called oversampling domain $\varOmega$ to explore possible solutions.
+In the literature, this oversampling problem is also expressed in terms of a \textit{transfer operator} that maps the values on the boundary $\partial\varOmega$ to the unknown solution restricted to the target subdomain $\bm{u}(\bm{\mu})\vert_{\varOmega_{\mathrm{in}}}$.
+% TODO continue with how to solve the oversampling problem, leading up to range finder
 
-\subsection{Problem setup}
+\section{METHOD}%
+\label{sec:method}
 
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=0.95\textwidth]{./figures/beam/beam_sketch.pdf}
-    \end{center}
-    \caption{Schematic representation of the structural problem. A simply supported beam is subjected to a constant load.}\label{fig:beam_sketch}
-\end{figure}
+\subsection{Auxiliary Problem} % (fold)
+\label{sub:Auxiliary Problem}
 
+% subsection Auxiliary Problem (end)
+
+\subsection{Construction of local approximation spaces} % (fold)
+\label{sub:Construction of local approximation spaces}
+
+% subsection Construction of local approximation spaces (end)
+
+\subsubsection{Randomized range finder and POD} % (fold)
+\label{sec:Randomized range finder and POD}
+
+% subsubsection Randomized range finder and POD (end)
+
+\subsubsection{Heuristic randomized range finder} % (fold)
+\label{sec:Heuristic randomized range finder}
+
+% subsubsection Heuristic randomized range finder (end)
+
+\subsection{Construction of a global approximation} % (fold)
+\label{sub:Construction of a global approximation}
+
+% subsection Construction of a global approximation (end)
+
+\subsection{Online efficiency} % (fold)
+\label{sub:Online efficiency}
+
+% subsection Online efficiency (end)
+
+\section{NUMERICAL EXPERIMENTS} % (fold)
+\label{sec:numerical experiments}
+
+% section numerical experiments (end)
+
+\subsection{Graded concrete slab} % (fold)
+\label{sub:Graded concrete slab}
+
+% \begin{figure}
+%     \begin{center}
+%         \includegraphics[width=0.95\textwidth]{./figures/beam/beam_sketch.pdf}
+%     \end{center}
+%     \caption{Schematic representation of the structural problem. A simply supported beam is subjected to a constant load.}\label{fig:beam_sketch}
+% \end{figure}
+%
 % TODO placement
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=0.65\textwidth]{./figures/beam/unit_cell.png}
-    \end{center}
-    \caption{Unit cell domain partition.}\label{fig:unit_cell_domain}
-\end{figure}
+% \begin{figure}
+%     \begin{center}
+%         \includegraphics[width=0.65\textwidth]{./figures/beam/unit_cell.png}
+%     \end{center}
+%     \caption{Unit cell domain partition.}\label{fig:unit_cell_domain}
+% \end{figure}
+%
+% \begin{figure}
+%     \begin{center}
+%         \includegraphics[width=0.95\textwidth]{./figures/beam/global_domain.png}
+%     \end{center}
+%     \caption{The partition of the global domain. The colors indicate the variation in Young's modulus for each subdomain.}\label{fig:global_domain}
+% \end{figure}
+%
+% subsection Graded concrete slab (end)
 
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=0.95\textwidth]{./figures/beam/global_domain.png}
-    \end{center}
-    \caption{The partition of the global domain. The colors indicate the variation in Young's modulus for each subdomain.}\label{fig:global_domain}
-\end{figure}
+\subsection{Projection error} % (fold)
+\label{sub:Projection error}
 
-\subsection{Basis Construction}
-For the beam problem, there are three oversampling problems to consider (left, inner, right).
-For each of the associated parametric transfer operators, $n_{train}$ parameter samples are chosen, and
-the range for each of these (fixed) transfer operators is approximated via random sampling. In
-the sampling \textit{normal} or \textit{multivariate normal} distribution is used.
-The range approximation of the $n_{train}$ transfer operators yields $n_{train}$ sets of
-basis functions, which are further compressed via POD, to obtain the final parameter
-independent basis functions (POD modes).
+% subsection Projection error (end)
 
-\subsubsection{Distributed approximate POD}
+\subsection{Reduced order model validation} % (fold)
+\label{sub:Reduced order model validation}
 
-\begin{figure}[!htb]
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_loc_svals_left.pdf}
-  \caption{Singular valus left}\label{fig:loc_svals_left}
-\endminipage\hfill
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_loc_svals_inner.pdf}
-  \caption{Singular valus inner}\label{fig:loc_svals_inner}
-\endminipage\hfill
-\minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{./figures/beam/fig_loc_svals_right.pdf}
-  \caption{Singular valus right}\label{fig:loc_svals_right}
-\endminipage
-\end{figure}
+% subsection Reduced order model validation (end)
 
-% \subsubsection{Projection Error}
-% The projection error is computed for a test set for each configuration that was computed
-% using the FOM. Each test set has size $n_{test}$.
+\subsection{Shape optimization} % (fold)
+\label{sub:Shape optimization}
 
-\begin{figure}[!htb]
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_left_hapod.pdf}
-  \caption{Projection error hapod}\label{fig:proj_error_left_hapod}
-\endminipage\hfill
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_inner_hapod.pdf}
-  \caption{Projection error hapod}\label{fig:proj_error_inner_hapod}
-\endminipage\hfill
-\minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_right_hapod.pdf}
-  \caption{Projection error hapod}\label{fig:proj_error_right_hapod}
-\endminipage
-\end{figure}
+% subsection Shape optimization (end)
 
-\begin{table}
-    \caption{HAPOD inner}\label{tab:hapod_inner}
-    \input{./src/beam/tables/hapod_inner.tex}
-\end{table}
+\section{CONCLUSIONS} % (fold)
+\label{sec:conclusions}
 
-\subsubsection{Heuristic range finder}
+% section CONCLUSION (end)
 
-\begin{figure}[!htb]
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_left_heuristic.pdf}
-  \caption{Projection error }\label{fig:proj_error_left_heuristic}
-\endminipage\hfill
-\minipage{0.32\textwidth}
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_inner_heuristic.pdf}
-  \caption{Projection error }\label{fig:proj_error_inner_heuristic}
-\endminipage\hfill
-\minipage{0.32\textwidth}%
-  \includegraphics[width=\linewidth]{./figures/beam/fig_proj_error_right_heuristic.pdf}
-  \caption{Projection error }\label{fig:proj_error_right_heuristic}
-\endminipage
-\end{figure}
+% \begin{figure}[!htb]
+% \minipage{0.49\textwidth}
+%   \includegraphics[]{./figures/beam/fig_loc_svals_left.pdf}
+%   \caption{Singular valus left}\label{fig:loc_svals_left}
+% \endminipage\hfill
+% \minipage{0.49\textwidth}
+%   \includegraphics[]{./figures/beam/fig_loc_svals_right.pdf}
+%   \caption{Singular valus right}\label{fig:loc_svals_right}
+% \endminipage\hfill\\%
+% \begin{center}
+% \minipage{0.49\textwidth}
+%   \includegraphics[]{./figures/beam/fig_loc_svals_inner.pdf}
+%   \caption{Singular valus inner}\label{fig:loc_svals_inner}
+% \endminipage
+% \end{center}
+% \end{figure}
 
-\begin{table}
-    \caption{Heuristic rrf inner}\label{tab:heuristic_inner}
-    \input{./src/beam/tables/heuristic_inner.tex}
-\end{table}
+% \begin{figure}[!htb]
+% \minipage{0.49\textwidth}
+%   \includegraphics{./figures/beam/fig_proj_error_left_hapod.pdf}
+%   \caption{Projection error hapod}\label{fig:proj_error_left_hapod}
+% \endminipage\hfill
+% \minipage{0.49\textwidth}
+%   \includegraphics{./figures/beam/fig_proj_error_right_hapod.pdf}
+%   \caption{Projection error hapod}\label{fig:proj_error_right_hapod}
+% \endminipage\hfill\\
+% \begin{center}
+% \minipage{0.49\textwidth}%
+%   \includegraphics{./figures/beam/fig_proj_error_inner_hapod.pdf}
+%   \caption{Projection error hapod}\label{fig:proj_error_inner_hapod}
+% \endminipage
+% \end{center}
+% \end{figure}
 
-\subsection{Localized ROM}
+% \begin{table}
+%     \centering
+%     \caption{HAPOD inner}\label{tab:hapod_inner}
+%     \input{./src/beam/tables/hapod_inner.tex}
+% \end{table}
 
-\begin{figure}[!htb]
-    \includegraphics[width=\linewidth]{./figures/beam/fig_loc_rom_error.pdf}
-    \caption{A really Awesome Image}\label{fig:loc_rom_error}
-\end{figure}
+% \begin{figure}[!htb]
+% \minipage{0.49\textwidth}
+%   \includegraphics{./figures/beam/fig_proj_error_left_heuristic.pdf}
+%   \caption{Projection error }\label{fig:proj_error_left_heuristic}
+% \endminipage\hfill
+% \minipage{0.49\textwidth}
+%   \includegraphics{./figures/beam/fig_proj_error_right_heuristic.pdf}
+%   \caption{Projection error }\label{fig:proj_error_right_heuristic}
+% \endminipage\hfill\\
+% \begin{center}
+% \minipage{0.49\textwidth}%
+%   \includegraphics{./figures/beam/fig_proj_error_inner_heuristic.pdf}
+%   \caption{Projection error }\label{fig:proj_error_inner_heuristic}
+% \endminipage
+% \end{center}
+% \end{figure}
 
-\section{FORMAT OF REFERENCES}
+% \begin{table}
+%     \centering
+%     \caption{Heuristic rrf inner}\label{tab:heuristic_inner}
+%     \input{./src/beam/tables/heuristic_inner.tex}
+% \end{table}
 
-References should be quoted in the text by superscript
-numbers \cite{McQuerry18,nssga,Oden18} and grouped together at the end of the Full Paper in numerical order as shown in these instructions.
+% \begin{figure}[!htb]
+%     \centering
+%     \includegraphics{./figures/beam/fig_loc_rom_error.pdf}
+%     \caption{Local ROM error relative to FOM.}\label{fig:loc_rom_error}
+% \end{figure}
 
-\section{CONCLUSIONS}
+% \begin{table}
+%     \caption{Result of the optimization with FOM and ROM.}\label{tab:minimization_data}
+%     \centering
+%     \input{./src/beam/tables/minimization_data.tex}
+% \end{table}
 
-\begin{itemize}
-\item[-] Full Papers in format for publication should be submitted electronically via the web page of the Conference, before \textbf{July 1st, 2024}.
-They must be converted to  Portable Document Format (PDF) before submission.
-The maximum size of the file is 4 Mb.
-\end{itemize}
+% \begin{table}
+%     \caption{Comparison of reduced optimal solution $\mu_N^{\ast}$ and true optimal solution $\mu^{\ast}$.}\label{tab:minimization_comparison}
+%     \centering
+%     \input{./src/beam/tables/minimization_comparison.tex}
+% \end{table}
 
+% TODO use biblatex instead if the template allows that
 \begin{thebibliography}{99}

### 2024-08-02 paper/paper.tex
[paper][intro] wip

@@
 \usepackage{url}
-% \usepackage{hyperref}
 \usepackage{amsmath}
@@
 \usepackage{physics}
-% \usepackage{siunitx}
 
@@
 \pgfplotsset{compat=1.18}
+
+\usepackage[%
+backend=biber,
+style=numeric, %alphabetic, numeric
+giveninits=true,
+natbib=true,
+url=false,
+doi=true,
+eprint=false,
+isbn=false,
+defernumbers=true,
+labelnumber,
+hyperref=false,
+maxbibnames=3,
+sorting=none,%remove this to have things sorted, e.g. use style=alphabetic
+]{biblatex}
+\addbibresource{main.bib}
+\addbibresource{references.bib}
 
@@
 
+% heading is supposed to contain only the authors (on all pages except the first page)
 \heading{Philipp Diercks, Karen Veroy, Annika Robens-Radermacher and Jörg F. Unger}
@@
 
-\keywords{Multiscale methods, Domain decomposition methods, Model order reduction, parameterized PDEs}
+\keywords{Multiscale methods, Domain decomposition methods, Model order reduction, parameterized PDEs, Shape optimization}
 
 % TODO abstract
-\abstract{Add abstract.}
+\abstract{\comment[id=pd]{TODO: add abstract}.}
 
@@
 The technology offers numerous advantages over conventional manufacturing, including greater design flexibility, reduced material waste, and the possibility to produce complex structures with tailored material properties.
-It has been used in a wide range of applications, including aerospace, biomechanical, automotive, and construction industries \comment[id=pd]{TODO:  add citations}.%~\cite{duplessis2022properties,wu2016critical}.
+It has been used in a wide range of applications, including aerospace, biomechanical, automotive, and construction industries~\cite{Plessis2022Properties,Wu2016critical}.
 
@@
 Due to the prohibitive cost of even a single FOM solution, the iterative design process cannot be performed using direct numerical simulations.
-\comment[id=pd]{TODO: add suitable refs for localized MOR}
 Therefore, we propose a framework based on localized, also called component-based (CB), parametric model order reduction (pMOR).
@@
 The global approximation is obtained by a suitable coupling of the local reduced spaces spanned by the aforementioned basis functions, in which one naturally relies on domain decomposition (DD) strategies.
+For a review of concepts in localized model order reduction it is referred to~\cite{BuhrReview}.
 In particular, regarding lattice structures, this approach allows to take advantage of the repetitiveness of the lattice, such that a computation of the local basis is required only for few components, \ie{} unit cells.
 
-The computation of the local basis is an essential task in localized pMOR and is often done using the concept of oversampling\comment[id=pd]{cite Hou}.
+The computation of the local basis is an essential task in localized pMOR and is often done using the concept of oversampling~\cite{Hou1997Multiscale}.
 In this approach, the target subdomain $\varOmega_{\mathrm{in}}$, \ie{} that part of the domain for which one would like to construct basis functions, is extended and boundary conditions are prescribed on the boundary of the larger so-called oversampling domain $\varOmega$ to explore possible solutions.
-In the literature, this oversampling problem is also expressed in terms of a \textit{transfer operator} that maps the values on the boundary $\partial\varOmega$ to the unknown solution restricted to the target subdomain $\bm{u}(\bm{\mu})\vert_{\varOmega_{\mathrm{in}}}$.
-% TODO continue with how to solve the oversampling problem, leading up to range finder
+In the literature, this oversampling problem is also expressed in terms of a \textit{transfer operator} $\bm{T}$ that maps the values on the boundary $\partial\varOmega$ to the unknown solution restricted to the target subdomain $\bm{u}(\bm{\mu})\vert_{\varOmega_{\mathrm{in}}}$.
+The construction of (optimal) local approximation spaces then comprises the calculation of the left singular vectors of this transfer operator~\cite{Babuska2011Optimal,Smetana2016Optimal}.
+The direct calculation via eigenvalue problems is, however, computationally expensive and the range of the transfer operator, and thus the optimal local approximation spaces can be efficiently approximated by random sampling~\cite{Buhr2018Randomized}.
+Here, the authors treat non-parametrized PDEs and to the authors knowledge the extension to the parametric setting for linear problems has not been done yet.
+In~\cite{Smetana2016Optimal}, the authors propose a spectral greedy algorithm to construct local approximation spaces.
+Taddei and Patera~\cite{Taddei2018Localization} propose a combination of transfer eigenproblems and proper orthogonal decomposition (POD).
+For parameterized nonlinear elliptic PDEs Smetana and Taddei~\cite{Smetana2023Localized} present a randomized local training procedure with global enrichment.
+
+The contributions of the present work are given as follows.
+First, as references~\cite{Smetana2016Optimal,Taddei2018Localization} do not make use of range approximation via random sampling, suitable training strategies to construct local approximation spaces for parameterized linear problems via random sampling are investigated.
+In particular, the approach given in~\cite{Taddei2018Localization}, identified as a \textit{distributed approximate POD} (see also~\cite{Himpe2018Hierarchical}), is compared to a \textit{heuristic randomized range finder} developed in this work.
+Second, a framework for the shape optimization of lattice structures is proposed.
+It combines the aforementioned algorithms to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final reduced order model.
+Furthermore, the global approximation space is constructed from the local spaces using the generalized finite element method (GFEM)~\cite{Babuska2004Generalized}.
+
+% TODO add paragraph outlining structure of the text
 
@@
 
-\subsubsection{Randomized range finder and POD} % (fold)
+\subsubsection{Randomized range finder and proper orthogonal decomposition} % (fold)
 \label{sec:Randomized range finder and POD}
@@
 
-\subsection{Projection error} % (fold)
-\label{sub:Projection error}
+\subsection{Projection error study} % (fold)
+\label{sub:Projection error study}
 
-% subsection Projection error (end)
+% subsection Projection error study (end)
 
@@
 
-% TODO use biblatex instead if the template allows that
-\begin{thebibliography}{99}
-\bibitem{McQuerry18}  McQuerry, M.
-2018.
-“Validation of a Clothing Heat Transfer Model in Nonisothermal Test Conditions.” J Test Eval 46, no.
-1 (January/February): 1–7.
-http://doi.org/10.1520/JTE20170073
-
-\bibitem{nssga}   National Stone, Sand, and Gravel Association (NSSGA). 2013. The Aggregates Handbook, 2nd ed. Englewood, CO: Society of Mining, Metallurgy, and Exploration.
-
-\bibitem{Oden18}  Oden, C.
-P., C.
-L.
-Ho, and H.
-F.
-Kashani.
-2018.
-“Man-Portable Real-Time Ballast Inspection Device Using Ground-Penetrating Radar.” In Railroad Ballast Testing and Properties, edited by T.
-Stark, R.
-Swan, and R.
-Szecsy, 77–104.
-West Conshohocken, PA: ASTM International.
-http://doi.org/10.1520/STP160520170023
-
-\end{thebibliography}
+% references
+\printbibliography[title=REFERENCES]
 

### 2024-08-06 paper/paper.tex
add last paragraph of introduction

@@
 \usepackage{physics}
+\usepackage{cleveref}
 
@@
 The direct calculation via eigenvalue problems is, however, computationally expensive and the range of the transfer operator, and thus the optimal local approximation spaces can be efficiently approximated by random sampling~\cite{Buhr2018Randomized}.
-Here, the authors treat non-parametrized PDEs and to the authors knowledge the extension to the parametric setting for linear problems has not been done yet.
-In~\cite{Smetana2016Optimal}, the authors propose a spectral greedy algorithm to construct local approximation spaces.
-Taddei and Patera~\cite{Taddei2018Localization} propose a combination of transfer eigenproblems and proper orthogonal decomposition (POD).
+Here, the authors treat non-parametrized partial differential equations (PDEs) and to the authors knowledge the extension to the parametric setting for linear problems has not been done yet.
+In~\cite{Smetana2016Optimal}, the authors propose a spectral greedy algorithm to construct parameter-independent local approximation spaces and Taddei and Patera~\cite{Taddei2018Localization} propose a combination of transfer eigenproblems and proper orthogonal decomposition (POD).
 For parameterized nonlinear elliptic PDEs Smetana and Taddei~\cite{Smetana2023Localized} present a randomized local training procedure with global enrichment.
@@
 
-% TODO add paragraph outlining structure of the text
+The rest of the article is organized as follows.
+In~\cref{sec:method}, the building blocks of the shape optimization framework are described.
+In particular, the aforementioned two different strategies for the construction of local approximation spaces are presented.
+\Cref{sec:numerical experiments} comprises the numerical experiments.
+Based on the example of a graded concrete slab, the methods for the construction of local approximation spaces described in~\cref{sec:Randomized range finder and POD,sec:Heuristic randomized range finder} are compared, the proposed reduced order model (ROM) is validated, and results for the solution of a shape optimization problem are presented.
+Finally, conclusions are given in~\cref{sec:conclusions}.
 

### 2024-08-06 paper/paper.tex
[writing] initial commit section 2

@@
 
+\usepackage{tikz}
 \usepackage{standalone}
@@
 Second, a framework for the shape optimization of lattice structures is proposed.
-It combines the aforementioned algorithms to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final reduced order model.
+It combines the aforementioned algorithms to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final reduced order model (ROM).
 Furthermore, the global approximation space is constructed from the local spaces using the generalized finite element method (GFEM)~\cite{Babuska2004Generalized}.
@@
 \Cref{sec:numerical experiments} comprises the numerical experiments.
-Based on the example of a graded concrete slab, the methods for the construction of local approximation spaces described in~\cref{sec:Randomized range finder and POD,sec:Heuristic randomized range finder} are compared, the proposed reduced order model (ROM) is validated, and results for the solution of a shape optimization problem are presented.
+Based on the example of a graded concrete slab, the methods for the construction of local approximation spaces described in~\cref{sec:Randomized range finder and POD,sec:Heuristic randomized range finder} are compared, the proposed ROM is validated, and results for the solution of a shape optimization problem are presented.
 Finally, conclusions are given in~\cref{sec:conclusions}.
@@
 \label{sec:method}
+The method proposed in this article is a localized pMOR framework for the shape optimization of lattice structures.
+First, the auxiliary problem to model geometrical parameterizations is introduced in~\cref{sub:Auxiliary Problem}.
+Second, the construction of local approximation spaces is described in~\cref{sub:Construction of local approximation spaces}.
+Herein, the parametric oversampling problem and two different strategies to approximate the range of the corresponding parametric transfer operator are discussed.
+Third, it is outlined how a global approximation is obtained from local approximation spaces via the GFEM.
+Finally, the necessity and use of hyperreduction (in the form of empirical interpolation) to ensure an online-efficient ROM is detailed in~\cref{sub:Online efficiency}.
 
@@
 \label{sub:Auxiliary Problem}
+The approach adopted in this work to facilitate geometrical parametrizations belongs to the class of surface-based deformations~\cite{Botsch2010Polygon} and is based on transformations $\Phi_{\mu}$ that map each material point $\bm{x}_{\mathrm{p}}$ of a parameter-independent reference or parent domain $\varOmega^{\mathrm{p}}$ to a point $\bm{x}_{\mu}$ in the parameter-dependent current or physical domain $\varOmega^{\mu}$\comment[id=pd]{Add figure and reference to figure}.
+An auxiliary problem based on the equations of linear elasticity is solved to obtain such domain transformations, following the approach outlined in~\cite{Guo2022Learning}, and briefly repeated here for completeness.
+
+The transformation map $\Phi_{\mu}$ from parent to physical domain $\Phi_{\mu}: \varOmega^{\mathrm{p}}\mapsto\varOmega^{\mu}$ is given by $\bm{x}_{\mu}=\Phi_{\mu}(\bm{x}_{\mathrm{p}}) = \bm{x}_{\mathrm{p}} + \bm{d}(\bm{x}_{\mathrm{p}}; \mu)$,
+% \begin{align}
+%     \Phi_{\mu}&: \varOmega^{\mathrm{p}}\mapsto\varOmega^{\mu}\,,\\\bm{x}_{\mu}&=\Phi_{\mu}(\bm{x}_{\mathrm{p}}) = \bm{x}_{\mathrm{p}} + \bm{d}(\bm{x}_{\mathrm{p}}; \mu)\,,
+%     \label{eq:trafo}
+% \end{align}
+with $\bm{d}(\bm{x}_{\mathrm{p}}; \bm{\mu})$ being the transformation displacement field.
+\comment[id=pd]{Consistent notation for $\mu$ and its components. Suffix must also be bold then?}
+The transformation displacement is determined by solving the following linear elastostatic auxiliary problem.
+\begin{align}
+    \div{\left(\mathbb{C}\vdot\!\vdot\frac{1}{2}\left(
+            \bm{d}\otimes\grad + \grad\otimes\bm{d}
+\right)\right)}
+            = \bm{0}\,,\quad&\mathrm{in}\;\varOmega^{\mathrm{p}}\,,\\
+    \bm{d} = \bm{0}\,,\quad&\mathrm{on}\;\partial\varOmega^{\mathrm{p}}\,,\\
+    \bm{d} = \bm{x}_{\mu} - \bm{x}_{\mathrm{p}}\,,\quad&\mathrm{on}\;\partial\varOmega^{\mathrm{p}}_{\mathrm{int}}\,.
+    \label{eq:aux}
+\end{align}
+\comment[id=pd]{Nabla should be explicitly defined?}
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=0.95\textwidth]{./img/transformationmap.pdf}
+    \end{center}
+    \caption{Mapping.}\label{fig:transformationmap}
+\end{figure}
+
+% TODO move these equation to proper section (3 or 4)
+% probably section 3, because there we would need to derive the global final system as well
+% Subdomain operator on the physical domain $\varOmega^{\mu}$:
+% \begin{equation}
+%     \int_{\varOmega^{\mu}} \frac{\partial\bm{v}}{\partial\bm{x}_{\mu}}
+%     \vdot\!\vdot\,\hat{\mathbb{C}}\vdot\!\vdot\,\frac{\partial\bm{w}}{\partial\bm{x}_{\mu}}\dd{\bm{x}_{\mu}}\,.
+% \end{equation}
+% Subdomain operator on the parent domain $\varOmega^{\mathrm{p}}$:
+% \begin{equation}
+%     \int_{\varOmega^{\mathrm{p}}} \frac{\partial\bm{v}}{\partial\bm{x}_{\mathrm{p}}}
+%     \vdot\!\vdot\,\left(\hat{\mathbb{C}}\vdot\!\vdot\,
+%     \left(\frac{\partial\bm{w}}{\partial\bm{x}_{\mathrm{p}}}\vdot\bm{F}_{\mu}^{-1}\right)\right)
+%     \vdot\bm{F}_{\mu}^{-\mathrm{T}}
+%     \det(\bm{F}_{\mu})
+%     \dd{\bm{x}_{\mathrm{p}}}\,,
+%     \label{eq:pullback}
+% \end{equation}
+% with
+% \begin{equation}
+%     \bm{F}_{\mu}:= \frac{\partial\bm{x}_{\mu}}{\partial\bm{x}_{\mathrm{p}}}\,,\quad
+%     \mathrm{and}\quad
+%     \dd{\bm{x}_{\mu}}=\det(\bm{F}_{\mu})\dd{\bm{x}_{\mathrm{p}}}\,.
+%     \label{eq:defgrad}
+% \end{equation}
 

### 2024-08-07 paper/paper.tex
[writing] finalize auxiliary problem

@@
 \pgfplotsset{compat=1.18}
+\usepackage{bamcolors}
 
@@
 \label{sub:Auxiliary Problem}
-The approach adopted in this work to facilitate geometrical parametrizations belongs to the class of surface-based deformations~\cite{Botsch2010Polygon} and is based on transformations $\Phi_{\mu}$ that map each material point $\bm{x}_{\mathrm{p}}$ of a parameter-independent reference or parent domain $\varOmega^{\mathrm{p}}$ to a point $\bm{x}_{\mu}$ in the parameter-dependent current or physical domain $\varOmega^{\mu}$\comment[id=pd]{Add figure and reference to figure}.
+The approach adopted in this work to facilitate geometrical parametrizations belongs to the class of surface-based deformations~\cite{Botsch2010Polygon} and is based on transformations $\Phi_{\bm\mu}$ that map each material point $\bm{x}_{\mathrm{p}}$ of a parameter-independent reference or parent domain $\varOmega^{\mathrm{p}}$ to a point $\bm{x}_{\bm\mu}$ in the parameter-dependent current or physical domain $\varOmega^{\bm\mu}$.
+In the context of shape optimization of lattice structures, our objective is to determine the mapping $\Phi_{\bm\mu}$ for a single unit cell, see~\cref{fig:transformationmap}, and use it to describe the change in geometry of each unit cell throughout the structure.
 An auxiliary problem based on the equations of linear elasticity is solved to obtain such domain transformations, following the approach outlined in~\cite{Guo2022Learning}, and briefly repeated here for completeness.
 
-The transformation map $\Phi_{\mu}$ from parent to physical domain $\Phi_{\mu}: \varOmega^{\mathrm{p}}\mapsto\varOmega^{\mu}$ is given by $\bm{x}_{\mu}=\Phi_{\mu}(\bm{x}_{\mathrm{p}}) = \bm{x}_{\mathrm{p}} + \bm{d}(\bm{x}_{\mathrm{p}}; \mu)$,
-% \begin{align}
-%     \Phi_{\mu}&: \varOmega^{\mathrm{p}}\mapsto\varOmega^{\mu}\,,\\\bm{x}_{\mu}&=\Phi_{\mu}(\bm{x}_{\mathrm{p}}) = \bm{x}_{\mathrm{p}} + \bm{d}(\bm{x}_{\mathrm{p}}; \mu)\,,
-%     \label{eq:trafo}
-% \end{align}
+The transformation map $\Phi_{\bm\mu}$ from parent to physical domain $\Phi_{\bm\mu}: \varOmega^{\mathrm{p}}\mapsto\varOmega^{\bm\mu}$ is given by $\bm{x}_{\bm\mu}=\Phi_{\bm\mu}(\bm{x}_{\mathrm{p}}) = \bm{x}_{\mathrm{p}} + \bm{d}(\bm{x}_{\mathrm{p}}; \bm\mu)$,
 with $\bm{d}(\bm{x}_{\mathrm{p}}; \bm{\mu})$ being the transformation displacement field.
-\comment[id=pd]{Consistent notation for $\mu$ and its components. Suffix must also be bold then?}
 The transformation displacement is determined by solving the following linear elastostatic auxiliary problem.
 \begin{align}
-    \div{\left(\mathbb{C}\vdot\!\vdot\frac{1}{2}\left(
+    \label{eq:aux_1}\div{\left(\hat{\mathbb{C}}\vdot\!\vdot\frac{1}{2}\left(
             \bm{d}\otimes\grad + \grad\otimes\bm{d}
@@
             = \bm{0}\,,\quad&\mathrm{in}\;\varOmega^{\mathrm{p}}\,,\\
-    \bm{d} = \bm{0}\,,\quad&\mathrm{on}\;\partial\varOmega^{\mathrm{p}}\,,\\
-    \bm{d} = \bm{x}_{\mu} - \bm{x}_{\mathrm{p}}\,,\quad&\mathrm{on}\;\partial\varOmega^{\mathrm{p}}_{\mathrm{int}}\,.
-    \label{eq:aux}
+    \label{eq:aux_2}\bm{d} = \bm{0}\,,\quad&\mathrm{on}\;\partial\varOmega^{\mathrm{p}}\,,\\
+    \label{eq:aux_3}\bm{d} = \bm{x}_{\mu} - \bm{x}_{\mathrm{p}}\,,\quad&\mathrm{on}\;\partial\varOmega^{\mathrm{p}}_{\mathrm{int}}\,.
 \end{align}
-\comment[id=pd]{Nabla should be explicitly defined?}
+The stiffness tetrad of the auxiliary problem is defined as
+\begin{equation}
+    \hat{\mathbb{C}} = \hat{\lambda} \bm{I}\otimes\bm{I} + 2\hat{\mu}\mathbb{I}\,,\quad\mathrm{with}\quad \hat{\lambda}=\frac{\nu}{(1+\nu)(1-2\nu)}\quad\mathrm{and}\quad\hat{\mu}=\frac{1}{2(1+\nu)}\,.
+    \label{eq:aux_tetrad}
+\end{equation}
+With $\bm{x}_{\bm\mu}$ known for all points on the parent interface $\partial\varOmega^{\mathrm{p}}_{\mathrm{int}}$, the desired transformation (\eg{} enlarging/shrinking of the voids radius) is enforced by prescribing~\cref{eq:aux_2,eq:aux_3}.
+
+Given the transformation displacement $\bm{d}(\bm{x}_{\mathrm{p}};\bm\mu)$, the variational formulation is formulated over the parameter-independent parent domain $\varOmega^{\mathrm{p}}$ instead of the parameter-dependent physical domain $\varOmega^{\bm\mu}$.
+This introduces the parameter dependence in the variational formulation as shown in the following subsections, but has the advantage that all computations (integrations) can be carried out over the fixed parent domain and thus no re-meshing is required.
+However, for each new parameter value $\bm\mu$ for which the model is to be evaluated, we first need to solve the auxiliary problem.
+Nevertheless, this does not pose a problem, since the solution of the auxiliary problem is well amenable to acceleration via established MOR techniques and it is referred to~\cite{Guo2022Learning} for further details.
 
 \begin{figure}
-    \begin{center}
-        \includegraphics[width=0.95\textwidth]{./img/transformationmap.pdf}
-    \end{center}
-    \caption{Mapping.}\label{fig:transformationmap}
+    \centering
+    \include{./img/transformationmap.tex}
+    \caption{Transformation map $\Phi_{\bm\mu}$ from parent to physical domain on the example of the unit square domain with a circular void placed in the center.
+The parameter $\bm\mu$ controls the radius of the void.}\label{fig:transformationmap}
 \end{figure}

### 2024-08-07 paper/paper.tex
[writing] add algorithms

@@
 \usepackage{bamcolors}
+\usepackage{algorithm, algorithmicx, algpseudocode}
 
@@
 
+% - setting, global domain = union of non overlapping subdomains = unit cells
+% - introduction of coarse grid where each cell corresponds to unit cell subdomain
+% - goal of constructing local spaces for only parts of the domain
+% - definition of oversampling problem and transfer operator
+% - the oversampling problem should be formulated for union of parent unit cells
+% - weak formulation is not required here? Should be first introduced in section 2.3?
+% - comment on number of archetypes/configurations
+
 % subsection Construction of local approximation spaces (end)
@@
 
+\begin{algorithm}
+    \caption{RRFPOD caption.}%
+    \label{algo:rrfpod}
+    \begin{algorithmic}[1]
+        \Function{HapodRangeApproximation}{$\bm{T}$, $\texttt{tol}$, $n_t$, $\varepsilon_{\mathrm{algofail}}$, $N_{\mathrm{train}}$}
+    \Statex{\textbf{Input:} Operator $\bm{T}_{\bm\mu}$, target tolerance $\texttt{tol}$, number of testvectors $n_t$, maximum failure probability $\varepsilon_{\mathrm{algofail}}$, number of parameter samples $N_{\mathrm{train}}$}
+    \Statex{\textbf{Output:} space $X^n$}
+
+    \State{$\bm{S}$ $\gets\;\emptyset$}\Comment{initialize snapshot set}
+    \State{$\bm{S}_{\mathrm{train}}\gets\;\{\bm\mu_1,\ldots,\bm\mu_{N_\mathrm{train}}\}$}\Comment{initialize training set}
+    \For{$\bm\mu_j$ in $\bm{S}_{\mathrm{train}}$}
+    \State{$\bm{R}\gets$ \textsc{AdaptiveRandomizedRangeApproximation}($\bm{T}_{\bm\mu_j}$, $\texttt{tol}$, $n_t$, $\varepsilon_{\mathrm{algofail}}$)}
+    \State{$\bm{u}_{\mathrm{Neumann}}(\bm\mu_j)\gets \bm{T}_{\bm\mu_j}(\bm{0})$}\Comment{solve additional Neumann problem if $\varGamma_{\mathrm{N}}\ne\emptyset$}
+    \State{$\bm{R}\gets \bm{R}\cup\left(\bm{u}_{\mathrm{Neumann}}(\bm\mu_j)\right)$}
+    \State{$\bm{S}\gets \bm{S}\cup\left(\mathrm{orthonormalize}\left(\bm{R}\right)\right)$}
+    \EndFor%
+    \State{$\bm{B}\gets\textsc{POD}(\bm{S})$}
+    \State{\textbf{return} {$X^n=\mathrm{span}\left(\bm{B}\right)$}}
+    \EndFunction%
+    \end{algorithmic} 
+\end{algorithm}
+
+test
+\comment[id=pd]{TODO: check how/when Neumann data is added in case of RRFPOD. This can also be done outside the loop and even after POD of spectral basis.}
+
 % subsubsection Randomized range finder and POD (end)
@@
 \label{sec:Heuristic randomized range finder}
+
+\begin{algorithm}
+    \caption{HRRF caption.}%
+    \label{algo:hrrf}
+    \begin{algorithmic}[1]
+        \Function{HeuristicRangeApproximation}{$\bm{T}$, $\texttt{tol}$, $n_t$, $\varepsilon_{\mathrm{algofail}}$, $N_{\mathrm{train}}$, $N_{\mathrm{test}}$}
+    \Statex{\textbf{Input:} Operator $\bm{T}_{\bm\mu}$, target tolerance $\texttt{tol}$, number of testvectors $n_t$, maximum failure probability $\varepsilon_{\mathrm{algofail}}$, number of parameter samples $N_{\mathrm{train}}$, number of parameter samples $N_{\mathrm{test}}$}
+    \Statex{\textbf{Output:} space $X^n$}
+
+    \State{$\bm{B}$ $\gets\;\emptyset$}\Comment{initialize basis}
+    \State{$\bm{S}_{\mathrm{train}}\gets\;\{\bm\mu_1,\ldots,\bm\mu_{N_\mathrm{train}}\}$}\Comment{initialize training set}
+    \State{$\bm{S}_{\mathrm{test}}\gets\;\{\bm\mu_1,\ldots,\bm\mu_{N_\mathrm{test}}\}$}\Comment{initialize testing set}
+    \State{$\bm{M}\gets\{\bm{T}_{\bm\mu_1}(\bm{g}^1),\ldots,\bm{T}_{\bm\mu_1}(\bm{g}^{n_t}),\ldots,\bm{T}_{\bm\mu_{N_{\mathrm{test}}}}(\bm{g}^1),\ldots,\bm{T}_{\bm\mu_{N_{\mathrm{test}}}}(\bm{g}^{n_t})\}$}\Comment{initialize $N_{\mathrm{test}}\cdot\,n_t$ test vectors}
+    \State{$\varepsilon_{\mathrm{testfail}}\gets\varepsilon_{\mathrm{algofail}}/N_{\bm{T}}$}
+    \State{$c_{\mathrm{est}}\gets{\left[\sqrt{2\lambda^{M_S}_{\min}}\mathrm{erf}^{-1}\left(\sqrt[n_t]{\varepsilon_{\mathrm{testfail}}}\right)\right]}^{-1}$}\Comment{determine error estimator factor}
+    \While{$\left(\max_{t\in \bm{M}} \norm{t}_R\right) \vdot c_{\mathrm{est}} > \texttt{tol}$}
+        \State{$\bm\mu_j\gets\mathrm{random.choice}(\bm{S}_{\mathrm{train}})$}
+        \State{$\bm{B}\gets \bm{B}\cup\left(\bm{T}_{\bm\mu_j}(\bm{g})\right)$}
+        \State{$\bm{B}\gets \mathrm{orthonormalize}\left(\bm{B}\right)$}
+        \State{$\bm{M}\gets\{t-P_{\mathrm{span}(\bm{B})}t \mid t\in \bm{M}\}$}\Comment{orthogonalize test vectors to $\mathrm{span}(\bm{B})$}
+    \EndWhile{}
+    \State{\textbf{return} {$X^n=\mathrm{span}\left(\bm{B}\right)$}}
+    \EndFunction%
+    \end{algorithmic} 
+\end{algorithm}
 

### 2024-08-19 paper/paper.tex
[sec 2.2] initial commit

@@
 The contributions of the present work are given as follows.
-First, as references~\cite{Smetana2016Optimal,Taddei2018Localization} do not make use of range approximation via random sampling, suitable training strategies to construct local approximation spaces for parameterized linear problems via random sampling are investigated.
-In particular, the approach given in~\cite{Taddei2018Localization}, identified as a \textit{distributed approximate POD} (see also~\cite{Himpe2018Hierarchical}), is compared to a \textit{heuristic randomized range finder} developed in this work.
+First, as references~\cite{Smetana2016Optimal,Taddei2018Localization} do not make use of range approximation via random sampling, a suitable training strategy to construct local approximation spaces for parameterized linear problems via random sampling is discussed.
+Herein, the approach given in~\cite{Taddei2018Localization}, identified as a \textit{distributed approximate POD} (see also~\cite{Himpe2018Hierarchical}), is adopted to range approximation via random sampling.
 Second, a framework for the shape optimization of lattice structures is proposed.
-It combines the aforementioned algorithms to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final reduced order model (ROM).
+It combines the aforementioned algorithm to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final reduced order model (ROM).
 Furthermore, the global approximation space is constructed from the local spaces using the generalized finite element method (GFEM)~\cite{Babuska2004Generalized}.
@@
 In~\cref{sec:method}, the building blocks of the shape optimization framework are described.
-In particular, the aforementioned two different strategies for the construction of local approximation spaces are presented.
+In particular, the range approximation of a parametric transfer operator is discussed.
 \Cref{sec:numerical experiments} comprises the numerical experiments.
-Based on the example of a graded concrete slab, the methods for the construction of local approximation spaces described in~\cref{sec:Randomized range finder and POD,sec:Heuristic randomized range finder} are compared, the proposed ROM is validated, and results for the solution of a shape optimization problem are presented.
+Based on the example of a graded concrete slab, the quality of the local spaces generated by method described in~\cref{sec:Randomized range finder and POD} is analyzed, the proposed ROM is validated, and results for the solution of a shape optimization problem are presented.
 Finally, conclusions are given in~\cref{sec:conclusions}.
@@
 Second, the construction of local approximation spaces is described in~\cref{sub:Construction of local approximation spaces}.
-Herein, the parametric oversampling problem and two different strategies to approximate the range of the corresponding parametric transfer operator are discussed.
+Herein, the parametric oversampling problem and training strategy to approximate the range of the corresponding parametric transfer operator are discussed.
 Third, it is outlined how a global approximation is obtained from local approximation spaces via the GFEM.
-Finally, the necessity and use of hyperreduction (in the form of empirical interpolation) to ensure an online-efficient ROM is detailed in~\cref{sub:Online efficiency}.
+Finally, the necessity and use of hyper-reduction (in the form of empirical interpolation) to ensure an online-efficient ROM is detailed in~\cref{sub:Online efficiency}.
 
@@
 \label{sub:Auxiliary Problem}
-The approach adopted in this work to facilitate geometrical parametrizations belongs to the class of surface-based deformations~\cite{Botsch2010Polygon} and is based on transformations $\Phi_{\bm\mu}$ that map each material point $\bm{x}_{\mathrm{p}}$ of a parameter-independent reference or parent domain $\varOmega^{\mathrm{p}}$ to a point $\bm{x}_{\bm\mu}$ in the parameter-dependent current or physical domain $\varOmega^{\bm\mu}$.
+The approach adopted in this work to facilitate geometrical parameterizations belongs to the class of surface-based deformations~\cite{Botsch2010Polygon} and is based on transformations $\Phi_{\bm\mu}$ that map each material point $\bm{x}_{\mathrm{p}}$ of a parameter-independent reference or parent domain $\varOmega^{\mathrm{p}}$ to a point $\bm{x}_{\bm\mu}$ in the parameter-dependent current or physical domain $\varOmega^{\bm\mu}$.
 In the context of shape optimization of lattice structures, our objective is to determine the mapping $\Phi_{\bm\mu}$ for a single unit cell, see~\cref{fig:transformationmap}, and use it to describe the change in geometry of each unit cell throughout the structure.
@@
 Given the transformation displacement $\bm{d}(\bm{x}_{\mathrm{p}};\bm\mu)$, the variational formulation is formulated over the parameter-independent parent domain $\varOmega^{\mathrm{p}}$ instead of the parameter-dependent physical domain $\varOmega^{\bm\mu}$.
-This introduces the parameter dependence in the variational formulation as shown in the following subsections, but has the advantage that all computations (integrations) can be carried out over the fixed parent domain and thus no re-meshing is required.
+This introduces the parameter dependence in the variational formulation as shown in the following subsections, but has the advantage that numerical integration can be carried out over the fixed parent domain and thus no re-meshing is required.
 However, for each new parameter value $\bm\mu$ for which the model is to be evaluated, we first need to solve the auxiliary problem.
@@
 \label{sub:Construction of local approximation spaces}
+In this paper, local approximation spaces are constructed by solving an oversampling problem many times for different parameter values $\bm\mu$ and different (random) boundary conditions.
+It is therefore useful to cast this oversampling problem in the form of a parameter-dependent transfer operator $\bm{T}_{\bm\mu}$ that
+maps the boundary function $\bm{g}$ to the solution $\bm{u}(\bm\mu)$ in the target subdomain $\varOmega_{\mathrm{in}}$.
+The approximation of the range of this transfer operator is then the sought after local approximation space.
 
+\added[id=pd]{TODO: need to introduce pull back here already because this is used in the training as well}
+% First, use physical domain (suffix mu everywhere)
+% Then introduce pull back and write omega_p everywhere ?
+
+First, the global domain $\varOmega_{\mathrm{gl}}^{\bm\mu}$ and a non-overlapping domain decomposition
+\begin{equation}
+	\varOmega_{\mathrm{gl}}^{\bm\mu}=\cup_{i=1}^{N_{\mathrm{cells}}} \varOmega_i^{\bm\mu}
+\end{equation}
+is introduced.
+In the context of lattice structures, each subdomain $\varOmega_i^{\bm\mu}$ corresponds to a unit cell.
+Next, a coarse grid partition of the global domain and a fine grid partition of the unit cell is introduced as shown in~\added[id=pd]{TODO: add figure showing coarse grid and fine grid for a single unit cell.}
+\added[id=pd]{Fig. XX} also shows one exemplary oversampling domain $\varOmega^{\bm\mu}$.
+The oversampling problem then comprises the solution of the following boundary value problem.
+\begin{align}
+	\begin{split}
+	\label{eq:osp_p}
+	u &= 0\,,\\
+	v &= 1
+	\end{split}
+\end{align}
+
+\added[id=pd]{TODO: variational formulation oversampling problem. Maybe skip BVP in order to reduce number of eqs.}
 % - setting, global domain = union of non overlapping subdomains = unit cells
@@
 % - comment on number of archetypes/configurations
+
+\added[id=pd]{TODO: add footnote about extension of Auxiliary Problem from unit cell to global domain / oversampling domain.}
+
+% oversampling problem pulled back to parent domain
+\begin{align}
+	\begin{split}
+		\label{eq:osp_mu}
+		u &= 0 \,,\\
+		v &= 1
+	\end{split}
+\end{align}
 
@@
 % subsubsection Randomized range finder and POD (end)
-
-\subsubsection{Heuristic randomized range finder} % (fold)
-\label{sec:Heuristic randomized range finder}
-
-\begin{algorithm}
-    \caption{HRRF caption.}%
-    \label{algo:hrrf}
-    \begin{algorithmic}[1]
-        \Function{HeuristicRangeApproximation}{$\bm{T}$, $\texttt{tol}$, $n_t$, $\varepsilon_{\mathrm{algofail}}$, $N_{\mathrm{train}}$, $N_{\mathrm{test}}$}
-    \Statex{\textbf{Input:} Operator $\bm{T}_{\bm\mu}$, target tolerance $\texttt{tol}$, number of testvectors $n_t$, maximum failure probability $\varepsilon_{\mathrm{algofail}}$, number of parameter samples $N_{\mathrm{train}}$, number of parameter samples $N_{\mathrm{test}}$}
-    \Statex{\textbf{Output:} space $X^n$}
-
-    \State{$\bm{B}$ $\gets\;\emptyset$}\Comment{initialize basis}
-    \State{$\bm{S}_{\mathrm{train}}\gets\;\{\bm\mu_1,\ldots,\bm\mu_{N_\mathrm{train}}\}$}\Comment{initialize training set}
-    \State{$\bm{S}_{\mathrm{test}}\gets\;\{\bm\mu_1,\ldots,\bm\mu_{N_\mathrm{test}}\}$}\Comment{initialize testing set}
-    \State{$\bm{M}\gets\{\bm{T}_{\bm\mu_1}(\bm{g}^1),\ldots,\bm{T}_{\bm\mu_1}(\bm{g}^{n_t}),\ldots,\bm{T}_{\bm\mu_{N_{\mathrm{test}}}}(\bm{g}^1),\ldots,\bm{T}_{\bm\mu_{N_{\mathrm{test}}}}(\bm{g}^{n_t})\}$}\Comment{initialize $N_{\mathrm{test}}\cdot\,n_t$ test vectors}
-    \State{$\varepsilon_{\mathrm{testfail}}\gets\varepsilon_{\mathrm{algofail}}/N_{\bm{T}}$}
-    \State{$c_{\mathrm{est}}\gets{\left[\sqrt{2\lambda^{M_S}_{\min}}\mathrm{erf}^{-1}\left(\sqrt[n_t]{\varepsilon_{\mathrm{testfail}}}\right)\right]}^{-1}$}\Comment{determine error estimator factor}
-    \While{$\left(\max_{t\in \bm{M}} \norm{t}_R\right) \vdot c_{\mathrm{est}} > \texttt{tol}$}
-        \State{$\bm\mu_j\gets\mathrm{random.choice}(\bm{S}_{\mathrm{train}})$}
-        \State{$\bm{B}\gets \bm{B}\cup\left(\bm{T}_{\bm\mu_j}(\bm{g})\right)$}
-        \State{$\bm{B}\gets \mathrm{orthonormalize}\left(\bm{B}\right)$}
-        \State{$\bm{M}\gets\{t-P_{\mathrm{span}(\bm{B})}t \mid t\in \bm{M}\}$}\Comment{orthogonalize test vectors to $\mathrm{span}(\bm{B})$}
-    \EndWhile{}
-    \State{\textbf{return} {$X^n=\mathrm{span}\left(\bm{B}\right)$}}
-    \EndFunction%
-    \end{algorithmic} 
-\end{algorithm}
-
-% subsubsection Heuristic randomized range finder (end)
 

### 2024-08-20 paper/paper.tex
wip sec 2.2

@@
 \usepackage{bm}
+\usepackage{upgreek}
 \usepackage{physics}
@@
 
-% custom commands
+% %%% Custom commands
 \newcommand{\ie}{i.\,e.\@\xspace}
 \newcommand{\eg}{e.\,g.\@\xspace}
+% suffixes
+\newcommand{\m}{\bm\mu}
+\newcommand{\gl}{\mathrm{gl}}
+\newcommand{\p}{\mathrm{p}}
+\newcommand{\out}{\mathrm{out}}
+\newcommand{\inrm}{\mathrm{in}}
 
@@
 Finally, conclusions are given in~\cref{sec:conclusions}.
+
+\added[id=pd]{TODO: add note on notation where appropriate.}
 
@@
 maps the boundary function $\bm{g}$ to the solution $\bm{u}(\bm\mu)$ in the target subdomain $\varOmega_{\mathrm{in}}$.
-The approximation of the range of this transfer operator is then the sought after local approximation space.
+The approximation of the range of this transfer operator is then the sought after local approximation space, see~\cite{Buhr2018Randomized}.
 
@@
 \added[id=pd]{Fig. XX} also shows one exemplary oversampling domain $\varOmega^{\bm\mu}$.
-The oversampling problem then comprises the solution of the following boundary value problem.
+The oversampling problem (given for linear elastostatics) then comprises the solution of the following boundary value problem.
 \begin{align}
@@
 	\label{eq:osp_p}
-	u &= 0\,,\\
-	v &= 1
+    - \div{\bm{\sigma}} (\bm{u}(\bm\mu)) &= \bm{0} \quad  \mathrm{in}\;{\varOmega^{\bm\mu}}\subset\varOmega^{\bm\mu}_{\mathrm{gl}}\,,\\
+    \bm{\sigma}(\bm{u}(\bm\mu)) \vdot \bm{n} &= \bm{0} \quad \mathrm{on}\; \varGamma^{\m}_{\mathrm{N}}:= \partial\varOmega^{\m}\cap\varSigma^{\m}_{\mathrm{N}}\,,\\
+    \bm{u}(\bm\mu) &= \bm{0} \quad  \mathrm{on}\;\varGamma^{\m}_{\mathrm{D}}:=\partial\varOmega^{\m}\cap\varSigma^{\m}_{\mathrm{D}}\,,\\
+    \bm{u}(\bm\mu) &= \bm{g}\quad\mathrm{on}\;\varGamma^{\m}_{\mathrm{out}}:= \partial\varOmega^{\m} \setminus\partial\varOmega^{\m}_{\mathrm{gl}}\,.
 	\end{split}
 \end{align}
+Here, $\bm\sigma$ is the \textsc{Cauchy} stress, $\bm{n}$ the normal vector and $\bm{g}$ \textsc{Dirichlet} boundary data to be prescribed on the boundary $\varGamma^{\m}_{\out}$.
+The boundaries $\varGamma^{\m}_{\mathrm{N}}$ and $\varGamma^{\m}_{\mathrm{D}}$ denote the part of the boundary of the oversampling domain that coincides with the global \textsc{Neumann} boundary $\varSigma^{\m}_{\mathrm{N}}$ or \textsc{Dirichlet} boundary $\varSigma^{\m}_{\mathrm{D}}$, respectively.
+Note that the topology is dependent on the target subdomain $\varOmega^{\m}_{\inrm}$ and the size of the oversampled region (\added[id=pd]{TODO: refer to figure as above}).
 
 \added[id=pd]{TODO: variational formulation oversampling problem. Maybe skip BVP in order to reduce number of eqs.}
+Introducing the linear strain tensor as
+\begin{equation}
+    \label{eq:eps}
+    \bm\varepsilon(\bm{w}):= \frac{1}{2}\left(\bm{w}\otimes\nabla + \nabla\otimes\bm{w}\right)\,,
+\end{equation}
+the weak form reads \added[id=pd]{TODO: need to write out derivatives here, because of the pull back later. Also need to write everything with stiffness tetrad.}
+\begin{equation}
+    \int_{\varOmega^{\m}} \bm\varepsilon(\updelta\bm{u})\vdot\!\vdot\bm\sigma(\bm{u}_0(\m)) \dd{\bm{x}_{\m}} = -\int_{\varOmega^{\m}} \bm\varepsilon(\updelta\bm{u})\vdot\!\vdot\bm\sigma(\hat{\bm{g}})\dd{\bm{x}_{\m}}\,.
+    \label{eq:weak_osp}
+\end{equation}
+Herein, $\updelta\bm{u}$ denotes the test function and $\hat{\bm{g}}$ denotes a suitable lifting function for the inhomogeneous \textsc{Dirichlet} boundary condition and the solution is sought as $\bm{u}(\m)=\bm{u}_0(\m)+\hat{\bm{g}}$.
+Finally, the solution is restricted to the target subdomain which is denoted as $\bm{u}(\m)\big\vert_{\varOmega^{\m}_{\inrm}}$.
+The weak form is implicitly dependent on the parameter $\m$ due to the integration carried out over the physical domain $\varOmega^{\m}$.
+By introducing the deformation gradient of the geometrical transformation
+\begin{equation}
+    \bm{F}_{\m}:= \dv{\bm{x}^{\m}}{\bm{x}^{\p}}\,,
+    \label{eq:F_mu}
+\end{equation}
+the integration can be carried out over the fixed parent domain
+\begin{equation}
+    \int_{\varOmega^{\p}} \bm\varepsilon(\updelta\bm{u})\vdot\!\vdot\bm\sigma(\bm{u}_0(\m)) \dd{\bm{x}_{\m}} = -\int_{\varOmega^{\p}} \bm\varepsilon(\updelta\bm{u})\vdot\!\vdot\bm\sigma(\hat{\bm{g}})\dd{\bm{x}_{\m}}\,.
+    \label{eq:weak_osp_parent}
+\end{equation}
 % - setting, global domain = union of non overlapping subdomains = unit cells
@@
 % - definition of oversampling problem and transfer operator
+
+% these are details of the specific problem we solve later
 % - the oversampling problem should be formulated for union of parent unit cells

### 2024-08-22 paper/paper.tex
add progress with paper

@@
 \usepackage{cleveref}
+\usepackage{algorithm, algorithmicx, algpseudocode}
 
@@
 \pgfplotsset{compat=1.18}
+\usetikzlibrary{%
+	calc,
+	patterns,
+	positioning,
+	decorations.pathreplacing,%
+	decorations.pathmorphing,%
+	arrows.meta,%
+}
 \usepackage{bamcolors}
-\usepackage{algorithm, algorithmicx, algpseudocode}
 
@@
 \newcommand{\inrm}{\mathrm{in}}
+
+\graphicspath{{./img/}}
 
@@
 \Cref{sec:numerical experiments} comprises the numerical experiments.
-Based on the example of a graded concrete slab, the quality of the local spaces generated by method described in~\cref{sec:Randomized range finder and POD} is analyzed, the proposed ROM is validated, and results for the solution of a shape optimization problem are presented.
+Based on the example of a graded concrete slab, the quality of the local spaces generated by the method described in~\cref{sec:Randomized range finder and POD} is analyzed, the proposed ROM is validated, and results for the solution of a shape optimization problem are presented.
 Finally, conclusions are given in~\cref{sec:conclusions}.
@@
 In the context of shape optimization of lattice structures, our objective is to determine the mapping $\Phi_{\bm\mu}$ for a single unit cell, see~\cref{fig:transformationmap}, and use it to describe the change in geometry of each unit cell throughout the structure.
+% The auxiliary problem is formulated here for a single unit cell, and for ease of notation the subscript $i$ as employed in the remainder of the article is dropped.
 An auxiliary problem based on the equations of linear elasticity is solved to obtain such domain transformations, following the approach outlined in~\cite{Guo2022Learning}, and briefly repeated here for completeness.
@@
 
-% TODO move these equation to proper section (3 or 4)
-% probably section 3, because there we would need to derive the global final system as well
-% Subdomain operator on the physical domain $\varOmega^{\mu}$:
-% \begin{equation}
-%     \int_{\varOmega^{\mu}} \frac{\partial\bm{v}}{\partial\bm{x}_{\mu}}
-%     \vdot\!\vdot\,\hat{\mathbb{C}}\vdot\!\vdot\,\frac{\partial\bm{w}}{\partial\bm{x}_{\mu}}\dd{\bm{x}_{\mu}}\,.
-% \end{equation}
-% Subdomain operator on the parent domain $\varOmega^{\mathrm{p}}$:
-% \begin{equation}
-%     \int_{\varOmega^{\mathrm{p}}} \frac{\partial\bm{v}}{\partial\bm{x}_{\mathrm{p}}}
-%     \vdot\!\vdot\,\left(\hat{\mathbb{C}}\vdot\!\vdot\,
-%     \left(\frac{\partial\bm{w}}{\partial\bm{x}_{\mathrm{p}}}\vdot\bm{F}_{\mu}^{-1}\right)\right)
-%     \vdot\bm{F}_{\mu}^{-\mathrm{T}}
-%     \det(\bm{F}_{\mu})
-%     \dd{\bm{x}_{\mathrm{p}}}\,,
-%     \label{eq:pullback}
-% \end{equation}
-% with
-% \begin{equation}
-%     \bm{F}_{\mu}:= \frac{\partial\bm{x}_{\mu}}{\partial\bm{x}_{\mathrm{p}}}\,,\quad
-%     \mathrm{and}\quad
-%     \dd{\bm{x}_{\mu}}=\det(\bm{F}_{\mu})\dd{\bm{x}_{\mathrm{p}}}\,.
-%     \label{eq:defgrad}
-% \end{equation}
-
 % subsection Auxiliary Problem (end)
@@
 It is therefore useful to cast this oversampling problem in the form of a parameter-dependent transfer operator $\bm{T}_{\bm\mu}$ that
-maps the boundary function $\bm{g}$ to the solution $\bm{u}(\bm\mu)$ in the target subdomain $\varOmega_{\mathrm{in}}$.
+maps the boundary function $\bm{g}$ to the solution $\bm{u}(\bm\mu)$ in the target subdomain $\varOmega^{\m}_{\mathrm{in}}$.
 The approximation of the range of this transfer operator is then the sought after local approximation space, see~\cite{Buhr2018Randomized}.
-
-\added[id=pd]{TODO: need to introduce pull back here already because this is used in the training as well}
-% First, use physical domain (suffix mu everywhere)
-% Then introduce pull back and write omega_p everywhere ?
 
@@
 Note that the topology is dependent on the target subdomain $\varOmega^{\m}_{\inrm}$ and the size of the oversampled region (\added[id=pd]{TODO: refer to figure as above}).
-
-\added[id=pd]{TODO: variational formulation oversampling problem. Maybe skip BVP in order to reduce number of eqs.}
-Introducing the linear strain tensor as
+Considering linear elastic isotropic materials
+\begin{equation}
+    \bm\sigma(\bm{u}(\m)) = \mathbb{C}\vdot\!\vdot\bm\varepsilon(\bm{u}(\m))\,,
+    \label{eq:hooke}
+\end{equation}
+with the linear strain tensor as
 \begin{equation}
     \label{eq:eps}
-    \bm\varepsilon(\bm{w}):= \frac{1}{2}\left(\bm{w}\otimes\nabla + \nabla\otimes\bm{w}\right)\,,
+    \bm\varepsilon(\bm{w}):= \frac{1}{2}\left(\bm{w}\otimes\grad + \grad\otimes\bm{w}\right)\,,
 \end{equation}
-the weak form reads \added[id=pd]{TODO: need to write out derivatives here, because of the pull back later. Also need to write everything with stiffness tetrad.}
+and the stiffness tetrad given in component form as
 \begin{equation}
-    \int_{\varOmega^{\m}} \bm\varepsilon(\updelta\bm{u})\vdot\!\vdot\bm\sigma(\bm{u}_0(\m)) \dd{\bm{x}_{\m}} = -\int_{\varOmega^{\m}} \bm\varepsilon(\updelta\bm{u})\vdot\!\vdot\bm\sigma(\hat{\bm{g}})\dd{\bm{x}_{\m}}\,.
+    C_{ijkl} = \lambda \delta_{ij}\delta_{kl} + \mu \left(\delta_{ik}\delta_{jl}+\delta_{il}\delta_{jk}\right)\,,
+    \label{eq:stiffness_tetrad}
+\end{equation}
+the weak form reads \added[id=pd]{TODO: is it necessary to write Dirichlet lift explicitly?}
+\begin{equation}
+    \int_{\varOmega^{\m}} \pdv{\updelta\bm{u}}{\bm{x}^{\m}} \vdot\!\vdot \mathbb{C} \vdot\!\vdot
+    \pdv{\bm{u}(\m)}{\bm{x}^{\m}} \dd{\bm{x}^{\m}} = \bm{0}\,.
+    % -\int_{\varOmega^{\m}} \pdv{\updelta\bm{u}}{\bm{x}^{\m}} \vdot\!\vdot \mathbb{C} \vdot\!\vdot
+    % \pdv{\hat{\bm{g}}}{\bm{x}^{\m}} \dd{\bm{x}^{\m}}\,.
     \label{eq:weak_osp}
 \end{equation}
-Herein, $\updelta\bm{u}$ denotes the test function and $\hat{\bm{g}}$ denotes a suitable lifting function for the inhomogeneous \textsc{Dirichlet} boundary condition and the solution is sought as $\bm{u}(\m)=\bm{u}_0(\m)+\hat{\bm{g}}$.
-Finally, the solution is restricted to the target subdomain which is denoted as $\bm{u}(\m)\big\vert_{\varOmega^{\m}_{\inrm}}$.
+Herein, $\updelta\bm{u}$ denotes the test function.
+% Herein, $\updelta\bm{u}$ denotes the test function and $\hat{\bm{g}}$ denotes a suitable lifting function for the inhomogeneous \textsc{Dirichlet} boundary condition and the solution is sought as $\bm{u}(\m)=\bm{u}_0(\m)+\hat{\bm{g}}$.
 The weak form is implicitly dependent on the parameter $\m$ due to the integration carried out over the physical domain $\varOmega^{\m}$.
-By introducing the deformation gradient of the geometrical transformation
+By introducing the deformation gradient of the geometrical transformation\footnote{We note here that the auxiliary problem described in~\cref{sub:Auxiliary Problem} needs to be extended to the oversampling domain, which consists of a union of unit cells.
+This is straightforward in the current setup, but has the limitation that \cref{eq:aux_2} needs to be satisfied for each unit cell, \ie{} the interfaces between them.} as introduced in~\cref{sub:Auxiliary Problem}
 \begin{equation}
-    \bm{F}_{\m}:= \dv{\bm{x}^{\m}}{\bm{x}^{\p}}\,,
+    \bm{F}_{\m}:= \dv{\bm{x}^{\m}}{\bm{x}^{\p}}\quad\mathrm{and}\quad
+    \dd{\bm{x}^{\m}}=\det(\bm{F}_{\m})\dd{\bm{x}^{\p}}\,,
     \label{eq:F_mu}
@@
 \begin{equation}
-    \int_{\varOmega^{\p}} \bm\varepsilon(\updelta\bm{u})\vdot\!\vdot\bm\sigma(\bm{u}_0(\m)) \dd{\bm{x}_{\m}} = -\int_{\varOmega^{\p}} \bm\varepsilon(\updelta\bm{u})\vdot\!\vdot\bm\sigma(\hat{\bm{g}})\dd{\bm{x}_{\m}}\,.
+    \int_{\varOmega^{\p}} \left(\pdv{\updelta\bm{u}}{\bm{x}^{\p}}\vdot\bm{F}^{-1}_{\m}\right)\vdot\!\vdot \mathbb{C} \vdot\!\vdot
+    \left(\pdv{\bm{u}(\m)}{\bm{x}^{\p}}\vdot\bm{F}^{-1}_{\m}\right)\det(\bm{F}_{\m})\dd{\bm{x}^{\p}} = \bm{0}\,.
     \label{eq:weak_osp_parent}
 \end{equation}
+The resolution of~\cref{eq:weak_osp_parent} and subsequent restriction of the solution to the target subdomain, denoted as $\bm{u}(\m)\big\vert_{\varOmega^{\p}_{\inrm}}$, comprises the solution of the oversampling problem.
+The transfer operator $\bm{T}_{\m}: \mathcal{S}\to\mathcal{R}$, maps the boundary data $\bm{g}$ to the solution on the target subdomain $\bm{u}(\m)\big\vert_{\varOmega^{\p}_{\inrm}} = \bm{T}_{\m}(\bm{g}).$
+Here, $\mathcal{S}$ and $\mathcal{R}$ are appropriate source and range spaces, respectively.
+Throughout the paper, we assume a suitable discretization and numerical solution of~\cref{eq:weak_osp_parent} using the FE method.
+For details concerning the discretization of the transfer operator (for a fixed parameter value) it is referred to~\cite{Buhr2018Randomized}.
+\added[id=pd]{Mention additional Neumann problem, data to be added to the final basis. Introduce terminology of spectral basis as part without Neumann}
 % - setting, global domain = union of non overlapping subdomains = unit cells
@@
 % - definition of oversampling problem and transfer operator
-
 % these are details of the specific problem we solve later
@@
 
-\added[id=pd]{TODO: add footnote about extension of Auxiliary Problem from unit cell to global domain / oversampling domain.}
+\begin{figure}
+    \centering
+    \include{./img/oversampling_domain.tex}
+    \caption{Exemplary coarse grid discretization of the global domain $\varOmega^{\m}_{\gl}$, with oversampling domain $\varOmega^{\m}$ in grey (a) and fine grid discretization of the unit cell (b).}\label{fig:oversampling_domain}
+\end{figure}
 
-% oversampling problem pulled back to parent domain
-\begin{align}
-	\begin{split}
-		\label{eq:osp_mu}
-		u &= 0 \,,\\
-		v &= 1
-	\end{split}
-\end{align}
+% TODO adjust abovecaptionskip if necessary
+% it seems the default is 6pt (see class file)
+% \setlength{\abovecaptionskip}{15pt plus 3pt minus 2pt}
 
@@
 \label{sec:Randomized range finder and POD}
+
+\begin{itemize}
+    \item assume discretization of T
+    \item explain challenge of finding appropriate space
+    \item go with idea from Taddei \& Patera, but range approx. instead of TE
+\end{itemize}
 
@@
 
-test
-\comment[id=pd]{TODO: check how/when Neumann data is added in case of RRFPOD. This can also be done outside the loop and even after POD of spectral basis.}
+\added[id=pd]{TODO: Given a discretization of T, we can now discuss how the range can be approximated.}
+
+\added[id=pd]{TODO: check how/when Neumann data is added in case of RRFPOD. This can also be done outside the loop and even after POD of spectral basis.}
+\added[id=pd]{TODO: FIX neumann data in the algorithm.}
 
@@
 
+\begin{itemize}
+    \item challenge that local spaces are not conforming on the interface
+    \item use GFEM as suitable method to construct conforming global approximation
+    \item how much detail should be necessary to explain GFEM?
+\end{itemize}
+
 % subsection Construction of a global approximation (end)
@@
 \label{sub:Online efficiency}
+
+\begin{itemize}
+    \item explain approach based on \textit{super elements}: each unit cell is just a quadrilateral cell with special ansatz functions
+    \item challenge: would still need to assemble $K$ of unit cell for each new parameter
+    \item therefore, employ EIM to speed up assembly of $K$ (reduced integration)
+\end{itemize}
 
@@
 
+\begin{itemize}
+    \item general stuff like FE cell type and degree
+    \item software: dolfinx version 0.8.0
+    \item PC hardware?
+\end{itemize}
+
 % section numerical experiments (end)
@@
 \label{sub:Graded concrete slab}
+
+\begin{itemize}
+    \item introduce graded concrete / problem (gradientenbeton eu); why this is interesting
+    \item sketch of the mechanical problem
+    \item setup regarding the new method?
+\end{itemize}
 

### 2024-08-25 src/parageom/pp_stress.py
[tasks] add optimization & pp_stress

@@ -4,6 +4,7 @@ import dolfinx as df
 import basix
 
 import numpy as np
+from pymor.core.pickle import load
 from pymor.vectorarrays.numpy import NumpyVectorSpace
 from pymor.operators.constructions import VectorOperator, LincombOperator, VectorFunctional, ConstantOperator
 from pymor.parameters.functionals import GenericParameterFunctional
@@ -50,14 +51,19 @@ def main(args):
     # ### Global function for displacment
     displacement = df.fem.Function(V)
 
-    # ps = fom.parameters.space(example.mu_range)
-    # mu = ps.sample_randomly(1)[0]
-    mu = fom.parameters.parse([0.1 for _ in range(10)])
+    # ### Get optimal solution μ*
+    fom_data = example.fom_minimization_data
+    with fom_data.open("rb") as fh:
+        data = load(fh)
+    mu = fom.parameters.parse(data["mu_min"])
 
-    s1_fom, s2_fom, s_fom = compute_principal_stress(fom, mu, displacement, stress, parageom)
+    # targets
+    xdmf_files = example.pp_stress(args.name)
+
+    s1_fom, s2_fom, s_fom = compute_principal_stress(fom, mu, displacement, stress, parageom, rec_data=None, xdmf_filename=xdmf_files["fom"].as_posix())
     displacement.x.array[:] = 0. # type: ignore
     stress.x.array[:] = 0. # type: ignore
-    s1_rom, s2_rom, s_rom = compute_principal_stress(rom, mu, displacement, stress, parageom, rec_data=rec_data)
+    s1_rom, s2_rom, s_rom = compute_principal_stress(rom, mu, displacement, stress, parageom, rec_data=rec_data, xdmf_filename=xdmf_files["rom"].as_posix())
 
     # stress error in euclidean norm
     def compute_norms(fom, rom):
@@ -81,12 +87,12 @@ def main(args):
     error.x.array[::2] /= sx_max # type: ignore
     error.x.array[1::2] /= sy_max # type: ignore
 
-    with df.io.XDMFFile(Q.mesh.comm, "output/stress_error.xdmf", "w") as xdmf: # type: ignore
+    with df.io.XDMFFile(Q.mesh.comm, xdmf_files["err"].as_posix(), "w") as xdmf: # type: ignore
         xdmf.write_mesh(Q.mesh)
         xdmf.write_function(error) # type: ignore
 
 
[…]

### 2024-08-25 src/parageom/tasks.py

@@ -298,3 +298,62 @@ def task_fig_locrom_error():
             "targets": [example.fig_locrom_error],
             "clean": True,
             }
+
+
+
+def task_optimization():
+    """ParaGeom: Determine optimal design"""
+    module = "src.parageom.optimization"
+    distr = example.distributions[0]
+
+    num_modes = 30 # TODO: rather use max value?
+    minimizer = "SLSQP"
+    omega = example.omega
+
+    nreal = 0 # do optimization only for single realization
+    for method in example.methods:
+        deps = [SRC / "optimization.py"]
+        deps.append(example.coarse_grid("global"))
+        deps.append(example.parent_domain("global"))
+        deps.append(example.parent_unit_cell)
+        for cell in range(5):
+            deps.append(example.local_basis_npy(nreal, method, distr, cell))
+        deps.append(example.local_basis_dofs_per_vert(nreal, method, distr))
+        targets = [example.fom_minimization_data,
+                   example.rom_minimization_data(distr, method)]
+        yield {
+                "name": method,
+                "file_dep": deps,
+                "actions": ["python3 -m {} {} {} {} --method {} --omega {}".format(module, distr, method, num_modes, minimizer, omega)],
+                "targets": targets,
+                "clean": True,
+                }
+
+
+def task_pp_stress():
+    """ParaGeom: Post-process stress"""
+    module = "src.parageom.pp_stress"
+    distr = "normal"
+    num_modes = 60 # local basis size
+    omega = example.omega # weighting factor for output functional
+    nreal = 0
+    for method in example.methods:
+        deps = [SRC / "pp_stress.py"]
+        # mesh and basis deps to construct rom
+        deps.append(example.coarse_grid("global"))
+        deps.append(example.parent_domain("global"))
+        deps.append(example.parent_unit_cell)
+        for cell in range(5):
+            deps.append(example.local_basis_npy(nreal, method, distr, cell))
+        deps.append(example.local_basis_dofs_per_vert(nreal, method, distr))
+        # optimization result
+        deps.append(example.fom_minimization_data)
+        # xdmf files as targets
+        targets = list(example.pp_stress(method).values())
+        yield {
+                "name": method,
+                "file_dep": deps,
+                "actions": ["python3 -m {} {} {} {} --omega {}".format(module, distr, method, num_modes, omega)],
+                "targets": targets,
+                "clean": True,
+                }

### 2024-08-25 src/parageom/definitions.py
[definitions] adjust neumann rtol

@@ -86,7 +86,7 @@ class BeamData:
     rrf_ttol: float = 10e-2
     rrf_ftol: float = 1e-10
     rrf_num_testvecs: int = 20
-    neumann_rtol: float = 1e-5
+    neumann_rtol: float = 1e-8
     run_mode: str = "DEBUG"
 
     def __post_init__(self):
@@ -221,9 +221,11 @@ class BeamData:
     def ntrain(self, config: str) -> int:
         """Define size of training set"""
         if self.run_mode == "DEBUG":
+            # 20 samples per unit cell
             map = {"left": 60, "inner": 80, "right": 60}
             return map[config]
         elif self.run_mode == "PRODUCTION":
+            # 50 samples per unit cell
             map = {"left": 150, "inner": 200, "right": 150}
             return map[config]
         else:
@@ -261,6 +263,15 @@ class BeamData:
     def rom_minimization_data(self, distr: str, name: str) -> Path:
         """ROM minimization data"""
         return self.rf / f"rom_minimization_data_{distr}_{name}.out"
+
+    def pp_stress(self, method: str) -> dict[str, Path]:
+        """Postprocessing of stress at optimal design"""
+        folder = self.method_folder(0, method)
+        return {
+                "fom": folder / "stress_fom.xdmf",
+                "rom": folder / "stress_rom.xdmf",
+                "err": folder / "stress_err.xdmf"
+                }
 
     # @property
     # def minimization_data_table(self) -> Path:

### 2024-08-25 paper/paper.tex
[sec 2.2] wip until 2.2.1

@@
 In the context of lattice structures, each subdomain $\varOmega_i^{\bm\mu}$ corresponds to a unit cell.
-Next, a coarse grid partition of the global domain and a fine grid partition of the unit cell is introduced as shown in~\added[id=pd]{TODO: add figure showing coarse grid and fine grid for a single unit cell.}
-\added[id=pd]{Fig. XX} also shows one exemplary oversampling domain $\varOmega^{\bm\mu}$.
+Next, a coarse grid partition of the global domain and a fine grid partition of the unit cell is introduced as shown in~\cref{fig:oversampling_domain}.
+\Cref{fig:oversampling_domain} also shows one exemplary oversampling domain $\varOmega^{\bm\mu}$.
 The oversampling problem (given for linear elastostatics) then comprises the solution of the following boundary value problem.
@@
 The boundaries $\varGamma^{\m}_{\mathrm{N}}$ and $\varGamma^{\m}_{\mathrm{D}}$ denote the part of the boundary of the oversampling domain that coincides with the global \textsc{Neumann} boundary $\varSigma^{\m}_{\mathrm{N}}$ or \textsc{Dirichlet} boundary $\varSigma^{\m}_{\mathrm{D}}$, respectively.
-Note that the topology is dependent on the target subdomain $\varOmega^{\m}_{\inrm}$ and the size of the oversampled region (\added[id=pd]{TODO: refer to figure as above}).
+Note that the topology is dependent on the target subdomain $\varOmega^{\m}_{\inrm}$ and the size of the oversampled region.
 Considering linear elastic isotropic materials
@@
 The resolution of~\cref{eq:weak_osp_parent} and subsequent restriction of the solution to the target subdomain, denoted as $\bm{u}(\m)\big\vert_{\varOmega^{\p}_{\inrm}}$, comprises the solution of the oversampling problem.
-The transfer operator $\bm{T}_{\m}: \mathcal{S}\to\mathcal{R}$, maps the boundary data $\bm{g}$ to the solution on the target subdomain $\bm{u}(\m)\big\vert_{\varOmega^{\p}_{\inrm}} = \bm{T}_{\m}(\bm{g}).$
+The transfer operator $\bm{T}_{\m}: \mathcal{S}\to\mathcal{R}$, maps for a given parameter value $\m$ the boundary data $\bm{g}$ to the solution on the target subdomain $\bm{u}(\m)\big\vert_{\varOmega^{\p}_{\inrm}} = \bm{T}_{\m}(\bm{g}).$
 Here, $\mathcal{S}$ and $\mathcal{R}$ are appropriate source and range spaces, respectively.
@@
 For details concerning the discretization of the transfer operator (for a fixed parameter value) it is referred to~\cite{Buhr2018Randomized}.
-\added[id=pd]{Mention additional Neumann problem, data to be added to the final basis. Introduce terminology of spectral basis as part without Neumann}
+
+Finally, we note that the images of $\bm{T}_{\m}$ are computed, such that they do not contain any translations or rotations (rigid body modes are removed via orthogonal projection), \ie{} they comprise purely \textit{deformational modes}.
+Moreover, (global) \textsc{Neumann} or \textsc{Dirichlet} boundary conditions prescribed on $\varSigma^{\m}_{\mathrm{N}}$ or $\varSigma^{\m}_{\mathrm{D}}$, respectively, have not been considered yet.
+Therefore, for example, an additional \textsc{Neumann} problem\footnote{The case of inhomogeneous \textsc{Dirichlet} boundary conditions is analogous.} has to be solved to enrich the local spaces.
+\begin{align}
+	\begin{split}
+	\label{eq:neumann_problem}
+    - \div{\bm{\sigma}} (\bm{u}(\bm\mu)) &= \bm{0} \quad  \mathrm{in}\;{\varOmega^{\bm\mu}}\subset\varOmega^{\bm\mu}_{\mathrm{gl}}\,,\\
+    \bm{\sigma}(\bm{u}(\bm\mu)) \vdot \bm{n} &= \hat{\bm{t}} \quad \mathrm{on}\; \varGamma^{\m}_{\mathrm{N}}:= \partial\varOmega^{\m}\cap\varSigma^{\m}_{\mathrm{N}}\,,\\
+    \bm{u}(\bm\mu) &= \bm{0} \quad  \mathrm{on}\;\varGamma^{\m}_{\mathrm{D}}:=\partial\varOmega^{\m}\cap\varSigma^{\m}_{\mathrm{D}}\,,\\
+    \bm{u}(\bm\mu) &= \bm{0}\quad\mathrm{on}\;\varGamma^{\m}_{\mathrm{out}}:= \partial\varOmega^{\m} \setminus\partial\varOmega^{\m}_{\mathrm{gl}}\,.
+	\end{split}
+\end{align}
+\Cref{eq:neumann_problem} is solved on the parent domain employing a pull back as above.
+Also, for construction of a sufficiently rich local basis it has to be solved for many parameter values $\m$, but a sampling of the boundary values on $\varGamma^{\m}_{\out}$ is not required.
 % - setting, global domain = union of non overlapping subdomains = unit cells
@@
     \include{./img/oversampling_domain.tex}
-    \caption{Exemplary coarse grid discretization of the global domain $\varOmega^{\m}_{\gl}$, with oversampling domain $\varOmega^{\m}$ in grey (a) and fine grid discretization of the unit cell (b).}\label{fig:oversampling_domain}
+    \caption{Exemplary coarse grid discretization of the global domain $\varOmega^{\m}_{\gl}$, with target subdomain $\varOmega^{\m}_i$ within oversampling domain $\varOmega^{\m}$ in grey (a) and fine grid discretization of the unit cell (b).}\label{fig:oversampling_domain}
 \end{figure}
@@
     \State{$\bm{R}\gets$ \textsc{AdaptiveRandomizedRangeApproximation}($\bm{T}_{\bm\mu_j}$, $\texttt{tol}$, $n_t$, $\varepsilon_{\mathrm{algofail}}$)}
-    \State{$\bm{u}_{\mathrm{Neumann}}(\bm\mu_j)\gets \bm{T}_{\bm\mu_j}(\bm{0})$}\Comment{solve additional Neumann problem if $\varGamma_{\mathrm{N}}\ne\emptyset$}
-    \State{$\bm{R}\gets \bm{R}\cup\left(\bm{u}_{\mathrm{Neumann}}(\bm\mu_j)\right)$}
-    \State{$\bm{S}\gets \bm{S}\cup\left(\mathrm{orthonormalize}\left(\bm{R}\right)\right)$}
+    % \State{$\bm{u}_{\mathrm{Neumann}}(\bm\mu_j)\gets \bm{T}_{\bm\mu_j}(\bm{0})$}\Comment{solve additional Neumann problem if $\varGamma_{\mathrm{N}}\ne\emptyset$}
+    % \State{$\bm{R}\gets \bm{R}\cup\left(\bm{u}_{\mathrm{Neumann}}(\bm\mu_j)\right)$}
+    \State{$\bm{S}\gets \mathrm{orthonormalize}\left(\bm{S}\cup\bm{R}\right)$}
     \EndFor%

### 2024-08-25 src/parageom/run_locrom.py
some tweaks

@@ -60,7 +60,7 @@ def main(args):
         # FIXME
         # store data of deim somewhere
         with Timer("EI of subdomain operator") as t:
-            mops, interpolation_matrix, idofs, magic_dofs, deim_data = interpolate_subdomain_operator(example, operator_local, design="uniform", ntrain=501, modes=None, atol=0., rtol=1e-12, method="method_of_snapshots")
+            mops, interpolation_matrix, idofs, magic_dofs, deim_data = interpolate_subdomain_operator(example, operator_local, design="uniform", ntrain=501, modes=None, atol=0., rtol=1e-8, method="method_of_snapshots")
             logger.info(f"EI of subdomain operator took {t.elapsed()[0]}.")
         m_dofs, m_inv = np.unique(magic_dofs, return_inverse=True)
         logger.debug(f"{magic_dofs=}")

### 2024-08-25 src/parageom/tasks.py

@@ -306,7 +306,7 @@ def task_optimization():
     module = "src.parageom.optimization"
     distr = example.distributions[0]
 
-    num_modes = 30 # TODO: rather use max value?
+    num_modes = 100 # TODO: rather use max value?
     minimizer = "SLSQP"
     omega = example.omega
 
@@ -334,7 +334,7 @@ def task_pp_stress():
     """ParaGeom: Post-process stress"""
     module = "src.parageom.pp_stress"
     distr = "normal"
-    num_modes = 60 # local basis size
+    num_modes = 100 # local basis size
     omega = example.omega # weighting factor for output functional
     nreal = 0
     for method in example.methods:

### 2024-08-25 src/parageom/parse_opt_data_mu_min.py
add parse_opt_data_mu_min.py

@@ -1,0 +1,12 @@
+from .tasks import example
+from pymor.core.pickle import load
+
+with example.fom_minimization_data.open('rb') as fh:
+    fomdata = load(fh)
+
+rom_data_path = example.rom_minimization_data("normal", "hapod")
+with rom_data_path.open('rb') as fh:
+    romdata = load(fh)
+
+print(f"FOM optimum: {fomdata['mu_min']}")
+print(f"ROM optimum: {romdata['mu_N_min']}")

### 2024-08-26 paper/paper.tex
wip: sec 2

@@
 It combines the aforementioned algorithm to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final reduced order model (ROM).
-Furthermore, the global approximation space is constructed from the local spaces using the generalized finite element method (GFEM)~\cite{Babuska2004Generalized}.
+Furthermore, the global approximation space is constructed from the local spaces using the partition of unity method or generalized finite element method (GFEM)~\cite{BABUSKA1997,Babuska2004Generalized}.
 
@@
 \end{equation}
-the weak form reads \added[id=pd]{TODO: is it necessary to write Dirichlet lift explicitly?}
+the weak form reads
 \begin{equation}
@@
 \end{align}
-\Cref{eq:neumann_problem} is solved on the parent domain employing a pull back as above.
+\Cref{eq:neumann_problem} is solved on the parent domain employing the pull back as above.
 Also, for construction of a sufficiently rich local basis it has to be solved for many parameter values $\m$, but a sampling of the boundary values on $\varGamma^{\m}_{\out}$ is not required.
@@
 
-\begin{itemize}
-    \item assume discretization of T
-    \item explain challenge of finding appropriate space
-    \item go with idea from Taddei \& Patera, but range approx. instead of TE
-\end{itemize}
+Assuming a proper FE discretization of $\bm{T}_{\m}$, next the algorithm to approximate the range of $\bm{T}_{\m}$ for any $\m$ is discussed.
+Herein, a strategy for the exploration of the parameter space and sampling of the random boundary conditions in the oversampling are the key challenges.
+In the present work, the strategy by Taddei \& Patera~\cite{Taddei2018Localization} is adopted in which based on the connection between the POD of images of the transfer operator (for certain boundary conditions) and the singular values of the transfer operator they devise a training strategy which is recognized as a distributed approximate POD (see \cite{Himpe2018Hierarchical}).
+For each $\m_j$ in a predefined training set, they compute the singular values of $\bm{T}_{\m_j}$ by resolution of an eigenvalue problem and subsequently compute the POD over the set of all eigenvectors.
+As the authors state, the resolution of the eigenvalue problems may be prohibitive in some cases, and therefore we employ range approximation by random sampling instead.
+
+The procedure for the computation of parameter-independent deformational modes is summarized in~\cref{algo:rrfpod}.
+First, an empty snapshot set $\bm{S}$ and a training set $\bm{S}_{\mathrm{train}}$ are initialized.
+The training set $\bm{S}_{\mathrm{train}}$ has to be sufficiently rich, such that it is a good representation of the parameter space.
+\comment[id=pd]{TODO: add reference to Haasdonk (standard sampling)?}
+Next, for each parameter value $\m_j\in\bm{S}_{\mathrm{train}}$ the range of $\bm{T}_{\m_j}$ which is now fixed with respect to the parameter value is approximated using Algorithm 1 \textsc{AdaptiveRandomizedRangeApproximation} from~\cite{Buhr2018Randomized} and the set of snapshots is extended by Gram Schmidt orthogonalization (line 5 \& 6).
+Note that the parameter value $\m$ and boundary conditions $\bm{g}$ are not varied at the same time.
+Finally, the POD is computed over the snapshot set $\bm{S}$ which yields the parameter-independent space $X^n$.
+The dimension of the local space $X^n$, \ie{} the number of local basis functions is denoted by $n$.
+
+In case of inhomogeneous \textsc{Neumann} boundary conditions, after termination of~\cref{algo:rrfpod}, the \textsc{Neumann} problem~\cref{eq:neumann_problem} is solved for each $\m_j\in\bm{S}_{\mathrm{train}}$ as well.
+Following compression via POD the space $X^n$ is enriched by the additional \textsc{Neumann} modes.
 
@@
 
-\added[id=pd]{TODO: Given a discretization of T, we can now discuss how the range can be approximated.}
-
-\added[id=pd]{TODO: check how/when Neumann data is added in case of RRFPOD. This can also be done outside the loop and even after POD of spectral basis.}
-\added[id=pd]{TODO: FIX neumann data in the algorithm.}
+\comment[id=pd]{TODO: explain tolerances to control approximation quality. Refer to HAPOD paper and mention that $\varepsilon^{\ast}$ can be used to compute tolerances at each node of the tree.}
+\comment[id=pd]{TODO: define notation for POD.}
 
@@
 \label{sub:Construction of a global approximation}
-
-\begin{itemize}
-    \item challenge that local spaces are not conforming on the interface
-    \item use GFEM as suitable method to construct conforming global approximation
-    \item how much detail should be necessary to explain GFEM?
-\end{itemize}
+Given local approximation spaces $X^n_i$ for each target subdomain $\varOmega_i^{\p}$, the construction of a global approximation via the GFEM is discussed.
+The difficulty is that the basis functions of the local spaces $X^n_i$ and $X^n_j$ of two neighboring subdomains $\varOmega_i^{\p}$ and $\varOmega_j^{\p}$ are not conforming on the shared interface $\varGamma^{\p}_{ij}=\partial\varOmega_i^{\p}\cap\partial\varOmega_j^{\p}$.
+However, given a suitably\footnote{According to Definition 1 of~\cite{BABUSKA1997}.} defined partition of unity $\{{\varphi}_i\}$ the local basis functions can be included in the global FE space by multiplication with the standard FE shape functions ${\varphi}_i$, such that a conforming approximation is obtained.
+The global GFEM space is defined as
+\begin{equation}
+	\label{eq:Xgfem}
+	X_{\mathrm{GFEM}}:= \sum_{i} \varphi_iX^n_i = \left\lbrace\sum_i\varphi_i\bm\xi_i \;\big|\; \bm\xi_i\in X^n_i	\right\rbrace\,.
+\end{equation}
+\comment[id=pd]{TODO: add figure. This explains best what is going on.}
+\comment[id=pd]{TODO: equation for partition of unity and final gfem space. Not sure if I can write this like Babuska.}
 
@@
 \label{sub:Online efficiency}
-
+With the global GFEM space, each coarse grid cell (\ie{} unit cell of the lattice) can be viewed as a high order FE.
+For example, it is just a quadrilateral cell with special ansatz functions 
 \begin{itemize}

### 2024-08-26 src/parageom/definitions.py
[WIP] fix kappa

@@ -71,7 +71,7 @@ class BeamData:
     )
     training_set_seed: int = 767667058
     testing_set_seed: int = 545445836
-    validation_set_seed: int = 9818890
+    validation_set_seed: int = 241690
     projerr_seed: int = 923719053
     configurations: tuple[str, str, str] = ("left", "inner", "right")
     distributions: tuple[str, ...] = ("normal",)
@@ -87,6 +87,7 @@ class BeamData:
     rrf_ftol: float = 1e-10
     rrf_num_testvecs: int = 20
     neumann_rtol: float = 1e-8
+    mdeim_rtol: float = 1e-5
     run_mode: str = "DEBUG"
 
     def __post_init__(self):
@@ -220,16 +221,14 @@ class BeamData:
 
     def ntrain(self, config: str) -> int:
         """Define size of training set"""
-        if self.run_mode == "DEBUG":
-            # 20 samples per unit cell
-            map = {"left": 60, "inner": 80, "right": 60}
-            return map[config]
-        elif self.run_mode == "PRODUCTION":
-            # 50 samples per unit cell
-            map = {"left": 150, "inner": 200, "right": 150}
-            return map[config]
-        else:
-            raise NotImplementedError
+        # num_samples = 51 # per unit cell; step size 0.004
+        num_samples = 300
+        if self.run_mode == "PRODUCTION":
+            num_samples = 101 # per unit cell; step size 0.002
+        # map = {"left": 3 * num_samples, "inner": 4 * num_samples, "right": 3 * num_samples}
+        map = {"left": 1000, "inner": 2000, "right": 1000}
+        # 2000 is too much though ...
+        return map[config]
 
     def cell_to_config(self, cell: int) -> str:
         """Maps global cell index to config."""
@@ -329,6 +328,10 @@ class BeamData:
     def hapod_singular_values(self, nr: int, distr: str, conf: str) -> Path:
         """singular values of final POD"""
         return self.method_folder(nr, "hapod") / f"singular_values_{distr}_{conf}.npy"
+
+    def hapod_neumann_svals(self, nr: int, distr: str, conf: str) -> Path:
+        """singular values of POD of neumann data"""
+        return self.method_folder(nr, "hapod") / f"neumann_singular_values_{distr}_{conf}.npy"
 
     def hapod_modes_xdmf(self, nr: int, distr: str, config: str) -> Path:
         """modes of the final POD"""

### 2024-08-26 src/parageom/gfem.py

@@ -144,9 +144,9 @@ def main(args):
 
     def enrich_with_constant(rb, x=False, y=False):
         dim = rb.shape[1]
-        xmode = np.ones((1, dim), dtype=np.float32)
+        xmode = np.ones((1, dim), dtype=np.float64)
         xmode[:, 1::2 ] *= 0.
-        ymode = np.ones((1, dim), dtype=np.float32)
+        ymode = np.ones((1, dim), dtype=np.float64)
         ymode[:, ::2] *= 0.
         basis = rb.copy()
         if y:
@@ -203,9 +203,9 @@ def main(args):
             dx_omega[:, 1:] = np.zeros_like(dx_omega[:, 1:])
             logger.debug(f"{dx_omega=}")
             if vertex in left_boundary:
-                dx_omega = np.array([[a, 0, 0]], dtype=np.float32)
+                dx_omega = np.array([[a, 0, 0]], dtype=np.float64)
             if vertex in right_boundary:
-                dx_omega = np.array([[4 * a, 0, 0]], dtype=np.float32)
+                dx_omega = np.array([[4 * a, 0, 0]], dtype=np.float64)
             omega[config].translate(dx_omega)
             logger.debug(f"{config=}, \tomega.xmin={omega[config].xmin}")
 

### 2024-08-26 src/parageom/hapod.py

@@ -107,17 +107,13 @@ def adaptive_rrf_normal(
             (source_product.range.dim, source_product.source.dim),  # type: ignore
             matvec=mvinv,  # type: ignore
         )
-        lambda_min = eigsh(
-            L, sigma=0, which="LM", return_eigenvectors=False, k=1, OPinv=Linv
-        )[0]
+        lambda_min = eigsh(L, sigma=0, which="LM", return_eigenvectors=False, k=1, OPinv=Linv)[0]
 
     # NOTE tp.source is the full space, while the source product
     # is of lower dimension
     num_source_dofs = tp.rhs.dofs.size
     testfail = failure_tolerance / min(num_source_dofs, tp.range.dim)
-    testlimit = (
-        np.sqrt(2.0 * lambda_min) * erfinv(testfail ** (1.0 / num_testvecs)) * error_tol
-    )
+    testlimit = np.sqrt(2.0 * lambda_min) * erfinv(testfail ** (1.0 / num_testvecs)) * error_tol
 
     logger.info(f"{lambda_min=}")
     logger.info(f"{testlimit=}")
@@ -127,9 +123,7 @@ def adaptive_rrf_normal(
         norm = error.norm2(product)
         return np.sum(norm) / len(U)
 
-    R = tp.generate_random_boundary_data(
-        count=num_testvecs, distribution=distribution, options=sampling_options
-    )
+    R = tp.generate_random_boundary_data(count=num_testvecs, distribution=distribution, options=sampling_options)
     M = tp.solve(R)
     B = tp.range.empty()
     maxnorm = np.inf
@@ -173,9 +167,7 @@ def main(args):
         loglevel = 20
 
     method = Path(__file__).stem  # hapod
-    logfilename = example.log_basis_construction(
-        args.nreal, method, args.distribution, args.configuration
-    ).as_posix()
+    logfilename = example.log_basis_construction(args.nreal, method, args.distribution, args.configuration).as_posix()
     set_defaults({"pymor.core.logger.getLogger.filename": logfilename})
     logger = getLogger(method, level=loglevel)
 
@@ -183,15 +175,11 @@ def main(args):
     training_seeds = {}
     for cfg, rndseed in zip(
         example.configurations,
-        np.random.SeedSequence(example.training_set_seed).generate_state(
-            len(example.configurations)
-        ),
+        np.random.SeedSequence(example.training_set_seed).generate_state(len(example.configurations)),
     ):
[…]

### 2024-08-26 src/parageom/lhs.py

@@ -25,7 +25,7 @@ def sample_lhs(parameter_space: ParameterSpace, name: str, **kwargs) -> list[Mu]
 
     """
     ndim = parameter_space.parameters[name]
-    ranges = np.array([parameter_space.ranges[name]], dtype=np.float32)
+    ranges = np.array([parameter_space.ranges[name]], dtype=np.float64)
     xrange = np.repeat(ranges, ndim, axis=0)
     samples = lhs(ndim, **kwargs)
     scale_range(samples, xrange)
@@ -39,7 +39,7 @@ def sample_lhs(parameter_space: ParameterSpace, name: str, **kwargs) -> list[Mu]
 
 if __name__ == "__main__":
     E_range = [1.0, 2.0]
-    xr = np.array([E_range], dtype=np.float32)
+    xr = np.array([E_range], dtype=np.float64)
     ndim = 2
     x = np.repeat(xr, ndim, axis=0)
     num_samples = 50

### 2024-08-26 src/parageom/locmor.py

@@ -16,14 +16,14 @@ import ufl
 from pymor.algorithms.gram_schmidt import gram_schmidt
 from pymor.algorithms.projection import project
 from pymor.bindings.fenicsx import FenicsxMatrixOperator, FenicsxVectorSpace
-from pymor.operators.constructions import VectorOperator
+from pymor.operators.constructions import VectorOperator, LincombOperator
 from pymor.operators.interface import Operator
 from pymor.operators.numpy import NumpyMatrixOperator
 from pymor.vectorarrays.interface import VectorArray
 from pymor.vectorarrays.numpy import NumpyVectorSpace
 from pymor.parameters.base import Parameters
 
-from multi.boundary import plane_at
+from multi.boundary import plane_at, point_at
 from multi.domain import RectangularDomain, StructuredQuadGrid
 from multi.dofmap import DofMap
 from multi.materials import LinearElasticMaterial
@@ -409,9 +409,17 @@ def assemble_gfem_system_with_ei(
 
     from .dofmap_gfem import select_modes
 
+    # TODO: try to set BCs to better condition system matrix
+
     # no need to apply bcs as basis functions should
     # satisfy these automatically
-    bc_dofs = np.array([], dtype=np.int32)
+    raise NotImplementedError("FIXME, condition number")
+    breakpoint()
+    left = dofmap.grid.locate_entities_boundary(0, plane_at(0.0, "x"))
+    bc_dofs = []
+    for vertex in left:
+        bc_dofs += dofmap.entity_dofs(vertex)
+    bc_dofs = np.array(bc_dofs, dtype=np.int32)
 
     lhs = defaultdict(list)
     rhs = defaultdict(list)
@@ -500,6 +508,8 @@ def assemble_gfem_system_with_ei(
     # coeff (M, 10)
     # indexptr --> defining 10 ranges within [0, nnz-1] that corresponds to values for each subdomain
 
+    # ### LHS
+    # Matrix operator
     rows = np.array(lhs["rows"], dtype=np.int32)
     cols = np.array(lhs["cols"], dtype=np.int32)
     indexptr = np.array(lhs["indexptr"], dtype=np.int32)
@@ -517,7 +527,17 @@ def assemble_gfem_system_with_ei(
         solver_options=options,
         name="K",
     )
+    # BC operator
+    bc_array = coo_array(
[…]

### 2024-08-26 src/parageom/optimization.py

@@ -151,7 +151,7 @@ def main(args):
     # ### Write outputs
     fom_minimization_data["method"] = args.method
     rom_minimization_data["method"] = args.method
-    rom_minimization_data["num_modes"] = args.num_modes
+    rom_minimization_data["num_modes"] = np.average(rec_data[0].dofs_per_vert)
 
     with example.fom_minimization_data.open("wb") as fh:
         dump(fom_minimization_data, fh)
@@ -337,7 +337,7 @@ def build_localized_rom(cli, example, global_auxp, trafo_disp, parameters, ω=0.
     operator_local, rhs_local = discretize_subdomain_operators(example)
     u_local = df.fem.Function(operator_local.source.V)
 
-    mops, interpolation_matrix, idofs, magic_dofs, deim_data = interpolate_subdomain_operator(example, operator_local, design="uniform", ntrain=501, modes=None, atol=0., rtol=1e-12, method="method_of_snapshots")
+    mops, interpolation_matrix, idofs, magic_dofs, deim_data = interpolate_subdomain_operator(example, operator_local, design="uniform", ntrain=501, modes=None, atol=0., rtol=example.mdeim_rtol, method="method_of_snapshots")
     m_dofs, m_inv = np.unique(magic_dofs, return_inverse=True)
     restricted_op, _ = operator_local.restricted(m_dofs, padding=1e-8)
     wrapped_op = EISubdomainOperatorWrapper(restricted_op, mops, interpolation_matrix, magic_dofs, m_inv)
@@ -377,7 +377,11 @@ def build_localized_rom(cli, example, global_auxp, trafo_disp, parameters, ω=0.
     assert max_dofs_per_vert.shape == (coarse_grid.num_cells, 4)
     assert np.allclose(np.array(bases_length), np.sum(max_dofs_per_vert, axis=1))
 
-    nmodes = cli.num_modes
+    Nmax = max_dofs_per_vert.max()
+    ΔN = 10
+    num_modes_per_vertex = list(range(Nmax // ΔN, Nmax + 1, 3 * (Nmax // ΔN) ))
+    nmodes = cli.num_modes or num_modes_per_vertex[-2] # second to last point in the validation
+
     dofs_per_vert = max_dofs_per_vert.copy()
     dofs_per_vert[max_dofs_per_vert > nmodes] = nmodes
     dofmap.distribute_dofs(dofs_per_vert)
[…]

### 2024-08-26 src/parageom/parse_opt_data_mu_min.py

@@ -10,3 +10,4 @@ with rom_data_path.open('rb') as fh:
 
 print(f"FOM optimum: {fomdata['mu_min']}")
 print(f"ROM optimum: {romdata['mu_N_min']}")
+breakpoint()

### 2024-08-26 src/parageom/plot_romerr.py

@@ -25,11 +25,16 @@ def main(cli):
     red = bamcd["red"][0]
 
     marker = {
-        "h1_semi": "x",
-        "max": "o",
+        "l2_err": "o",
+        "max_relerr_h1_semi": "s",
+        "max_relerr_max": "+",
     }
-    norm_label = {"h1_semi": "H1-semi", "max": "max"}
-    err_ = "max_relerr"
+    # marker_ei = {
+    #     "l2_err": "o",
+    #     "max_relerr_h1_semi": "o",
+    #     "max_relerr_max": "o",
+    # }
+    norm_label = {"max_relerr_h1_semi": "H1-semi", "max_relerr_max": "max"}
 
     args = [__file__, cli.outfile]
     styles = [example.plotting_style.as_posix()]
@@ -40,7 +45,7 @@ def main(cli):
             data = np.load(infile)
             num_dofs = data["ndofs"]
             for norm in cli.norm:
-                err = data["_".join([err_, norm])]
+                err = data[norm]
                 mark = marker[norm]
                 if method == "heuristic":
                     label = f"HRRF, {norm_label[norm]}"
@@ -55,7 +60,7 @@ def main(cli):
                 data = np.load(infile)
                 num_dofs = data["ndofs"]
                 for norm in cli.norm:
-                    err = data["_".join([err_, norm])]
+                    err = data[norm]
                     mark = marker[norm]
                     if method == "heuristic":
                         label = f"HRRF+EI, {norm_label[norm]}"

### 2024-08-26 src/parageom/pp_stress.py

@@ -193,7 +193,11 @@ def build_localized_rom(cli, example, global_auxp, trafo_disp, parameters, ω=0.
     assert max_dofs_per_vert.shape == (coarse_grid.num_cells, 4)
     assert np.allclose(np.array(bases_length), np.sum(max_dofs_per_vert, axis=1))
 
-    nmodes = cli.num_modes
+    Nmax = max_dofs_per_vert.max()
+    ΔN = 10
+    num_modes_per_vertex = list(range(Nmax // ΔN, Nmax + 1, 3 * (Nmax // ΔN) ))
+    nmodes = num_modes_per_vertex[-2] # second to last point in the validation
+
     dofs_per_vert = max_dofs_per_vert.copy()
     dofs_per_vert[max_dofs_per_vert > nmodes] = nmodes
     dofmap.distribute_dofs(dofs_per_vert)
@@ -201,7 +205,6 @@ def build_localized_rom(cli, example, global_auxp, trafo_disp, parameters, ω=0.
     operator, rhs, local_bases = assemble_gfem_system_with_ei(
         dofmap, wrapped_op, rhs_local, local_bases, dofs_per_vert, max_dofs_per_vert, parameters)
 
-    # TODO: add output functional
     # definition of ParaGeom Problem for volume computation
     omega = global_auxp.problem.domain
     matparam = {"gdim": omega.gdim, "E": 1.0, "NU": example.poisson_ratio, "plane_stress": example.plane_stress}
@@ -260,11 +263,6 @@ if __name__ == "__main__":
         type=str,
         help="The name of the training strategy.",
         choices=("hapod", "heuristic"),
-    )
-    parser.add_argument(
-        "num_modes",
-        type=int,
-        help="Local basis size to be used with local ROM.",
     )
     parser.add_argument(
             "--omega",

### 2024-08-26 src/parageom/run_locrom.py

@@ -14,6 +14,7 @@ from pymor.operators.constructions import VectorOperator
 from pymor.models.basic import StationaryModel
 from pymor.tools.random import new_rng
 
+
 def main(args):
     from .tasks import example
     from .auxiliary_problem import discretize_auxiliary_problem
@@ -24,7 +25,13 @@ def main(args):
     from .dofmap_gfem import GFEMDofMap
 
     # ### logger
-    set_defaults({"pymor.core.logger.getLogger.filename": example.log_run_locrom(args.nreal, args.method, args.distr, ei=args.ei)})
+    set_defaults(
+        {
+            "pymor.core.logger.getLogger.filename": example.log_run_locrom(
+                args.nreal, args.method, args.distr, ei=args.ei
+            )
+        }
+    )
     if args.debug:
         loglevel = "DEBUG"
     else:
@@ -60,7 +67,16 @@ def main(args):
         # FIXME
         # store data of deim somewhere
         with Timer("EI of subdomain operator") as t:
-            mops, interpolation_matrix, idofs, magic_dofs, deim_data = interpolate_subdomain_operator(example, operator_local, design="uniform", ntrain=501, modes=None, atol=0., rtol=1e-8, method="method_of_snapshots")
+            mops, interpolation_matrix, idofs, magic_dofs, deim_data = interpolate_subdomain_operator(
+                example,
+                operator_local,
+                design="uniform",
+                ntrain=501,
+                modes=None,
+                atol=0.0,
+                rtol=example.mdeim_rtol,
+                method="method_of_snapshots",
+            )
             logger.info(f"EI of subdomain operator took {t.elapsed()[0]}.")
         m_dofs, m_inv = np.unique(magic_dofs, return_inverse=True)
         logger.debug(f"{magic_dofs=}")
@@ -68,12 +84,12 @@ def main(args):
         wrapped_op = EISubdomainOperatorWrapper(restricted_op, mops, interpolation_matrix, magic_dofs, m_inv)
 
         # convert `rhs_local` to NumPy
-        vector = rhs_local.as_range_array().to_numpy() # type: ignore
+        vector = rhs_local.as_range_array().to_numpy()  # type: ignore
         rhs_va = mops[0].range.from_numpy(vector)
         rhs_local = VectorOperator(rhs_va)
 
[…]

### 2024-08-26 src/parageom/tasks.py

@@ -125,6 +125,8 @@ def task_hapod():
             )
             targets.append(example.hapod_modes_npy(nreal, distr, config))
             targets.append(example.hapod_singular_values(nreal, distr, config))
+            if config == "left":
+                targets.append(example.hapod_neumann_svals(nreal, distr, config))
             yield {
                 "name": config + ":" + str(nreal),
                 "file_dep": deps,
@@ -286,7 +288,7 @@ def task_fig_locrom_error():
     module = "src.parageom.plot_romerr"
     nreal = 0
     distr = "normal"
-    norm = "h1_semi"
+    norm = "max_relerr_h1_semi"
 
     deps = []
     for method in example.methods:
@@ -306,7 +308,6 @@ def task_optimization():
     module = "src.parageom.optimization"
     distr = example.distributions[0]
 
-    num_modes = 100 # TODO: rather use max value?
     minimizer = "SLSQP"
     omega = example.omega
 
@@ -324,7 +325,7 @@ def task_optimization():
         yield {
                 "name": method,
                 "file_dep": deps,
-                "actions": ["python3 -m {} {} {} {} --method {} --omega {}".format(module, distr, method, num_modes, minimizer, omega)],
+                "actions": ["python3 -m {} {} {} --method {} --omega {}".format(module, distr, method, minimizer, omega)],
                 "targets": targets,
                 "clean": True,
                 }
@@ -334,7 +335,6 @@ def task_pp_stress():
     """ParaGeom: Post-process stress"""
     module = "src.parageom.pp_stress"
     distr = "normal"
-    num_modes = 100 # local basis size
     omega = example.omega # weighting factor for output functional
     nreal = 0
     for method in example.methods:
@@ -353,7 +353,7 @@ def task_pp_stress():
         yield {
                 "name": method,
                 "file_dep": deps,
-                "actions": ["python3 -m {} {} {} {} --omega {}".format(module, distr, method, num_modes, omega)],
+                "actions": ["python3 -m {} {} {} --omega {}".format(module, distr, method, omega)],
                 "targets": targets,
                 "clean": True,
                 }

### 2024-08-26 src/parageom/test_dirichlet_lift.py

@@ -36,7 +36,7 @@ def test():
             print("pass")
 
     num_vals = n + 1
-    g = np.arange(num_vals, dtype=np.float32)
+    g = np.arange(num_vals, dtype=np.float64)
     compare(mat, lift, g)
     compare(mat, lift, np.random.rand(num_vals))
     compare(mat, lift, np.random.rand(num_vals))

### 2024-08-27 paper/paper.tex
finish first draft of methodology

@@
 \usepackage{graphicx}
+\usepackage{subfig}
 \usepackage{url}
@@
 \newcommand{\inrm}{\mathrm{in}}
+% operators
+\newcommand{\vecm}{\mathrm{vec}}
+\renewcommand{\top}{\mathrm{T}}
 
@@
 Moreover, (global) \textsc{Neumann} or \textsc{Dirichlet} boundary conditions prescribed on $\varSigma^{\m}_{\mathrm{N}}$ or $\varSigma^{\m}_{\mathrm{D}}$, respectively, have not been considered yet.
-Therefore, for example, an additional \textsc{Neumann} problem\footnote{The case of inhomogeneous \textsc{Dirichlet} boundary conditions is analogous.} has to be solved to enrich the local spaces.
+Therefore, for example, an additional \textsc{Neumann} problem\footnote{The case of inhomogeneous \textsc{Dirichlet} boundary conditions can be treated in the same way.} has to be solved to enrich the local spaces.
 \begin{align}
@@
 The procedure for the computation of parameter-independent deformational modes is summarized in~\cref{algo:rrfpod}.
+The target tolerance $\varepsilon^{\ast}$ is used to control the quality of the approximation space (by bounding the mean $\ell^2$-error), and to compute the local tolerances $\varepsilon_{\alpha}$ that steer the range approximation of each transfer operator and the tolerance $\varepsilon_{\rho}$ for the final POD.
 First, an empty snapshot set $\bm{S}$ and a training set $\bm{S}_{\mathrm{train}}$ are initialized.
-The training set $\bm{S}_{\mathrm{train}}$ has to be sufficiently rich, such that it is a good representation of the parameter space.
-\comment[id=pd]{TODO: add reference to Haasdonk (standard sampling)?}
+The training set $\bm{S}_{\mathrm{train}}$ has to be sufficiently rich, such that it is a good representation of the parameter space, and the reader is referred to~\cite{Haasdonk2017Chapter}.
 Next, for each parameter value $\m_j\in\bm{S}_{\mathrm{train}}$ the range of $\bm{T}_{\m_j}$ which is now fixed with respect to the parameter value is approximated using Algorithm 1 \textsc{AdaptiveRandomizedRangeApproximation} from~\cite{Buhr2018Randomized} and the set of snapshots is extended by Gram Schmidt orthogonalization (line 5 \& 6).
@@
 \begin{algorithm}
-    \caption{RRFPOD caption.}%
+    \caption{Randomized range finder (RRF) combined with POD.}%
     \label{algo:rrfpod}
     \begin{algorithmic}[1]
-        \Function{HapodRangeApproximation}{$\bm{T}$, $\texttt{tol}$, $n_t$, $\varepsilon_{\mathrm{algofail}}$, $N_{\mathrm{train}}$}
-    \Statex{\textbf{Input:} Operator $\bm{T}_{\bm\mu}$, target tolerance $\texttt{tol}$, number of testvectors $n_t$, maximum failure probability $\varepsilon_{\mathrm{algofail}}$, number of parameter samples $N_{\mathrm{train}}$}
+        \Function{HapodRangeApproximation}{$\bm{T}$, $\varepsilon^{\ast}$, $n_t$, $\varepsilon_{\mathrm{algofail}}$, $N_{\mathrm{train}}$}
+        \Statex{\textbf{Input:} Operator $\bm{T}_{\bm\mu}$, target tolerance $\varepsilon^{\ast}$, number of testvectors $n_t$, maximum failure probability $\varepsilon_{\mathrm{algofail}}$, number of parameter samples $N_{\mathrm{train}}$}
     \Statex{\textbf{Output:} space $X^n$}
 
+    \State{Compute $\varepsilon_{\rho}$ and $\varepsilon_{\alpha}$ from $\varepsilon^{\ast}$ according to Theorem 10 of~\cite{Himpe2018Hierarchical}}
     \State{$\bm{S}$ $\gets\;\emptyset$}\Comment{initialize snapshot set}
@@
     \For{$\bm\mu_j$ in $\bm{S}_{\mathrm{train}}$}
-    \State{$\bm{R}\gets$ \textsc{AdaptiveRandomizedRangeApproximation}($\bm{T}_{\bm\mu_j}$, $\texttt{tol}$, $n_t$, $\varepsilon_{\mathrm{algofail}}$)}
+    \State{$\bm{R}\gets$ \textsc{AdaptiveRandomizedRangeApproximation}($\bm{T}_{\bm\mu_j}$, $\varepsilon_{\alpha}$, $n_t$, $\varepsilon_{\mathrm{algofail}}$)}
     % \State{$\bm{u}_{\mathrm{Neumann}}(\bm\mu_j)\gets \bm{T}_{\bm\mu_j}(\bm{0})$}\Comment{solve additional Neumann problem if $\varGamma_{\mathrm{N}}\ne\emptyset$}
@@
     \EndFor%
-    \State{$\bm{B}\gets\textsc{POD}(\bm{S})$}
+    \State{$\bm{B}\gets\textsc{POD}(\bm{S}, \varepsilon_{\rho})$}
     \State{\textbf{return} {$X^n=\mathrm{span}\left(\bm{B}\right)$}}
@@
 
-\comment[id=pd]{TODO: explain tolerances to control approximation quality. Refer to HAPOD paper and mention that $\varepsilon^{\ast}$ can be used to compute tolerances at each node of the tree.}
-\comment[id=pd]{TODO: define notation for POD.}
+% \comment[id=pd]{TODO: explain tolerances to control approximation quality.
+Refer to HAPOD paper and mention that $\varepsilon^{\ast}$ can be used to compute tolerances at each node of the tree (epsilon star drives the RRF and the final POD.
+However, this means that we have to use l2-mean error as termination criterion.
+Too detailed?}
+% \comment[id=pd]{TODO: define notation for POD.}
 
@@
 \end{equation}
-\comment[id=pd]{TODO: add figure. This explains best what is going on.}
-\comment[id=pd]{TODO: equation for partition of unity and final gfem space. Not sure if I can write this like Babuska.}
+Furthermore, the GFEM function denoted as $\bm{\psi}(\bm{x}) = \varphi(\bm{x})\bm{\xi}(\bm{x}), \bm{x}\in\omega_{\alpha}$ for a patch $\omega_{\alpha}$ is shown in~\cref{fig:gfem} to illustrate the procedure of constructing the GFEM functions.
+\begin{figure}[h]
+    \centering
+    \subfloat[Linear hat function $\varphi(\bm{x})$]{%
+          \includegraphics[height=5cm]{phi.pdf}%
+          \label{fig:gfem_left}%
+       } 
+       \subfloat[Enrichment function $\bm{\xi}(\bm{x})$]{%
+          \includegraphics[height=5cm]{enrichment.pdf}%
+          \label{fig:gfem_middle}%
+       }
+       \subfloat[GFEM function $\bm\psi(\bm{x})$]{%
+          \includegraphics[height=5cm]{gfem_function.pdf}%
+          \label{fig:gfem_right}%
+       }
+       \caption{Construction of GFEM functions.}
+       \label{fig:gfem}
+\end{figure}
 
@@
 With the global GFEM space, each coarse grid cell (\ie{} unit cell of the lattice) can be viewed as a high order FE.
-For example, it is just a quadrilateral cell with special ansatz functions 
-\begin{itemize}
-    \item explain approach based on \textit{super elements}: each unit cell is just a quadrilateral cell with special ansatz functions
-    \item challenge: would still need to assemble $K$ of unit cell for each new parameter
-    \item therefore, employ EIM to speed up assembly of $K$ (reduced integration)
-\end{itemize}
+For instance, assuming a coarse grid partition as shown in~\cref{fig:oversampling_domain}, the coarse grid cell is a quadrilateral cell with special ansatz functions $\bm{\psi}_i$.
+The components of the FE matrix of the $i-$th coarse grid cell are given by
+\begin{align}
+    \begin{split}
+        a^i_{kl} &= \int_{\varOmega^{\p}_i} \left(\pdv{\bm{\psi}_k}{\bm{x}^{\p}}\vdot\bm{F}^{-1}_{\m}\right)\vdot\!\vdot \mathbb{C} \vdot\!\vdot
+    \left(\pdv{\bm{\psi}_l}{\bm{x}^{\p}}\vdot\bm{F}^{-1}_{\m}\right)\det(\bm{F}_{\m})\dd{\bm{x}^{\p}}\,,\\
+                 &= ({V}^\top)_{km} {K}_{mn}(\mu){V}_{nl}\,.
+    \end{split}
+    \label{eq:element_matrix}
+\end{align}
+Here, $\bm{V}$ is the matrix that holds the FE vectors of the GFEM functions $\bm\psi_i$ as column vectors and $\bm{K}(\mu)$ is the stiffness matrix of the subdomain $\varOmega^{\p}_i$ (\ie{} unit cell).
+Since $\bm{K}$ is dependent on the parameter $\m$, integration over the full fine scale mesh of the unit cell is required whenever the model is evaluated for a new parameter value which will drastically reduce the computational efficiency of our method.
+
+To overcome this problem, the stiffness matrix $\bm{K}(\m)$ is approximated by Empirical Interpolation~(see \cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} and in particular \cite{Negri2015Efficient} for the matrix version termed MDEIM employed here), such that
+\begin{equation}
+    \bm{K}(\m) \approx \bm{K}_Q(\m) =\sum_{q=1}^{Q} \theta_q(\m) \bm{K}_q\,.
+    \label{eq:K_ei}
+\end{equation}
+Herein, $\bm{K}_q$ are parameter-independent matrices that can be precomputed\footnote{Due to the Galerkin projection onto $\bm{V}$ (\cref{eq:element_matrix}) the matrices to be precomputed and stored are also sufficiently small.}.
+The $\bm\theta(\m)\in\mathbb{R}^Q$ are called interpolation coefficients and can be determined via the interpolation equation
+\begin{equation}
+    \bm{\Phi}_{\mathcal{I}}\bm\theta(\m) = \bm{k}_{\mathcal{I}}(\m)\,.
+    \label{eq:interp}
+\end{equation}
+Considering the matrix version of DEIM, the interpolation is constructed for matrices that are vectorized by stacking their columns on top of each other which is denoted by $\bm{k}=\vecm(\bm{K})$.
+The suffix $(\cdot)_{\mathcal{I}}$ indicates the restricted evaluation, reduced integration respectively.
+The basis $\bm{\Phi}=[\vecm(\bm{K}_1), \ldots, \vecm(\bm{K}_Q)]$ and the interpolation indices $\mathcal{I}$ are determined by applying the DEIM algorithm as proposed in~\cite{Chaturantabut2010Nonlinear} to a set of snapshots $\{\vecm(\bm{K}(\m_1)), \ldots, \vecm(\bm{K}(\m_m))\}$, $m>0$.
+By reversing the $\vecm(\cdot)$ operation, the approximation $\bm{K}_Q(\m)$ can be formed using~\cref{eq:K_ei} once the interpolation coefficients are determined by solution of \cref{eq:interp}.
+For more details on the overall procedure it is referred to~\cite{Negri2015Efficient}.
+
+Finally, with a fast method to compute $\bm{K}_Q(\m)$, the calculation of the FE matrices $a^i_{kl}$ of each coarse grid cell are independent of the underlying fine grid discretization of the unit cell.
+The assembly of the global system matrix follows the usual assembly routine for FE methods.
 

### 2024-08-28 src/parageom/auxiliary_problem.py
[WIP] current state 28.8

@@ -300,13 +300,13 @@ def main():
     # phyiscal domains/meshes
     # need to use same degree as degree that should be
     # used for the geometry interpolation afterwards
-    degree = example.geom_deg
+    assert example.fe_deg == example.geom_deg
 
     # discretize auxiliary problem for parent unit cell
     mshfile = example.parent_unit_cell.as_posix()
     ftags = {"bottom": 11, "left": 12, "right": 13, "top": 14, "interface": 15}
     param = {"R": 1}
-    auxp = discretize_auxiliary_problem(mshfile, degree, ftags, param)
+    auxp = discretize_auxiliary_problem(example, mshfile, ftags, param)
 
     # output function
     d = df.fem.Function(auxp.problem.V)
@@ -327,7 +327,7 @@ def main():
     global_coarse_domain = example.coarse_grid("global").as_posix()
     param = {"R": 10}
     int_tags = [i for i in range(15, 25)]
-    auxp = discretize_auxiliary_problem(global_domain_msh, degree, int_tags, param, coarse_grid=global_coarse_domain)
+    auxp = discretize_auxiliary_problem(example, global_domain_msh, int_tags, param, coarse_grid=global_coarse_domain)
 
     d = df.fem.Function(auxp.problem.V)
     xdmf = XDMFFile(d.function_space.mesh.comm, "./transformation_global_domain.xdmf", "w")

### 2024-08-28 src/parageom/definitions.py

@@ -50,7 +50,7 @@ class BeamData:
 
     name: str = "parageom"
     gdim: int = 2
-    l_char: float = 100.0  # [mm], characteristic length = unit length
+    l_char: float = 1.0  # [mm], characteristic length = unit length
     unit_length: float = 1.0  # dimensionless unit length
     nx: int = 10
     ny: int = 1
@@ -78,8 +78,8 @@ class BeamData:
     methods: tuple[str, ...] = ("hapod", ) # "heuristic")
     epsilon_star: dict = field(
             default_factory=lambda: {
-                "heuristic": 0.01,
-                "hapod": 0.01,
+                "heuristic": 0.001,
+                "hapod": 0.001,
                 })
     epsilon_star_projerr: float = 0.001
     omega: float = 0.5 # ω related to HAPOD (not output functional)
@@ -221,13 +221,10 @@ class BeamData:
 
     def ntrain(self, config: str) -> int:
         """Define size of training set"""
-        # num_samples = 51 # per unit cell; step size 0.004
-        num_samples = 300
+        num_samples = 50 # per unit cell; step size 0.004
         if self.run_mode == "PRODUCTION":
             num_samples = 101 # per unit cell; step size 0.002
-        # map = {"left": 3 * num_samples, "inner": 4 * num_samples, "right": 3 * num_samples}
-        map = {"left": 1000, "inner": 2000, "right": 1000}
-        # 2000 is too much though ...
+        map = {"left": 3 * num_samples, "inner": 5 * num_samples, "right": 3 * num_samples}
         return map[config]
 
     def cell_to_config(self, cell: int) -> str:

### 2024-08-28 src/parageom/fom.py

@@ -146,7 +146,7 @@ def discretize_subdomain_operators(example):
     d = df.fem.Function(aux.problem.V, name="d_trafo")
 
     omega = aux.problem.domain
-    matparam = {"gdim": omega.gdim, "E": 1.0, "NU": example.poisson_ratio, "plane_stress": example.plane_stress}
+    matparam = {"gdim": omega.gdim, "E": example.youngs_modulus, "NU": example.poisson_ratio, "plane_stress": example.plane_stress}
     problem = ParaGeomLinEla(omega, aux.problem.V, d, matparam)
 
     # ### wrap stiffness matrix as pymor operator
@@ -161,8 +161,9 @@ def discretize_subdomain_operators(example):
     )
 
     # ### wrap external force as pymor operator
-    EMOD = example.youngs_modulus
-    TY = -example.traction_y / EMOD
+    # EMOD = example.youngs_modulus
+    # TY = -example.traction_y / EMOD
+    TY = -example.traction_y
     traction = df.fem.Constant(
         omega.grid, (df.default_scalar_type(0.0), df.default_scalar_type(TY))
     )
@@ -202,7 +203,7 @@ def discretize_fom(example: BeamData, auxiliary_problem, trafo_disp, ω=0.5):
     assert neumann is not None
     facet_tags, _ = create_meshtags(domain, fdim, {"top": neumann})
     omega = RectangularDomain(domain, facet_tags=facet_tags)
-    matparam = {"gdim": omega.gdim, "E": 1.0, "NU": example.poisson_ratio, "plane_stress": example.plane_stress}
+    matparam = {"gdim": omega.gdim, "E": example.youngs_modulus, "NU": example.poisson_ratio, "plane_stress": example.plane_stress}
     problem = ParaGeomLinEla(omega, auxiliary_problem.problem.V, trafo_disp, matparam)
 
     # Dirichlet BCs
@@ -243,8 +244,7 @@ def discretize_fom(example: BeamData, auxiliary_problem, trafo_disp, ω=0.5):
     bcs = _create_dirichlet_bcs(tuple(dirichlet_bcs))
 
     # Neumann BCs
-    EMOD = example.youngs_modulus
-    TY = -example.traction_y / EMOD
+    TY = -example.traction_y
     traction = df.fem.Constant(
         domain, (df.default_scalar_type(0.0), df.default_scalar_type(TY))
     )
[…]

### 2024-08-28 src/parageom/hapod.py

@@ -129,14 +129,11 @@ def adaptive_rrf_normal(
     maxnorm = np.inf
     l2 = np.sum(M.norm2(range_product)) / len(M)
 
-    l2_errors = [
-        l2,
-    ]
-    max_norms = [
-        maxnorm,
-    ]
+    l2_errors = [l2,]
+    max_norms = [maxnorm,]
 
-    while (maxnorm > testlimit) and (l2 > l2_err**2.0):
+    while (l2 > l2_err**2.0):
+    # while (maxnorm > testlimit) and (l2 > l2_err**2.0):
         basis_length = len(B)
         v = tp.generate_random_boundary_data(1, distribution, options=sampling_options)
 
@@ -151,6 +148,8 @@ def adaptive_rrf_normal(
         max_norms.append(maxnorm)
 
     reason = "maxnorm" if maxnorm < testlimit else "l2err"
+    logger.info(f"{maxnorm < testlimit =}\t{maxnorm=}")
+    logger.info(f"{l2 < l2_err ** 2 =}\t{l2=}")
     logger.info(f"Finished RRF in {len(B)} iterations ({reason=}).")
 
     return B

### 2024-08-28 src/parageom/locmor.py

@@ -413,13 +413,21 @@ def assemble_gfem_system_with_ei(
 
     # no need to apply bcs as basis functions should
     # satisfy these automatically
-    raise NotImplementedError("FIXME, condition number")
-    breakpoint()
-    left = dofmap.grid.locate_entities_boundary(0, plane_at(0.0, "x"))
-    bc_dofs = []
-    for vertex in left:
-        bc_dofs += dofmap.entity_dofs(vertex)
-    bc_dofs = np.array(bc_dofs, dtype=np.int32)
+    # raise NotImplementedError("FIXME, condition number")
+    # left = dofmap.grid.locate_entities_boundary(0, plane_at(0.0, "x"))
+    # right = dofmap.grid.locate_entities_boundary(0, point_at([10., 0., 0.]))
+
+    # assert left.size == 2
+    # assert right.size == 1
+
+    # bc_dofs = []
+    # for vertex in left:
+    #     bc_dofs.append(dofmap.entity_dofs(vertex)[0])
+    # for vertex in right:
+    #     bc_dofs.append(dofmap.entity_dofs(vertex)[0])
+    #     bc_dofs.append(dofmap.entity_dofs(vertex)[1])
+    # bc_dofs = np.array(bc_dofs, dtype=np.int32)
+    bc_dofs = np.array([], dtype=np.int32)
 
     lhs = defaultdict(list)
     rhs = defaultdict(list)
@@ -528,14 +536,17 @@ def assemble_gfem_system_with_ei(
         name="K",
     )
     # BC operator
-    bc_array = coo_array(
-        (bc_mat["data"], (bc_mat["rows"], bc_mat["cols"])), shape=shape
-    )
-    bc_array.eliminate_zeros()
-    bc_op = NumpyMatrixOperator(
-        bc_array.tocsr(), op.source.id, op.range.id, op.solver_options, "bc_mat"
-    )
-    lhs_op = LincombOperator([op, bc_op], [1.0, 1.0])
+    if np.any(bc_dofs):
+        bc_array = coo_array(
+            (bc_mat["data"], (bc_mat["rows"], bc_mat["cols"])), shape=shape
+        )
+        bc_array.eliminate_zeros()
+        bc_op = NumpyMatrixOperator(
+            bc_array.tocsr(), op.source.id, op.range.id, op.solver_options, "bc_mat"
+        )
+        lhs_op = LincombOperator([op, bc_op], [1.0, 1.0])
+    else:
+        lhs_op = op
 
     # ### RHS
[…]

### 2024-08-28 src/parageom/optimization.py

@@ -380,7 +380,7 @@ def build_localized_rom(cli, example, global_auxp, trafo_disp, parameters, ω=0.
     Nmax = max_dofs_per_vert.max()
     ΔN = 10
     num_modes_per_vertex = list(range(Nmax // ΔN, Nmax + 1, 3 * (Nmax // ΔN) ))
-    nmodes = cli.num_modes or num_modes_per_vertex[-2] # second to last point in the validation
+    nmodes = cli.num_modes or num_modes_per_vertex[-2]
 
     dofs_per_vert = max_dofs_per_vert.copy()
     dofs_per_vert[max_dofs_per_vert > nmodes] = nmodes
@@ -391,7 +391,7 @@ def build_localized_rom(cli, example, global_auxp, trafo_disp, parameters, ω=0.
 
     # definition of ParaGeom Problem for volume computation
     omega = global_auxp.problem.domain
-    matparam = {"gdim": omega.gdim, "E": 1.0, "NU": example.poisson_ratio, "plane_stress": example.plane_stress}
+    matparam = {"gdim": omega.gdim, "E": example.youngs_modulus, "NU": example.poisson_ratio, "plane_stress": example.plane_stress}
     parageom = ParaGeomLinEla(omega, global_auxp.problem.V, trafo_disp, matparam)
 
     def param_setter(mu):

### 2024-08-28 src/parageom/projerr.py

@@ -78,7 +78,15 @@ def main(args):
 
         Nin = transfer.rhs.dofs.size
         epsilon_alpha = np.sqrt(Nin) * np.sqrt(1 - example.omega**2.0) * epsilon_star
-        epsilon_pod = epsilon_star * np.sqrt(Nin * ntrain)
+        # total number of input vectors is at most Nin * ntrain
+        epsilon_pod = np.sqrt(Nin * ntrain) * example.omega * epsilon_star
+        # but as usually much less vectors than Nin are computed per transfer operator
+        # however epsilon_pod can simply be computed after number of snapshots is known
+
+        # ε_α = np.sqrt(Nin) ... <-- cardinality of snapshot set
+        # l2-mean should be computed over set of size Nin then ...
+
+        # but since Nin is 
 
         # scaling
         epsilon_alpha /= example.l_char
@@ -92,13 +100,14 @@ def main(args):
                     transfer,
                     error_tol=example.rrf_ttol / example.l_char,
                     failure_tolerance=example.rrf_ftol,
-                    num_testvecs=example.rrf_num_testvecs,
+                    num_testvecs=Nin,
+                    # num_testvecs=20,
                     l2_err=epsilon_alpha,
-                    sampling_options={"scale": 0.1},
+                    # sampling_options={"scale": 0.1},
                 )
                 logger.info(f"\nSpectral Basis length: {len(rb)}.")
                 spectral_basis_sizes.append(len(rb))
-                snapshots.append(rb)
+                snapshots.append(rb) # type: ignore
         logger.info(
             f"Average length of spectral basis: {np.average(spectral_basis_sizes)}."
         )
@@ -132,14 +141,17 @@ def main(args):
         raise ValueError("Basis is not orthonormal wrt range product.")
 
     # Definition of test set (μ) and test data (g)
-    test_set = parameter_space.sample_randomly(50)
+    test_set = parameter_space.sample_randomly(ntrain)
     test_data = transfer.range.empty(reserve=len(test_set))
+
+    # use ntrain and Nin to define test data, but with different seed?
 
     logger.info(f"Computing test set of size {len(test_set)}...")
     with new_rng(example.projerr_seed):
         for mu in test_set:
             transfer.assemble_operator(mu)
-            g = transfer.generate_random_boundary_data(1, args.distr, {"scale": 0.1})
[…]

### 2024-08-28 src/parageom/run_locrom.py

@@ -52,7 +52,8 @@ def main(args):
     )
     trafo_d_gl = df.fem.Function(global_auxp.problem.V, name="d_trafo")
     fom = discretize_fom(example, global_auxp, trafo_d_gl)
-    h1_product = fom.products["h1_0_semi"]
+    # h1_product = fom.products["h1_0_semi"]
+    energy_product = fom.products["energy"]
 
     # ### Discretize subdomain operators
     # NOTE
@@ -130,8 +131,8 @@ def main(args):
     Nmax = max_dofs_per_vert.max()
     ΔN = 10
     num_modes_per_vertex = list(range(Nmax // ΔN, Nmax + 1, 3 * (Nmax // ΔN)))
-    num_modes_per_vertex = [20, 60, 100, 120]
-    # breakpoint()
+    # num_modes_per_vertex = [20, 40, 60, 80, 100, 120] #, 100, 120]
+    num_modes_per_vertex = num_modes_per_vertex[-1:]
     logger.debug(f"{Nmax=}")
     logger.debug(f"{num_modes_per_vertex=}")
 
@@ -166,16 +167,16 @@ def main(args):
                 logger.info(f"AssemblyEI took {t.elapsed()[0]}.")
             rom = StationaryModel(operator, rhs, name="locROM_with_ei")
 
-            cond_A = []
-            for mu in validation_set:
-                cond_A.append(compute_kappa(rom, mu))
-            breakpoint()
-            print("check condition number")
+            # cond_A = []
+            # for mu in validation_set:
+            #     cond_A.append(compute_kappa(rom, mu))
+            # κ = np.array(cond_A)
+            # logger.debug(f"{κ=}")
 
         fom_solutions = fom.solution_space.empty()
         rom_solutions = fom.solution_space.empty()
 
-        cond_A_without_ei = []
+        # cond_A_without_ei = []
 
         for mu in validation_set:
             U_fom = fom.solve(mu)
@@ -200,8 +201,8 @@ def main(args):
                     )
                     logger.info(f"{nmodes=}, \tAssembly took {t.elapsed()[0]}.")
                 rom = StationaryModel(operator, rhs, name="locROM")
-                kappa = compute_kappa(rom, mu)
-                cond_A_without_ei.append(kappa)
+                # kappa = compute_kappa(rom, mu)
+                # cond_A_without_ei.append(kappa)
 
                 with Timer("Solve") as t:
[…]

### 2024-08-28 src/parageom/stress_validation.py

@@ -4,11 +4,11 @@ import dolfinx as df
 import basix
 
 import numpy as np
-from pymor.core.pickle import load
 from pymor.vectorarrays.numpy import NumpyVectorSpace
 from pymor.operators.constructions import VectorOperator, LincombOperator, VectorFunctional, ConstantOperator
 from pymor.parameters.functionals import GenericParameterFunctional
 from pymor.models.basic import StationaryModel
+from pymor.tools.random import new_rng
 
 
 def main(args):
@@ -18,6 +18,8 @@ def main(args):
     from .fom import discretize_fom, ParaGeomLinEla
 
     # ### Build FOM
+    print("Discretizing FOM ...")
+
     coarse_grid_path = example.coarse_grid("global").as_posix()
     parent_domain_path = example.parent_domain("global").as_posix()
     interface_tags = [
@@ -39,6 +41,7 @@ def main(args):
     parageom = ParaGeomLinEla(auxiliary_problem.problem.domain, V, d_trafo, matparam) # type: ignore
 
     # ### Build localized ROM
+    print("Building ROM ...")
     rom, rec_data = build_localized_rom(args, example, auxiliary_problem, d_trafo, fom.parameters, ω=args.omega)
 
     # ### Global function for stress
@@ -52,18 +55,10 @@ def main(args):
     displacement = df.fem.Function(V)
 
     # ### Get optimal solution μ*
-    fom_data = example.fom_minimization_data
-    with fom_data.open("rb") as fh:
-        data = load(fh)
-    mu = fom.parameters.parse(data["mu_min"])
-
-    # targets
-    xdmf_files = example.pp_stress(args.name)
-
-    s1_fom, s2_fom, s_fom = compute_principal_stress(fom, mu, displacement, stress, parageom, rec_data=None, xdmf_filename=xdmf_files["fom"].as_posix())
-    displacement.x.array[:] = 0. # type: ignore
-    stress.x.array[:] = 0. # type: ignore
-    s1_rom, s2_rom, s_rom = compute_principal_stress(rom, mu, displacement, stress, parageom, rec_data=rec_data, xdmf_filename=xdmf_files["rom"].as_posix())
+    # ### ROM Assembly and Error Analysis
+    P = fom.parameters.space(example.mu_range)
+    with new_rng(example.validation_set_seed):
[…]

### 2024-08-28 src/parageom/tasks.py

@@ -220,6 +220,13 @@ def task_gfem():
     """ParaGeom: Build GFEM approximation"""
     module = "src.parageom.gfem"
     distr = example.distributions[0]
+
+    def create_action(module, nreal, method, distr, debug=False):
+        action = "python3 -m {} {} {} {}".format(module, nreal, method, distr)
+        if debug:
+            action += " --debug"
+        return action
+
     for nreal in range(example.num_real):
         for method in example.methods:
             deps = [SRC / "gfem.py"]
@@ -240,7 +247,7 @@ def task_gfem():
             yield {
                     "name": method + ":" + str(nreal),
                     "file_dep": deps,
-                    "actions": ["python3 -m {} {} {} {}".format(module, nreal, method, distr)],
+                    "actions": [create_action(module, nreal, method, distr, debug=True)],
                     "targets": targets,
                     "clean": True,
                     }
@@ -309,6 +316,7 @@ def task_optimization():
     module = "src.parageom.optimization"
     distr = example.distributions[0]
 
+    num_modes = 120
     minimizer = "SLSQP"
     omega = example.omega
 
@@ -326,7 +334,7 @@ def task_optimization():
         yield {
                 "name": method,
                 "file_dep": deps,
-                "actions": ["python3 -m {} {} {} --method {} --omega {}".format(module, distr, method, minimizer, omega)],
+                "actions": ["python3 -m {} {} {} --num_modes {} --method {} --omega {}".format(module, distr, method, num_modes, minimizer, omega)],
                 "targets": targets,
                 "clean": True,
                 }

### 2024-08-28 src/parageom/test_pull_back_global_domain.py

@@ -3,11 +3,9 @@ import numpy as np
 from mpi4py import MPI
 import dolfinx as df
 from dolfinx.io import gmshio
-from dolfinx.io.utils import XDMFFile
 import ufl
 import basix
 
-from multi.boundary import plane_at
 from multi.preprocessing import create_meshtags
 from multi.domain import RectangularDomain, StructuredQuadGrid
 from multi.materials import LinearElasticMaterial
@@ -22,11 +20,9 @@ from .definitions import BeamData
 
 def compute_reference_solution(example: BeamData, mshfile, degree, d):
     from .matrix_based_operator import BCTopo, _create_dirichlet_bcs
-    # translate parent domain
-    domain = gmshio.read_from_msh(
-        mshfile, MPI.COMM_WORLD, gdim=2
-    )[0]
 
+    # translate parent domain
+    domain = gmshio.read_from_msh(mshfile, MPI.COMM_WORLD, gdim=2)[0]
     x_domain = domain.geometry.x
     disp = np.pad(
         d.x.array.reshape(x_domain.shape[0], -1),  # type: ignore
@@ -34,11 +30,21 @@ def compute_reference_solution(example: BeamData, mshfile, degree, d):
     )
     x_domain += disp
 
-    # Physical Domain
-    top_locator = plane_at(example.height, "y")
+    # Coarse grid
+    grid, _, _ = read_mesh(
+        example.coarse_grid("global"),
+        MPI.COMM_SELF,
+        kwargs={"gdim": example.gdim},
+    )
+    coarse_grid = StructuredQuadGrid(grid)
+
+    # Definition of Neumann marker function
+    neumann = example.get_neumann(coarse_grid.grid, "left")
+    assert neumann is not None
+    top_marker, top_locator = neumann
+
     tdim = domain.topology.dim
     fdim = tdim - 1
-    top_marker = int(194)
     facet_tags, _ = create_meshtags(domain, fdim, {"top": (top_marker, top_locator)})
     omega = RectangularDomain(domain, facet_tags=facet_tags)
 
@@ -50,12 +56,6 @@ def compute_reference_solution(example: BeamData, mshfile, degree, d):
     problem = LinearElasticityProblem(omega, V, phases=material)
 
     # Dirichlet BCs
-    grid, _, _ = read_mesh(
-        example.coarse_grid("global"),
-        MPI.COMM_SELF,
-        kwargs={"gdim": example.gdim},
-    )
[…]

### 2024-08-28 paper/paper.tex
[review jörg] wording and commas

@@
 The global approximation is obtained by a suitable coupling of the local reduced spaces spanned by the aforementioned basis functions, in which one naturally relies on domain decomposition (DD) strategies.
-For a review of concepts in localized model order reduction it is referred to~\cite{BuhrReview}.
+A review of concepts in localized model order reduction is presented in~\cite{BuhrReview}.
 In particular, regarding lattice structures, this approach allows to take advantage of the repetitiveness of the lattice, such that a computation of the local basis is required only for few components, \ie{} unit cells.
@@
 The direct calculation via eigenvalue problems is, however, computationally expensive and the range of the transfer operator, and thus the optimal local approximation spaces can be efficiently approximated by random sampling~\cite{Buhr2018Randomized}.
-Here, the authors treat non-parametrized partial differential equations (PDEs) and to the authors knowledge the extension to the parametric setting for linear problems has not been done yet.
+Here, the authors treat non-parametrized partial differential equations (PDEs) and to our knowledge the extension to the parametric setting for linear problems has not been done yet.
 In~\cite{Smetana2016Optimal}, the authors propose a spectral greedy algorithm to construct parameter-independent local approximation spaces and Taddei and Patera~\cite{Taddei2018Localization} propose a combination of transfer eigenproblems and proper orthogonal decomposition (POD).
@@
 \Cref{sec:numerical experiments} comprises the numerical experiments.
-Based on the example of a graded concrete slab, the quality of the local spaces generated by the method described in~\cref{sec:Randomized range finder and POD} is analyzed, the proposed ROM is validated, and results for the solution of a shape optimization problem are presented.
+Based on the example of a graded concrete slab, the quality of the local spaces generated by the proposed method is analyzed, the resulting ROM is validated, and results for the solution of a shape optimization problem are presented.
 Finally, conclusions are given in~\cref{sec:conclusions}.
@@
 Third, it is outlined how a global approximation is obtained from local approximation spaces via the GFEM.
-Finally, the necessity and use of hyper-reduction (in the form of empirical interpolation) to ensure an online-efficient ROM is detailed in~\cref{sub:Online efficiency}.
+Finally, the hyper-reduction (in the form of empirical interpolation) to ensure an online-efficient ROM is detailed in~\cref{sub:Online efficiency}.
 
@@
 \end{align}
-The stiffness tetrad of the auxiliary problem is defined as
+The stiffness tetrad $\hat{\mathbb{C}}$ of the auxiliary problem is defined as
 \begin{equation}
@@
 However, for each new parameter value $\bm\mu$ for which the model is to be evaluated, we first need to solve the auxiliary problem.
-Nevertheless, this does not pose a problem, since the solution of the auxiliary problem is well amenable to acceleration via established MOR techniques and it is referred to~\cite{Guo2022Learning} for further details.
+Nevertheless, this does not pose a problem, since the solution of the auxiliary problem is well amenable to acceleration via established MOR techniques and we refer to~\cite{Guo2022Learning} for further details.
 
@@
 maps the boundary function $\bm{g}$ to the solution $\bm{u}(\bm\mu)$ in the target subdomain $\varOmega^{\m}_{\mathrm{in}}$.
-The approximation of the range of this transfer operator is then the sought after local approximation space, see~\cite{Buhr2018Randomized}.
+The approximation of the range of this transfer operator is then the local approximation space, see~\cite{Buhr2018Randomized}.
 
@@
 \end{equation}
-The resolution of~\cref{eq:weak_osp_parent} and subsequent restriction of the solution to the target subdomain, denoted as $\bm{u}(\m)\big\vert_{\varOmega^{\p}_{\inrm}}$, comprises the solution of the oversampling problem.
-The transfer operator $\bm{T}_{\m}: \mathcal{S}\to\mathcal{R}$, maps for a given parameter value $\m$ the boundary data $\bm{g}$ to the solution on the target subdomain $\bm{u}(\m)\big\vert_{\varOmega^{\p}_{\inrm}} = \bm{T}_{\m}(\bm{g}).$
+The solution of the oversampling problem~\cref{eq:osp_p} is then given as the solution of~\cref{eq:weak_osp_parent} and subsequent restriction of the solution to the target subdomain, denoted as $\bm{u}(\m)\big\vert_{\varOmega^{\p}_{\inrm}}$.
+The transfer operator $\bm{T}_{\m}: \mathcal{S}\to\mathcal{R}$ maps, for a given parameter value $\m$, the boundary data $\bm{g}$ to the solution on the target subdomain $\bm{u}(\m)\big\vert_{\varOmega^{\p}_{\inrm}} = \bm{T}_{\m}(\bm{g}).$
 Here, $\mathcal{S}$ and $\mathcal{R}$ are appropriate source and range spaces, respectively.
 Throughout the paper, we assume a suitable discretization and numerical solution of~\cref{eq:weak_osp_parent} using the FE method.
-For details concerning the discretization of the transfer operator (for a fixed parameter value) it is referred to~\cite{Buhr2018Randomized}.
+For details concerning the discretization of the transfer operator (for a fixed parameter value), we refer to~\cite{Buhr2018Randomized}.
 
@@
 Moreover, (global) \textsc{Neumann} or \textsc{Dirichlet} boundary conditions prescribed on $\varSigma^{\m}_{\mathrm{N}}$ or $\varSigma^{\m}_{\mathrm{D}}$, respectively, have not been considered yet.
-Therefore, for example, an additional \textsc{Neumann} problem\footnote{The case of inhomogeneous \textsc{Dirichlet} boundary conditions can be treated in the same way.} has to be solved to enrich the local spaces.
+Therefore, an additional \textsc{Neumann} problem\footnote{The case of inhomogeneous \textsc{Dirichlet} boundary conditions can be treated in the same way.} has to be solved to enrich the local spaces.
 \begin{align}
@@
 
-Assuming a proper FE discretization of $\bm{T}_{\m}$, next the algorithm to approximate the range of $\bm{T}_{\m}$ for any $\m$ is discussed.
+Assuming a proper FE discretization of $\bm{T}_{\m}$, the algorithm to approximate the range of $\bm{T}_{\m}$ for any $\m$ is discussed next.
 Herein, a strategy for the exploration of the parameter space and sampling of the random boundary conditions in the oversampling are the key challenges.
 In the present work, the strategy by Taddei \& Patera~\cite{Taddei2018Localization} is adopted in which based on the connection between the POD of images of the transfer operator (for certain boundary conditions) and the singular values of the transfer operator they devise a training strategy which is recognized as a distributed approximate POD (see \cite{Himpe2018Hierarchical}).
-For each $\m_j$ in a predefined training set, they compute the singular values of $\bm{T}_{\m_j}$ by resolution of an eigenvalue problem and subsequently compute the POD over the set of all eigenvectors.
-As the authors state, the resolution of the eigenvalue problems may be prohibitive in some cases, and therefore we employ range approximation by random sampling instead.
+For each $\m_j$ in a predefined training set, they compute the singular values of $\bm{T}_{\m_j}$ by solution of an eigenvalue problem and, subsequently, compute the POD over the set of all eigenvectors.
+As the authors state, the resolution of the eigenvalue problems may be prohibitive in some cases, and therefore we employ range approximation by random sampling.
 
@@
 The training set $\bm{S}_{\mathrm{train}}$ has to be sufficiently rich, such that it is a good representation of the parameter space, and the reader is referred to~\cite{Haasdonk2017Chapter}.
-Next, for each parameter value $\m_j\in\bm{S}_{\mathrm{train}}$ the range of $\bm{T}_{\m_j}$ which is now fixed with respect to the parameter value is approximated using Algorithm 1 \textsc{AdaptiveRandomizedRangeApproximation} from~\cite{Buhr2018Randomized} and the set of snapshots is extended by Gram Schmidt orthogonalization (line 5 \& 6).
+Next, for each parameter value $\m_j\in\bm{S}_{\mathrm{train}}$, the range of $\bm{T}_{\m_j}$ which is now fixed with respect to the parameter value is approximated using Algorithm 1 \textsc{AdaptiveRandomizedRangeApproximation} from~\cite{Buhr2018Randomized} and the set of snapshots is extended by Gram Schmidt orthogonalization (line 5 \& 6).
 Note that the parameter value $\m$ and boundary conditions $\bm{g}$ are not varied at the same time.
@@
 The difficulty is that the basis functions of the local spaces $X^n_i$ and $X^n_j$ of two neighboring subdomains $\varOmega_i^{\p}$ and $\varOmega_j^{\p}$ are not conforming on the shared interface $\varGamma^{\p}_{ij}=\partial\varOmega_i^{\p}\cap\partial\varOmega_j^{\p}$.
-However, given a suitably\footnote{According to Definition 1 of~\cite{BABUSKA1997}.} defined partition of unity $\{{\varphi}_i\}$ the local basis functions can be included in the global FE space by multiplication with the standard FE shape functions ${\varphi}_i$, such that a conforming approximation is obtained.
+However, given a suitably\footnote{According to Definition 1 of~\cite{BABUSKA1997}.} defined partition of unity $\{{\varphi}_i\}$, the local basis functions can be included in the global FE space by multiplication with the standard FE shape functions ${\varphi}_i$, such that a conforming approximation is obtained.
 The global GFEM space is defined as
@@
 
-\subsection{Online efficiency} % (fold)
+\subsection{Global solution procedure} % (fold)
 \label{sub:Online efficiency}
@@
 By reversing the $\vecm(\cdot)$ operation, the approximation $\bm{K}_Q(\m)$ can be formed using~\cref{eq:K_ei} once the interpolation coefficients are determined by solution of \cref{eq:interp}.
-For more details on the overall procedure it is referred to~\cite{Negri2015Efficient}.
+For more details on the overall procedure the reader is referred to~\cite{Negri2015Efficient}.
 

### 2024-08-28 paper/paper.tex
add gradientenbeton image

@@
 
+\urldef{\gradientenbetonurl}\url{https://gradientenbeton.eu/}
 \graphicspath{{./img/}}
@@
 \label{sec:numerical experiments}
-
-\begin{itemize}
-    \item general stuff like FE cell type and degree
-    \item software: dolfinx version 0.8.0
-    \item PC hardware?
-\end{itemize}
+The proposed methodology to construct the ROM is evaluated based on the example of a graded concrete slab.
+In~\cref{sub:Projection error study}, the approximation quality of the local spaces is accessed by studying the projection error.
+The ROM is validated against the FOM (\cref{sub:Reduced order model validation}) and the solution of a shape optimization problem is presented in~\cref{sub:Shape optimization}.
+\comment[id=pd]{TODO: remove these sentences in case the article is too long.}
 
@@
 \label{sub:Graded concrete slab}
+The numerical example investigated here is motivated by \textit{graded concrete} in which voids are created by inserting hollow bodies into the concrete formwork with the goal of mass reduction, see~\cref{fig:gradientenbeton}.
+The problem is simplified and we consider a two dimensional beam under plane stress conditions as shown in~\cref{fig:beam_sketch}.
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[keepaspectratio,height=4cm]{gradientenbeton_formwork.jpg}
+    \end{center}
+    \caption{Formwork for a graded concrete slab.
+Weight savings are achieved by placing hollow spheres along a structured grid inside the formwork.
+Source:\gradientenbetonurl.}\label{fig:gradientenbeton}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=0.95\textwidth]{example-image}
+    \end{center}
+    \caption{}\label{fig:beam_sketch}
+\end{figure}
 

### 2024-08-30 src/parageom/definitions.py
rom validation and optimization

@@ -78,8 +78,8 @@ class BeamData:
     methods: tuple[str, ...] = ("hapod", ) # "heuristic")
     epsilon_star: dict = field(
             default_factory=lambda: {
-                "heuristic": 0.01,
-                "hapod": 0.01,
+                "heuristic": 0.001,
+                "hapod": 0.001,
                 })
     epsilon_star_projerr: float = 0.001
     omega: float = 0.5 # ω related to HAPOD (not output functional)
@@ -221,16 +221,12 @@ class BeamData:
 
     def ntrain(self, k: int) -> int:
         """Define size of training set for k-th transfer problem"""
-
-        # FIXME
-        # put meaningful numbers
-        # this is just to test if the conditioning improves with new oversampling
         if k in (0, 10):
-            return 30
-        elif k in (1, 9):
             return 50
-        else:
+        elif k in (1, 9):
             return 100
+        else:
+            return 200
 
     def cell_to_config(self, cell: int) -> str:
         """Maps global cell index to config."""
@@ -257,6 +253,19 @@ class BeamData:
         else:
             return dir / f"locrom_error_{distr}.npz"
 
+    def rom_error_u(self, nreal: int, num_modes: int, method="hapod", ei=False) -> Path:
+        dir = self.method_folder(nreal, method)
+        if ei:
+            return dir / f"rom_error_u_ei_{num_modes}.npz"
+        else:
+            return dir / f"rom_error_u_{num_modes}.npz"
+
+    def rom_error_s(self, nreal: int, num_modes: int, method="hapod", ei=False) -> Path:
+        dir = self.method_folder(nreal, method)
+        if ei:
+            return dir / f"rom_error_s_ei_{num_modes}.npz"
+        else:
+            return dir / f"rom_error_s_{num_modes}.npz"
     @property
     def fom_minimization_data(self) -> Path:
         """FOM minimization data"""
@@ -327,6 +336,17 @@ class BeamData:
             return dir / f"run_locrom_ei_{distr}.log"
         else:
             return dir / f"run_locrom_{distr}.log"
+
+    def log_validate_rom(self, nr: int, modes: int, method="hapod", distr="normal", ei=True) -> Path:
+        dir = self.logs_path(nr, method)
+        if ei:
+            return dir / f"validate_rom_{modes}_{distr}_with_ei.log"
+        else:
+            return dir / f"validate_rom_{modes}_{distr}.log"
+
+    @property
+    def log_optimization(self) -> Path:
[…]

### 2024-08-30 src/parageom/fom.py

@@ -13,8 +13,10 @@ from multi.preprocessing import create_meshtags
 from multi.product import InnerProduct
 from multi.io import read_mesh
 
-from pymor.basic import VectorOperator, StationaryModel, GenericParameterFunctional, NumpyVectorSpace, VectorFunctional
-from pymor.operators.constructions import ConstantOperator, LincombOperator
+from pymor.parameters.functionals import GenericParameterFunctional
+from pymor.operators.constructions import ConstantOperator, LincombOperator, VectorOperator, VectorFunctional
+from pymor.vectorarrays.numpy import NumpyVectorSpace
+from pymor.models.basic import StationaryModel
 from pymor.bindings.fenicsx import (
     FenicsxVectorSpace,
     FenicsxVisualizer,
@@ -138,13 +140,14 @@ def discretize_subdomain_operators(example):
     from .auxiliary_problem import discretize_auxiliary_problem
     from .matrix_based_operator import FenicsxMatrixBasedOperator
 
+    # discretize auxiliary problem on unit cell domain
     parent_subdomain_msh = example.parent_unit_cell.as_posix()
     ftags = {"bottom": 11, "left": 12, "right": 13, "top": 14, "interface": 15}
-    aux = discretize_auxiliary_problem(
-        example, parent_subdomain_msh, ftags, example.parameters["subdomain"]
-    )
-    d = df.fem.Function(aux.problem.V, name="d_trafo")
+    params = example.parameters["subdomain"]
+    aux = discretize_auxiliary_problem(example, parent_subdomain_msh, ftags, params)
+    d = df.fem.Function(aux.problem.V, name="d_trafo_unit_cell")
 
+    # create problem to define (stiffness matrix) operator
     omega = aux.problem.domain
     matparam = {"gdim": omega.gdim, "E": example.youngs_modulus, "NU": example.poisson_ratio, "plane_stress": example.plane_stress}
     problem = ParaGeomLinEla(omega, aux.problem.V, d, matparam)
@@ -155,14 +158,11 @@ def discretize_subdomain_operators(example):
         aux.solve(d, mu)
         d.x.scatter_forward()
 
-    params = example.parameters["subdomain"]
     operator = FenicsxMatrixBasedOperator(
         problem.form_lhs, params, param_setter=param_setter, name="ParaGeom"
     )
 
     # ### wrap external force as pymor operator
-    # EMOD = example.youngs_modulus
[…]

### 2024-08-30 src/parageom/optimization.py

@@ -5,47 +5,143 @@
 # "Model order reduction for PDE-constrained optimization problems"
 # See https://docs.pymor.org/2023-2-0/tutorial_optimization.html
 
+import pathlib
+from collections import namedtuple
+
+from mpi4py import MPI
 import dolfinx as df
 import basix
+import ufl
 
 import numpy as np
-from pymor.vectorarrays.numpy import NumpyVectorSpace
-from pymor.operators.constructions import VectorOperator, LincombOperator, VectorFunctional, ConstantOperator
-from pymor.parameters.functionals import GenericParameterFunctional
-from pymor.models.basic import StationaryModel
+from pymor.core.defaults import set_defaults
+from pymor.core.logger import getLogger
 from pymor.core.pickle import dump
+from multi.io import read_mesh
+from multi.domain import StructuredQuadGrid
+from scipy.optimize import minimize, NonlinearConstraint, Bounds
+
+StressWrapper = namedtuple(
+        "StressWrapper", ["expr", "cells", "fun"]
+        )
+ModelWrapper = namedtuple(
+        "ModelWrapper", ["model", "solution", "dofmap", "modes", "sol_local"],
+        defaults=(None, ) * 3
+        )
 
 
 def main(args):
     """solve optimization problem for different models"""
     from .tasks import example
-    from .auxiliary_problem import discretize_auxiliary_problem
-    from .fom import discretize_fom, ParaGeomLinEla
+    from .validate_rom import build_fom, build_rom
+    from .dofmap_gfem import GFEMDofMap
+    from .locmor import reconstruct
 
-    # TODO add logger
+    stem = pathlib.Path(__file__).stem
+    # always use first realization for optimization, nreal=0
+    logfilename = example.log_optimization.as_posix()
+    set_defaults({"pymor.core.logger.getLogger.filename": logfilename})
+    if args.debug:
+        loglevel = 10 # debug
+    else:
+        loglevel = 20 # info
+    logger = getLogger(stem, level=loglevel)
 
     # ### Build FOM
-    coarse_grid_path = example.coarse_grid("global").as_posix()
-    parent_domain_path = example.parent_domain("global").as_posix()
-    interface_tags = [
-        i for i in range(15, 25)
-    ]
-    auxiliary_problem = discretize_auxiliary_problem(
-        example,
[…]

### 2024-08-30 src/parageom/osp_v2.py

@@ -1,6 +1,5 @@
 """oversampling version 2"""
 
-import pathlib
 import typing
 from dataclasses import dataclass
 import dolfinx as df

### 2024-08-30 src/parageom/plot_romerr.py

@@ -1,76 +1,61 @@
 import numpy as np
+from collections import defaultdict
 from multi.postprocessing import read_bam_colors
 from multi.plotting_context import PlottingContext
-
-
-def postproc(nreal: int, method: str, distr: str):
-    from .tasks import example
-
-    instream = example.locrom_error(nreal, method, distr)
-    data = np.load(instream)
-    max_dpv = np.load(example.local_basis_dofs_per_vert(nreal, method, distr))
-
-    # now check values in data
-    # and max dofs per vert
-
-    breakpoint()
 
 
 def main(cli):
     from .tasks import example
 
-    distr = example.distributions[0]
     bamcd = read_bam_colors()
     blue = bamcd["blue"][0]
     red = bamcd["red"][0]
 
-    marker = {
-        "l2_err": "o",
-        "max_relerr_h1_semi": "s",
-        "max_relerr_max": "+",
-    }
-    # marker_ei = {
-    #     "l2_err": "o",
-    #     "max_relerr_h1_semi": "o",
-    #     "max_relerr_max": "o",
-    # }
-    norm_label = {"max_relerr_h1_semi": "H1-semi", "max_relerr_max": "max"}
+    nreal = cli.nreal
+    number_of_modes = [20, 40, 60, 80, 100]
+
+    erru = defaultdict(list)
+    errs = defaultdict(list)
+
+    # define data to gather via keys
+    keys = ["relerr"]
+
+    # append 1.0 for num_modes=0
+    for key in keys:
+        erru["min_"+key].append(1.0)
+        erru["max_"+key].append(1.0)
+        erru["avg_"+key].append(1.0)
+        errs["min_"+key].append(1.0)
+        errs["max_"+key].append(1.0)
+        errs["avg_"+key].append(1.0)
+
+    for num_modes in number_of_modes:
+        data_u = np.load(example.rom_error_u(nreal, num_modes, ei=cli.ei))
+        data_s = np.load(example.rom_error_s(nreal, num_modes, ei=cli.ei))
+
+        for key in keys:
+            erru["min_"+key].append(np.min(data_u[key]))
+            erru["max_"+key].append(np.max(data_u[key]))
+            erru["avg_"+key].append(np.average(data_u[key]))
+
+            errs["min_"+key].append(np.min(data_s[key]))
+            errs["max_"+key].append(np.max(data_s[key]))
+            errs["avg_"+key].append(np.average(data_s[key]))
+
[…]

### 2024-08-30 src/parageom/tasks.py

@@ -65,7 +65,7 @@ def task_coarse_grid():
     """ParaGeom: Create structured coarse grids"""
     module = "src.parageom.preprocessing"
 
-    for config in list(CONFIGS) + ["global", "target"]:
+    for config in ["global"]:
         yield {
             "name": config,
             "file_dep": [],
@@ -83,7 +83,7 @@ def task_coarse_grid():
 def task_fine_grid():
     """ParaGeom: Create parent domain mesh"""
     module = "src.parageom.preprocessing"
-    for config in list(CONFIGS) + ["global", "target"]:
+    for config in ["global"]:
         yield {
             "name": config,
             "file_dep": [example.coarse_grid(config), example.parent_unit_cell],
@@ -105,41 +105,6 @@ def task_preproc():
     }
 
 
-# def task_hapod():
-#     """ParaGeom: Construct basis via HAPOD"""
-#     module = "src.parageom.hapod"
-#     nworkers = 4  # number of workers in pool
-#     distr = example.distributions[0]
-#     for nreal in range(example.num_real):
-#         for config in CONFIGS:
-#             deps = [SRC / "hapod.py"]
-#             deps.append(example.coarse_grid("global"))
-#             deps.append(example.parent_domain("global"))
-#             deps.append(example.coarse_grid("target"))
-#             deps.append(example.parent_domain("target"))
-#             deps.append(example.coarse_grid(config))
-#             deps.append(example.parent_domain(config))
-#             targets = []
-#             targets.append(
-#                 example.log_basis_construction(nreal, "hapod", distr, config)
-#             )
-#             targets.append(example.hapod_modes_npy(nreal, distr, config))
-#             targets.append(example.hapod_singular_values(nreal, distr, config))
-#             if config == "left":
-#                 targets.append(example.hapod_neumann_svals(nreal, distr, config))
-#             yield {
-#                 "name": config + ":" + str(nreal),
-#                 "file_dep": deps,
-#                 "actions": [
-#                     "python3 -m {} {} {} {} --max_workers {}".format(
-#                         module, distr, config, nreal, nworkers
-#                     )
-#                 ],
-#                 "targets": targets,
-#                 "clean": True,
-#             }
-
-
 def task_hapod():
     """ParaGeom: Construct basis via HAPOD"""
     module = "src.parageom.osp_v2"
@@ -148,20 +113,16 @@ def task_hapod():
         for k in range(11):
             deps = [SRC / "osp_v2.py"]
             deps.append(example.coarse_grid("global"))
[…]

### 2024-08-30 src/parageom/validate_rom.py

@@ -1,23 +1,36 @@
+import pathlib
+from time import perf_counter
+
 from mpi4py import MPI
 import ufl
 import basix
 import dolfinx as df
 import numpy as np
-from pymor.tools.random import new_rng
-from pymor.models.basic import StationaryModel
 from multi.io import read_mesh
 from multi.domain import StructuredQuadGrid
 
+from pymor.core.defaults import set_defaults
+from pymor.core.logger import getLogger
+from pymor.tools.random import new_rng
+from pymor.models.basic import StationaryModel
+from pymor.parameters.functionals import GenericParameterFunctional
+from pymor.vectorarrays.numpy import NumpyVectorSpace
+from pymor.operators.constructions import ConstantOperator, LincombOperator, VectorFunctional
 
-def main():
-    """Eval ROM for same parameter value twice.
-    Call reconstruct in-between.
-    See if solutions are equal.
-    """
+
+def main(args):
     from .tasks import example
     from .dofmap_gfem import GFEMDofMap
     from .locmor import reconstruct, assemble_gfem_system
-    # from .stress_analysis import project
+
+    stem = pathlib.Path(__file__).stem
+    logfilename = example.log_validate_rom(args.nreal, args.num_modes, ei=args.ei).as_posix()
+    set_defaults({"pymor.core.logger.getLogger.filename": logfilename})
+    if args.debug:
+        loglevel = 10 # debug
+    else:
+        loglevel = 20 # info
+    logger = getLogger(stem, level=loglevel)
 
     # ### FOM
     fom, parageom_fom = build_fom(example)
@@ -26,7 +39,6 @@ def main():
     # ### Global function for displacment
     d_rom = df.fem.Function(V, name="urom")
     d_fom = df.fem.Function(V, name="ufom")
-    other_d_rom = df.fem.Function(V)
 
     # ### Quadrature space for stress
     basix_celltype = getattr(basix.CellType, V.mesh.topology.cell_type.name)
@@ -53,8 +65,6 @@ def main():
     cells = np.arange(0, num_cells, dtype=np.int32)
 
     def compute_principal_components(f):
-        # FIXME
-        # how to avoid hardcoding reshape?
         values = f.reshape(cells.size, 4, 4)
         fxx = values[:, :, 0]
         fyy = values[:, :, 1]
@@ -63,17 +73,19 @@ def main():
[…]

### 2024-08-30 paper/paper.tex
progress writing

@@
 \usepackage{algorithm, algorithmicx, algpseudocode}
-
-\usepackage{tikz}
-\usepackage{standalone}
-\usepackage{booktabs}
-\usepackage{pgfplotstable}
-\pgfplotsset{compat=1.18}
-\usetikzlibrary{%
-	calc,
-	patterns,
-	positioning,
-	decorations.pathreplacing,%
-	decorations.pathmorphing,%
-	arrows.meta,%
-}
-\usepackage{bamcolors}
 
@@
 
+\usepackage{tikz}
+\usepackage{standalone}
+\usepackage{booktabs}
+\usepackage{pgfplotstable}
+\pgfplotsset{compat=1.18}
+\usetikzlibrary{%
+	calc,
+	patterns,
+	positioning,
+	decorations.pathreplacing,%
+	decorations.pathmorphing,%
+	arrows.meta,%
+}
+\usepackage{bamcolors}
+
 \definechangesauthor[color=blue, name={Philipp Diercks}]{pd}
@@
 \newcommand{\inrm}{\mathrm{in}}
+\newcommand{\fom}{\mathrm{fom}}
+\newcommand{\rom}{\mathrm{rom}}
+\newcommand{\train}{\mathrm{train}}
 % operators
@@
 \graphicspath{{./img/}}
+
+% \usepackage{stanli}
 
@@
 
-\added[id=pd]{TODO: add note on notation where appropriate.}
+\added[id=pd]{TODO: add note on notation where appropriate.}\\
+\added[id=pd]{TODO: shorten introduction.}
 
@@
 The training set $\bm{S}_{\mathrm{train}}$ has to be sufficiently rich, such that it is a good representation of the parameter space, and the reader is referred to~\cite{Haasdonk2017Chapter}.
-Next, for each parameter value $\m_j\in\bm{S}_{\mathrm{train}}$, the range of $\bm{T}_{\m_j}$ which is now fixed with respect to the parameter value is approximated using Algorithm 1 \textsc{AdaptiveRandomizedRangeApproximation} from~\cite{Buhr2018Randomized} and the set of snapshots is extended by Gram Schmidt orthogonalization (line 5 \& 6).
+Next, for each parameter value $\m_j\in\bm{S}_{\mathrm{train}}$, the range of $\bm{T}_{\m_j}$ which is now fixed with respect to the parameter value is approximated using Algorithm 1 \textsc{AdaptiveRandomizedRangeApproximation} from~\cite{Buhr2018Randomized} (line 6).
+If $\partial\varOmega^{\p}\cap\varSigma^{\p}_{\mathrm{N}}\ne\emptyset$, the additional \textsc{Neumann} problem~\cref{eq:neumann_problem} is solved and the solution appended to the current range $\bm{R}$ (lines 7-9).
+The set of snapshots is then extended by the current range $\bm{R}$ (line 11).
 Note that the parameter value $\m$ and boundary conditions $\bm{g}$ are not varied at the same time.
@@
 The dimension of the local space $X^n$, \ie{} the number of local basis functions is denoted by $n$.
-
-In case of inhomogeneous \textsc{Neumann} boundary conditions, after termination of~\cref{algo:rrfpod}, the \textsc{Neumann} problem~\cref{eq:neumann_problem} is solved for each $\m_j\in\bm{S}_{\mathrm{train}}$ as well.
-Following compression via POD the space $X^n$ is enriched by the additional \textsc{Neumann} modes.
 
@@
     % \State{$\bm{R}\gets \bm{R}\cup\left(\bm{u}_{\mathrm{Neumann}}(\bm\mu_j)\right)$}
-    \State{$\bm{S}\gets \mathrm{orthonormalize}\left(\bm{S}\cup\bm{R}\right)$}
+    \If{$\partial\varOmega^{\p}\cap\varSigma^{\p}_{\mathrm{N}}\ne\emptyset$}
+        \State{$\bm{u}_{\textsc{Neumann}}\gets$ Solve \cref{eq:neumann_problem} for $\m_j$ }
+        \State{$\bm{R} \gets \bm{R}\cup\bm{u}_{\mathrm{Neumann}}$}
+    \EndIf
+    \State{$\bm{S}\gets \bm{S}\cup\bm{R}$}
     \EndFor%
@@
 \end{algorithm}
-
-% \comment[id=pd]{TODO: explain tolerances to control approximation quality.
-Refer to HAPOD paper and mention that $\varepsilon^{\ast}$ can be used to compute tolerances at each node of the tree (epsilon star drives the RRF and the final POD.
-However, this means that we have to use l2-mean error as termination criterion.
-Too detailed?}
-% \comment[id=pd]{TODO: define notation for POD.}
 
@@
           \label{fig:gfem_left}%
-       } 
+       }\hfill 
        \subfloat[Enrichment function $\bm{\xi}(\bm{x})$]{%
@@
           \label{fig:gfem_middle}%
-       }
+       }\hfill
        \subfloat[GFEM function $\bm\psi(\bm{x})$]{%
@@
 \label{sec:numerical experiments}
-The proposed methodology to construct the ROM is evaluated based on the example of a graded concrete slab.
+% The proposed methodology to construct the ROM is evaluated based on the example of a graded concrete slab.
 In~\cref{sub:Projection error study}, the approximation quality of the local spaces is accessed by studying the projection error.
 The ROM is validated against the FOM (\cref{sub:Reduced order model validation}) and the solution of a shape optimization problem is presented in~\cref{sub:Shape optimization}.
-\comment[id=pd]{TODO: remove these sentences in case the article is too long.}
+% \comment[id=pd]{TODO: remove these sentences in case the article is too long.}
 
@@
 The problem is simplified and we consider a two dimensional beam under plane stress conditions as shown in~\cref{fig:beam_sketch}.
+The coarse grid partition of half of the beam body $\varOmega^{\p}_{\gl}=[0, 10a]\times[0, a]$ is given by $10$ quadrilateral cells of size $a$, where $a$ is the unit length of the unit cell.
+Each unit cell, \ie{} subdomain $\varOmega^{\p}_i$, is taken as a square of size $a$ with a circular void with radius $R$ and is partitioned in \added{384} isoparametric quadrilateral FEs of polynomial degree two, see~\cref{fig:oversampling_domain}\,(b).
+With regard to the geometrical parametrization, we consider the variation of the radius $R$ in each subdomain.
+For the global problem this results in a parameter space $\mathcal{P}_{\gl}=[R_{-}, R_{+}]^{10}$, such that a parameter value $\m\in\mathbb{P}_{\gl}$ has parameter components $\m_i\in[R_{-},R_{+}], i=1, \ldots, 10$, with ranges chosen as $R_{-}=0.1a, R_{+}=0.3a$.
+The point load is distributed as a constant traction over the length of one unit cell (on the far left).
+
+\paragraph{Full order model}
+An auxiliary problem is defined on the global domain $\varOmega^{\p}_{\gl}$ to account for the variation in the geometry.
+The full order model is given by a standard FE model using the fine grid discretization of $\varOmega^{\p}_{\gl}$ as described above.
+
+\paragraph{Reduced order model}
+The ROM is constructed using the procedure for range approximation described in~\cref{sub:Construction of local approximation spaces}.
+Using the coarse grid as above, we choose to solve $k=1, \ldots, 11$ oversampling problems, such that for each of the target subdomains $\varOmega^{\p, k}_{\inrm}$ and $\varOmega^{\p, k+1}_{\inrm}$ an overlap of size $a\times a$ is created.
+The pointwise overlap condition is required for the construction of the GFEM functions.
+The size of the target subdomain $\varOmega^{\p, k}_{\inrm}$ is $2a \times a$ for $k=2, \ldots, 10$ and $a \times a$ for $k = 1, 11$.
+For each problem the oversampling layer has size $a \times a$, \ie{} the target subdomain is extended by a unit cell on each side unless the boundary of the target subdomain intersects with the boundary of the global domain ($\partial\varOmega^{\p,k}_{\inrm}\cap\partial\varOmega^{\p}_{\gl}\ne\emptyset$).
+
+The input parameters to~\cref{algo:rrfpod} are $\varepsilon^{\ast}=0.001\,, n_t=\dim(\mathcal{S})$ and $\varepsilon_{\mathrm{algofail}}=1\cdot\,10^{-14}$ for $k=1, \ldots, 11$, where $\dim(\mathcal{S})$ denotes the dimension of the source space of the transfer operator and corresponds to the number of degrees of freedom on the boundary $\varGamma^{\p}_{\out}$ in the FE model.
+The number of training samples is chosen as $N_{\train}=50$ ($k=1, 11$), $N_{\train}=100$ ($k=2, 10$) and $N_{\train}=200$ elsewhere, as the local parameter spaces have different dimensions.
+
+\begin{itemize}
+    \item \added{number of snapshots for each POD}
+    \item difference to full order model: only use operators on unit cell domain in the online phase
+\end{itemize}
 
@@
 
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=0.95\textwidth]{example-image}
-    \end{center}
-    \caption{}\label{fig:beam_sketch}
+\begin{figure}[h]
+    \centering
+    \subfloat[Mechanical model of the beam]{%
+          \includegraphics[keepaspectratio]{traeger.pdf}%
+          \label{fig:beam_sketch_left}%
+       }\hfill 
+       \subfloat[Simulation model of half of the beam]{%
+          \includegraphics[keepaspectratio, scale=0.8]{halfbeam.pdf}%
+          \label{fig:beam_sketch_right}%
+       }
+       \caption{Mechanical model of the beam and representation of the simulation model with loads and kinematical constraints.}
+       \label{fig:beam_sketch}
 \end{figure}
 
-\begin{itemize}
-    \item introduce graded concrete / problem (gradientenbeton eu); why this is interesting
-    \item sketch of the mechanical problem
-    \item setup regarding the new method?
-\end{itemize}
-
-% \begin{figure}
-%     \begin{center}
-%         \includegraphics[width=0.95\textwidth]{./figures/beam/beam_sketch.pdf}
-%     \end{center}
-%     \caption{Schematic representation of the structural problem. A simply supported beam is subjected to a constant load.}\label{fig:beam_sketch}
-% \end{figure}
-%
-% TODO placement
-% \begin{figure}
-%     \begin{center}
-%         \includegraphics[width=0.65\textwidth]{./figures/beam/unit_cell.png}
-%     \end{center}
-%     \caption{Unit cell domain partition.}\label{fig:unit_cell_domain}
-% \end{figure}
-%
-% \begin{figure}
-%     \begin{center}
-%         \includegraphics[width=0.95\textwidth]{./figures/beam/global_domain.png}
-%     \end{center}
-%     \caption{The partition of the global domain. The colors indicate the variation in Young's modulus for each subdomain.}\label{fig:global_domain}
-% \end{figure}
-%
 % subsection Graded concrete slab (end)
-
-\subsection{Projection error study} % (fold)
-\label{sub:Projection error study}
-
-% subsection Projection error study (end)
 
@@
 \label{sub:Reduced order model validation}
+
+For the validation of the ROM, the relative error with respect to the FOM is computed over a parameter set of size $200$.
+The error in the displacements in measured in the energy norm (over the domain $\varOmega^{\p}_{\gl}$) and defined as
+\begin{equation}
+    e_u(\m) := \frac{\norm{\bm{u}_{\fom} - \bm{u}_{\rom}}}{\norm{\bm{u}_{\fom}}}\,.
+    \label{eq:error_u}
+\end{equation}
+The error in the first principal component of the \textsc{Cauchy} stress is defined as
+\begin{equation}
+    e_{\sigma}(\m) = \frac{\vert\vb{s}_{\fom} - \vb{s}_{\rom}\vert}{\vert \mathrm{max}(\vb{s}_{\fom})\vert}\,.
+    \label{eq:error_sigma}
+\end{equation}
+Here, $\vb{s}\in\mathbb{R}^{N_g}$ denotes the vector holding the first principal stress values at the $N_g$ integration points of the mesh.
+For ease of notation, the parameter-dependence is omitted for $\bm{u}$ and $\vb{s}$.
+The minimum, average and maximum of $e_u$ and $e_{\sigma}$ over the validation set are shown in~\cref{fig:relerr}.
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=0.65\textwidth]{relerr.pdf}
+    \end{center}
+    \caption{}\label{fig:relerr}
+\end{figure}
 
@@
 \label{sub:Shape optimization}
+
+\begin{itemize}
+    \item problem definiot
+    \item output computation (difference with ROM)
+    \item result for both models
+\end{itemize}
 

### 2024-08-30 paper/paper.tex
add relerr and singular values

@@
 \label{sub:Graded concrete slab}
-The numerical example investigated here is motivated by \textit{graded concrete} in which voids are created by inserting hollow bodies into the concrete formwork with the goal of mass reduction, see~\cref{fig:gradientenbeton}.
+The numerical example investigated here is motivated by \textit{graded concrete} in which voids are created by inserting hollow bodies into the concrete formwork with the goal of mass reduction, see~\gradientenbetonurl.
 The problem is simplified and we consider a two dimensional beam under plane stress conditions as shown in~\cref{fig:beam_sketch}.
@@
 With regard to the geometrical parametrization, we consider the variation of the radius $R$ in each subdomain.
-For the global problem this results in a parameter space $\mathcal{P}_{\gl}=[R_{-}, R_{+}]^{10}$, such that a parameter value $\m\in\mathbb{P}_{\gl}$ has parameter components $\m_i\in[R_{-},R_{+}], i=1, \ldots, 10$, with ranges chosen as $R_{-}=0.1a, R_{+}=0.3a$.
+For the global problem this results in a parameter space $\mathcal{P}_{\gl}=[R_{-}, R_{+}]^{10}$, such that a parameter value $\m\in\mathcal{P}_{\gl}$ has parameter components $\m_i\in[R_{-},R_{+}], i=1, \ldots, 10$, with ranges chosen as $R_{-}=0.1a, R_{+}=0.3a$.
+For the parent domain we set $\bar{\m}_i=0.2a, i=1, \ldots, 10$ as the reference value.
 The point load is distributed as a constant traction over the length of one unit cell (on the far left).
@@
 The number of training samples is chosen as $N_{\train}=50$ ($k=1, 11$), $N_{\train}=100$ ($k=2, 10$) and $N_{\train}=200$ elsewhere, as the local parameter spaces have different dimensions.
+The singular values of the POD at the end of~\cref{algo:rrfpod} for each oversampling problem show a rapid decay~(\cref{fig:singular_values}).
+As expected, a less rapid decay is observed for the larger oversampling domains.
+Furthermore, for $k=1, 2, 3$ in comparison to the oversampling problems of the same size on the right ($k=9, 10, 11$), a difference in the total number of snapshots generated is observed which is due to the additional \textsc{Neumann} problems.
 
-\begin{itemize}
-    \item \added{number of snapshots for each POD}
-    \item difference to full order model: only use operators on unit cell domain in the online phase
-\end{itemize}
-
-\begin{figure}
-    \begin{center}
-        \includegraphics[keepaspectratio,height=4cm]{gradientenbeton_formwork.jpg}
-    \end{center}
-    \caption{Formwork for a graded concrete slab.
+% There is not enough space for this figure
+% \begin{figure}
+%     \begin{center}
+%         \includegraphics[keepaspectratio,height=4cm]{gradientenbeton_formwork.jpg}
+%     \end{center}
+%     \caption{Formwork for a graded concrete slab.
 Weight savings are achieved by placing hollow spheres along a structured grid inside the formwork.
 Source:\gradientenbetonurl.}\label{fig:gradientenbeton}
-\end{figure}
+% \end{figure}
 
@@
 
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=0.6\textwidth]{singular_values.pdf}
+    \end{center}
+    \caption{Decay of singular values for each of the $11$ oversampling problems.
+The total number of basis functions and the total number of snapshots are given in the legend.}\label{fig:singular_values}
+\end{figure}
+
 % subsection Graded concrete slab (end)
@@
 For the validation of the ROM, the relative error with respect to the FOM is computed over a parameter set of size $200$.
-The error in the displacements in measured in the energy norm (over the domain $\varOmega^{\p}_{\gl}$) and defined as
+The error in the displacements is measured in the energy norm (over the domain $\varOmega^{\p}_{\gl}$) and defined as
 \begin{equation}
@@
 \end{equation}
-The error in the first principal component of the \textsc{Cauchy} stress is defined as
+Apart from the displacement field as the primary unknown, we are interested in the first principal component of the \textsc{Cauchy} stress as this will be used to define a stress constraint for the optimization problem.
+The error in the stress is defined as
 \begin{equation}
@@
 The minimum, average and maximum of $e_u$ and $e_{\sigma}$ over the validation set are shown in~\cref{fig:relerr}.
+With regard to the error in the displacement, the average value is much smaller than the maximum over the validation set.
+In both cases, for local basis size ca. $> 75$ the error decreases only very slowly with increasing number of basis functions.
+This is due to the rather moderate size of the training set, as \eg{} $N_{\train}=200$ is used for a local parameter space $\mathcal{P}^k=[R_{-}, R_{+}]^4$ for $k=3, \ldots, 9$.
+However, the error in the stress is less sensitive to variations in the parameter $\m$ and a pointwise maximum relative error below $1\%$ is acceptable for the current application.
 
@@
     \begin{center}
-        \includegraphics[width=0.65\textwidth]{relerr.pdf}
+        \includegraphics[width=0.6\textwidth]{relerr.pdf}
     \end{center}
-    \caption{}\label{fig:relerr}
+    \caption{Minimum, average (dashed) and maximum relative error over the validation set of size $200$.}\label{fig:relerr}
 \end{figure}

### 2024-08-30 src/parageom/definitions.py
add progress

@@ -271,9 +271,10 @@ class BeamData:
         """FOM minimization data"""
         return self.rf / "fom_minimization_data.out"
 
-    def rom_minimization_data(self, distr: str, name: str) -> Path:
+    @property
+    def rom_minimization_data(self) -> Path:
         """ROM minimization data"""
-        return self.rf / f"rom_minimization_data_{distr}_{name}.out"
+        return self.rf / "rom_minimization_data.out"
 
     def pp_stress(self, method: str) -> dict[str, Path]:
         """Postprocessing of stress at optimal design"""

### 2024-08-30 src/parageom/optimization.py

@@ -248,7 +248,7 @@ def main(args):
     with example.fom_minimization_data.open("wb") as fh:
         dump(fom_minimization_data, fh)
 
-    with example.rom_minimization_data(args.distr, args.name).open("wb") as fh:
+    with example.rom_minimization_data.open("wb") as fh:
         dump(rom_minimization_data, fh)
 
 
@@ -360,12 +360,12 @@ def solve_optimization_problem(
 
     c2 = eval_rf_2(opt_result.x)
     strictly_positive = np.all(c2 >= 0.)
-    negative_vals = c2[c2 < 0.]
-    min_value = np.amin(negative_vals)
-    # nearly_zero = np.all(np.isclose(negative_vals, 0., atol=1e-4))
     if strictly_positive:
         return opt_result
     else:
+        negative_vals = c2[c2 < 0.]
+        assert negative_vals.size > 0
+        min_value = np.amin(negative_vals)
         logger.warning(f"Constraint not satisfied by optimal solution (max constraint violation is {np.abs(min_value):2.e}!")
         return opt_result
 

### 2024-08-30 src/parageom/plot_romerr.py

@@ -12,7 +12,7 @@ def main(cli):
     red = bamcd["red"][0]
 
     nreal = cli.nreal
-    number_of_modes = [20, 40, 60, 80, 100]
+    number_of_modes = [20, 40, 60, 80, 100, 120, 140, 160]
 
     erru = defaultdict(list)
     errs = defaultdict(list)

### 2024-08-30 src/parageom/tasks.py

@@ -231,8 +231,9 @@ def task_validate_rom():
         return [action]
 
     num_params = 200
-    number_of_modes = [20, 40, 60, 80, 100]
-    with_ei = {"no_ei": False, "ei": True}
+    number_of_modes = [20, 40, 60, 80, 100, 120, 140, 160]
+    # with_ei = {"no_ei": False, "ei": True}
+    with_ei = {"ei": True}
     num_cells = example.nx * example.ny
 
     for nreal in range(example.num_real):
@@ -281,33 +282,31 @@ def task_validate_rom():
 
 
 
-# def task_optimization():
-#     """ParaGeom: Determine optimal design"""
-#     module = "src.parageom.optimization"
-#     distr = example.distributions[0]
-#
-#     num_modes = 120
-#     minimizer = "SLSQP"
-#     omega = example.omega
-#
-#     nreal = 0 # do optimization only for single realization
-#     for method in example.methods:
-#         deps = [SRC / "optimization.py"]
-#         deps.append(example.coarse_grid("global"))
-#         deps.append(example.parent_domain("global"))
-#         deps.append(example.parent_unit_cell)
-#         for cell in range(5):
-#             deps.append(example.local_basis_npy(nreal, method, distr, cell))
-#         deps.append(example.local_basis_dofs_per_vert(nreal, method, distr))
-#         targets = [example.fom_minimization_data,
-#                    example.rom_minimization_data(distr, method)]
-#         yield {
-#                 "name": method,
-#                 "file_dep": deps,
-#                 "actions": ["python3 -m {} {} {} --num_modes {} --method {} --omega {}".format(module, distr, method, num_modes, minimizer, omega)],
-#                 "targets": targets,
-#                 "clean": True,
-#                 }
+def task_optimization():
+    """ParaGeom: Determine optimal design"""
+    module = "src.parageom.optimization"
+
+    num_modes = 100
+    minimizer = "SLSQP"
+    omega = example.omega
+
+    nreal = 0 # do optimization only for single realization
+    deps = [SRC / "optimization.py"]
+    deps.append(example.coarse_grid("global"))
+    deps.append(example.parent_domain("global"))
+    deps.append(example.parent_unit_cell)
+    for cell in range(example.nx * example.ny):
+        deps.append(example.local_basis_npy(nreal, cell))
+        deps.append(example.local_basis_dofs_per_vert(nreal, cell))
+    targets = [example.fom_minimization_data,
+               example.rom_minimization_data,
[…]

### 2024-08-30 paper/paper.tex
add conclusion

@@
 \usepackage{booktabs}
+\usepackage{tabularx}
 \usepackage{pgfplotstable}
@@
     \left(\pdv{\bm{\psi}_l}{\bm{x}^{\p}}\vdot\bm{F}^{-1}_{\m}\right)\det(\bm{F}_{\m})\dd{\bm{x}^{\p}}\,,\\
-                 &= ({V}^\top)_{km} {K}_{mn}(\mu){V}_{nl}\,.
+                 &= ({V}^\top)_{km} {K}_{mn}(\m){V}_{nl}\,.
     \end{split}
@@
 \end{align}
-Here, $\bm{V}$ is the matrix that holds the FE vectors of the GFEM functions $\bm\psi_i$ as column vectors and $\bm{K}(\mu)$ is the stiffness matrix of the subdomain $\varOmega^{\p}_i$ (\ie{} unit cell).
+Here, $\bm{V}$ is the matrix that holds the FE vectors of the GFEM functions $\bm\psi_i$ as column vectors and $\bm{K}(\m)$ is the stiffness matrix of the subdomain $\varOmega^{\p}_i$ (\ie{} unit cell).
 Since $\bm{K}$ is dependent on the parameter $\m$, integration over the full fine scale mesh of the unit cell is required whenever the model is evaluated for a new parameter value which will drastically reduce the computational efficiency of our method.
@@
 The coarse grid partition of half of the beam body $\varOmega^{\p}_{\gl}=[0, 10a]\times[0, a]$ is given by $10$ quadrilateral cells of size $a$, where $a$ is the unit length of the unit cell.
-Each unit cell, \ie{} subdomain $\varOmega^{\p}_i$, is taken as a square of size $a$ with a circular void with radius $R$ and is partitioned in \added{384} isoparametric quadrilateral FEs of polynomial degree two, see~\cref{fig:oversampling_domain}\,(b).
+Each unit cell, \ie{} subdomain $\varOmega^{\p}_i$, is taken as a square of size $a$ with a circular void with radius $R$ and is partitioned in 384 isoparametric quadrilateral FEs of polynomial degree two, see~\cref{fig:oversampling_domain}\,(b).
 With regard to the geometrical parametrization, we consider the variation of the radius $R$ in each subdomain.
-For the global problem this results in a parameter space $\mathcal{P}_{\gl}=[R_{-}, R_{+}]^{10}$, such that a parameter value $\m\in\mathcal{P}_{\gl}$ has parameter components $\m_i\in[R_{-},R_{+}], i=1, \ldots, 10$, with ranges chosen as $R_{-}=0.1a, R_{+}=0.3a$.
-For the parent domain we set $\bar{\m}_i=0.2a, i=1, \ldots, 10$ as the reference value.
+For the global problem this results in a parameter space $\mathcal{P}_{\gl}=[R_{-}, R_{+}]^{10}$, such that a parameter value $\m\in\mathcal{P}_{\gl}$ has parameter components $\mu_i\in[R_{-},R_{+}], i=1, \ldots, 10$, with ranges chosen as $R_{-}=0.1a, R_{+}=0.3a$.
+For the parent domain we set $\bar{\mu}_i=0.2a, i=1, \ldots, 10$ as the reference value.
 The point load is distributed as a constant traction over the length of one unit cell (on the far left).
@@
 However, the error in the stress is less sensitive to variations in the parameter $\m$ and a pointwise maximum relative error below $1\%$ is acceptable for the current application.
+% \comment[id=pd]{Actually, this is probably only because the ROM stress is computed on the full mesh but with reduced $u$ as input. Reduced stress computation?}
 
@@
 \label{sub:Shape optimization}
+The formulation of the shape optimization problem seeks to minimize an objective function, $J(\m, \bm{u}(\m))$, with constraints, $g_q(\m, \bm{u}(\m)) \le 0$, used to limit the first principal stress at each integration point, $q$, with $q=1, \ldots, N_g$.
+\begin{align}
+    \begin{split}
+        \min_{\m\in\mathcal{P}_{\gl}} &J(\m, \bm{u}(\m))\,,\\
+        \mathrm{s.t.}\quad & g_q(\m, \bm{u}(\m))\le 0, \quad q=1, \ldots, N_q\,,\\
+        \mathrm{with}\quad &\bm{K}(\m)\bm{u}(\m)=\bm{f}\,.
+    \end{split}
+    \label{eq:shape_opt}
+\end{align}
+The objective function is defined as a weighted sum of mass $m(\m)$ and compliance $C(\bm{u}(\m))$.
+\begin{equation}
+    J(\m, \bm{u}(\m))= (1-\omega)\frac{m(\m)}{m(\bar\m)}+\omega \frac{C(\bm{u}(\m))}{C(\bm{u}(\bar\m))}\,,\quad 0\le \omega\le 1\,.
+    \label{eq:output}
+\end{equation}
+The compliance is included, since we do not actively constrain the second principal component and, therefore, designs leading to high compressive stresses would not be detected.
+The reference values $m(\bar\m)$ and $C(\bm{u}(\bar\m))$ are thus used to normalize.
+The definition of the stress constraint follows a risk factor approach~\cite{Pastore2019Topology} and is written as
+\begin{equation}
+    g_q(\m, \bm{u}(\m)) = \frac{\vb{s}_q}{\sigma_{+}} - 1\,,\quad q=1,\ldots,N_q\,,
+    \label{eq:constraint}
+\end{equation}
+with the components of the vector of first principal stress $\vb{s}_q$ and an upper bound $\sigma_{+}$ on the admissible tensile stress.
+Note that $\vb{s}$ depends on $\bm{u}$ and thus also $\m$ although not explicitly written.
 
-\begin{itemize}
-    \item problem definiot
-    \item output computation (difference with ROM)
-    \item result for both models
-\end{itemize}
+The shape optimization problem~\cref{eq:shape_opt} is then solved using the FOM and the ROM.
+Note that the computation of the ROM output $J_n(\m, \bm{u}_n(\m))$, with reduced displacement solution $\bm{u}_n(\m)$, is not dependent on the dimension of the FOM.
+The results are obtained using sequential least squares programming~\cite{kraft1988software} (SLSQP) and finite difference approximation of the gradients, and are given in~\cref{tab:opt_result}.
+% The relative error in the output $J$ is $\approx\,0.4\%$, while the relative error in the mass $m$ is $\approx\,0.5\%$.
+The absolute error in the optimum is $\sqrt{\sum_{i=1}^{10}(\mu_i^{\ast}-\mu_{n,i}^{\ast})^2} \approx 2.44\cdot 10^{-2}$, with the optimum, $\m^{\ast}$, found by the FOM and the optimum, $\m_n^{\ast}$, found by the ROM, respectively.
+While the ROM is able to converge to a solution close to the FOM solution, over three times more iterations and function evaluations are necessary.
+
+\begin{table}
+    \caption{Result of the optimization. For the ROM a local basis size of $100$ is used. $\m^{\ast}$ is used as a placeholder for the optimum found by the respective model.}\label{tab:opt_result}
+    \begin{center}
+        \begin{tabularx}{\textwidth}{XXXXX}
+            \toprule
+            Model & $J(\m^{\ast})$ & $m(\m^{\ast})$ & Iterations & Function calls\\\midrule
+            FOM   &  0.941         &  7.855  & 4          & 45\\
+            ROM   &  0.946         &  7.897  & 13         & 155\\
+            \bottomrule
+        \end{tabularx}
+    \end{center}
+\end{table}
 
@@
 \label{sec:conclusions}
+In this contribution a hyper-reduced localized model order reduction framework for the shape optimization of lattice structures is presented.
+The data compression strategy, adopted from~\cite{Taddei2018Localization}, is shown to work well with range approximation by random sampling.
+Furthermore, the ROM solution of the considered shape optimization problem is in good agreement with the FOM solution and this is regarded as a first proof of concept.
+A thorough comparison of the computational efficiency including runtime measurements is planned for future work, as this is out of the scope of the current investigation.
+Other directions of future research include the exploration of offline training stategies in which the variation in the parameter value $\m$ and the boundary data $\bm{g}$ are addressed simultaneously, instead of sequentially as this may lead to redundant data in the snapshot set $\bm{S}$.
 
 % section CONCLUSION (end)
-
-% \begin{figure}[!htb]
-% \minipage{0.49\textwidth}
-%   \includegraphics[]{./figures/beam/fig_loc_svals_left.pdf}
-%   \caption{Singular valus left}\label{fig:loc_svals_left}
-% \endminipage\hfill
-% \minipage{0.49\textwidth}
-%   \includegraphics[]{./figures/beam/fig_loc_svals_right.pdf}
-%   \caption{Singular valus right}\label{fig:loc_svals_right}
-% \endminipage\hfill\\%
-% \begin{center}
-% \minipage{0.49\textwidth}
-%   \includegraphics[]{./figures/beam/fig_loc_svals_inner.pdf}
-%   \caption{Singular valus inner}\label{fig:loc_svals_inner}
-% \endminipage
-% \end{center}
-% \end{figure}
-
-% \begin{figure}[!htb]
-% \minipage{0.49\textwidth}
-%   \includegraphics{./figures/beam/fig_proj_error_left_hapod.pdf}
-%   \caption{Projection error hapod}\label{fig:proj_error_left_hapod}
-% \endminipage\hfill
-% \minipage{0.49\textwidth}
-%   \includegraphics{./figures/beam/fig_proj_error_right_hapod.pdf}
-%   \caption{Projection error hapod}\label{fig:proj_error_right_hapod}
-% \endminipage\hfill\\
-% \begin{center}
-% \minipage{0.49\textwidth}%
-%   \includegraphics{./figures/beam/fig_proj_error_inner_hapod.pdf}
-%   \caption{Projection error hapod}\label{fig:proj_error_inner_hapod}
-% \endminipage
-% \end{center}
-% \end{figure}
-
-% \begin{table}
-%     \centering
-%     \caption{HAPOD inner}\label{tab:hapod_inner}
-%     \input{./src/beam/tables/hapod_inner.tex}
-% \end{table}
-
-% \begin{figure}[!htb]
-% \minipage{0.49\textwidth}
-%   \includegraphics{./figures/beam/fig_proj_error_left_heuristic.pdf}
-%   \caption{Projection error }\label{fig:proj_error_left_heuristic}
-% \endminipage\hfill
-% \minipage{0.49\textwidth}
-%   \includegraphics{./figures/beam/fig_proj_error_right_heuristic.pdf}
-%   \caption{Projection error }\label{fig:proj_error_right_heuristic}
-% \endminipage\hfill\\
-% \begin{center}
-% \minipage{0.49\textwidth}%
-%   \includegraphics{./figures/beam/fig_proj_error_inner_heuristic.pdf}
-%   \caption{Projection error }\label{fig:proj_error_inner_heuristic}
-% \endminipage
-% \end{center}
-% \end{figure}
-
-% \begin{table}
-%     \centering
-%     \caption{Heuristic rrf inner}\label{tab:heuristic_inner}
-%     \input{./src/beam/tables/heuristic_inner.tex}
-% \end{table}
-
-% \begin{figure}[!htb]
-%     \centering
-%     \includegraphics{./figures/beam/fig_loc_rom_error.pdf}
-%     \caption{Local ROM error relative to FOM.}\label{fig:loc_rom_error}
-% \end{figure}
-
-% \begin{table}
-%     \caption{Result of the optimization with FOM and ROM.}\label{tab:minimization_data}
-%     \centering
-%     \input{./src/beam/tables/minimization_data.tex}
-% \end{table}
-
-% \begin{table}
-%     \caption{Comparison of reduced optimal solution $\mu_N^{\ast}$ and true optimal solution $\mu^{\ast}$.}\label{tab:minimization_comparison}
-%     \centering
-%     \input{./src/beam/tables/minimization_comparison.tex}
-% \end{table}
 

### 2024-08-30 paper/paper.tex
add abstract

@@
 % TODO abstract
-\abstract{\comment[id=pd]{TODO: add abstract}.}
+\abstract{A common engineering practice is to optimize the geometry of a structure by an iterative process, in which an objective function is minimized by systematically choosing the value of design variables and computing the value of the objective function many times.
+However, regarding multiscale problems, this direct numerical approach is not feasible for a number of reasons.
+The model has to take into account both the structural and unit cell (UC) scale, because the objective function to be minimized is a global quantity, while the parameters are related to the shape or material parameters of the UC.
+In the absence of a clear separation of scales, the direct resolution of both scales in the numerical model leads to a significant increase in computational cost, which makes it impossible to repeatedly evaluate the model during the optimization process.
+
+The main goal of this contribution is to overcome the aforementioned limitations and develop an efficient computational framework for the optimal design of lattice structures.
+To this end, parametric model order reduction is combined with domain decomposition methods.
+% The change in geometry of each UC within the lattice is modelled by introducing a transformation that maps the reference to the physical domain.
+In offline simulations, suitable localized approximation spaces, that account for the change in geometry of the UC and its neighbours, are constructed using the concept of oversampling and random sampling.
+The global approximation space is constructed from the local spaces via the partition of unity method.
+The reduced order model is validated on the example of a stress-constrained shape optimization problem.
+}
+
+% TODO adjust abovecaptionskip if necessary
+% it seems the default is 6pt (see class file)
+% \setlength{\abovecaptionskip}{3pt plus 3pt minus 2pt}
 
@@
 
-\added[id=pd]{TODO: add note on notation where appropriate.}\\
-\added[id=pd]{TODO: shorten introduction.}
+\comment[id=pd]{TODO: add note on notation where appropriate.}
+
+\comment[id=pd]{TODO: shorten introduction.}
 
@@
 \end{figure}
-
-% TODO adjust abovecaptionskip if necessary
-% it seems the default is 6pt (see class file)
-% \setlength{\abovecaptionskip}{15pt plus 3pt minus 2pt}
 
@@
 \begin{equation}
-    e_{\sigma}(\m) = \frac{\vert\vb{s}_{\fom} - \vb{s}_{\rom}\vert}{\vert \mathrm{max}(\vb{s}_{\fom})\vert}\,.
+    e_{\sigma}(\m) = \mathrm{max} \frac{\vert\vb{s}_{\fom} - \vb{s}_{\rom}\vert}{\norm{\vb{s}_{\fom}}_2}\,.
     \label{eq:error_sigma}
 \end{equation}
-Here, $\vb{s}\in\mathbb{R}^{N_g}$ denotes the vector holding the first principal stress values at the $N_g$ integration points of the mesh.
+Here, $\norm{\cdot}_2$ denotes the Euclidean norm and $\vb{s}\in\mathbb{R}^{N_g}$ denotes the vector holding the first principal stress values at the $N_g$ integration points of the mesh.
 For ease of notation, the parameter-dependence is omitted for $\bm{u}$ and $\vb{s}$.

### 2024-09-02 paper/paper.tex

@@
 \usepackage{cleveref}
+\usepackage{csquotes}
 \usepackage{algorithm, algorithmicx, algpseudocode}
@@
 
-\usepackage[commentmarkup=uwave]{changes}
+\usepackage[final,commentmarkup=uwave]{changes}
 \usepackage{xspace}
@@
 In the absence of a clear separation of scales, the direct resolution of both scales in the numerical model leads to a significant increase in computational cost, which makes it impossible to repeatedly evaluate the model during the optimization process.
-
 The main goal of this contribution is to overcome the aforementioned limitations and develop an efficient computational framework for the optimal design of lattice structures.
@@
 }
-
 % TODO adjust abovecaptionskip if necessary
@@
 % \setlength{\abovecaptionskip}{3pt plus 3pt minus 2pt}
-
 \begin{document}
@@
 \label{sec:introduction}
+Additive manufacturing \deleted{(AM), commonly known as 3D printing, }is a manufacturing technique that allows for the production of a wide range of structures and complex geometries.
+\deleted{The objects are build successively by adding material layer by layer from three-dimensional models.}
+The technology offers numerous advantages over conventional manufacturing, including greater design flexibility, reduced material waste, and the possibility to produce complex structures with tailored material properties.
+It has been used in a wide range of applications, including the construction industry~\cite{Wu2016critical}.
+\added{The great flexibility regarding the geometry motivates the solution of topology or shape optimization problems to determine the optimal design for a structural problem.}
+\added{In this contribution, we focus on shape optimization in the context of multiscale problems, \ie{} our aim is to optimize the geometrical parameters of the microstructure, \eg{} the unit cell of a lattice structure.}
+\added{However, the solution of the optimization problem requires many evaluations of the numerical model, which may be infeasible if solving the high fidelity finite element (FE) model, also called full order model (FOM) is a computationally demanding task.}
+\added{Reduced order models (ROMs) are powerful tools to mitigate this problem, but for multiscale or large-scale industrial applications with high-dimensional parameter spaces, the nowadays established \enquote{single-domain} approaches are limited~\cite{Ballani2018component}, and the construction of a fast-to-evaluate and accurate ROM is a challenging task.}
+\deleted{A common engineering practice is to optimize the geometry of a structure by an iterative process, in which an objective function is minimized by systematically choosing parameter values $\bm\mu$ (design variables) and computing the value of the function.}
+\deleted{In structural mechanics, a prominent example of an objective function is mass or compliance which is a function of the parameter-dependent displacement field $\bm{u}(\bm\mu)$.}
+\deleted{In each iteration of the optimization process, the geometry is manually changed in the CAD model and the high fidelity finite element (FE) model, also called full order model (FOM), is simply re-evaluated.
+While this approach is not only an ineffective use of resources, it is also infeasible when solving the FE model is a computationally demanding task, \eg{} in multiscale or large-scale industrial applications.
+Due to the prohibitive cost of even a single FOM solution, the iterative design process cannot be performed using direct numerical simulations.}
 
-Additive manufacturing (AM), commonly known as 3D printing, is a manufacturing technique that allows for the production of a wide range of structures and complex geometries.
-The objects are build successively by adding material layer by layer from three-dimensional models.
-The technology offers numerous advantages over conventional manufacturing, including greater design flexibility, reduced material waste, and the possibility to produce complex structures with tailored material properties.
-It has been used in a wide range of applications, including aerospace, biomechanical, automotive, and construction industries~\cite{Plessis2022Properties,Wu2016critical}.
-
-A common engineering practice is to optimize the geometry of a structure by an iterative process, in which an objective function is minimized by systematically choosing parameter values $\bm\mu$ (design variables) and computing the value of the function.
-In structural mechanics, a prominent example of an objective function is mass or compliance which is a function of the parameter-dependent displacement field $\bm{u}(\bm\mu)$.
-In each iteration of the optimization process, the geometry is manually changed in the CAD model and the high fidelity finite element (FE) model, also called full order model (FOM), is simply re-evaluated.
-While this approach is not only an ineffective use of resources, it is also infeasible when solving the FE model is a computationally demanding task, \eg{} in multiscale or large-scale industrial applications.
-Due to the prohibitive cost of even a single FOM solution, the iterative design process cannot be performed using direct numerical simulations.
 Therefore, we propose a framework based on localized, also called component-based (CB), parametric model order reduction (pMOR).
@@
 The global approximation is obtained by a suitable coupling of the local reduced spaces spanned by the aforementioned basis functions, in which one naturally relies on domain decomposition (DD) strategies.
-A review of concepts in localized model order reduction is presented in~\cite{BuhrReview}.
+A review of concepts in localized \replaced{pMOR}{model order reduction} is presented in~\cite{BuhrReview}.
 In particular, regarding lattice structures, this approach allows to take advantage of the repetitiveness of the lattice, such that a computation of the local basis is required only for few components, \ie{} unit cells.
-
 The computation of the local basis is an essential task in localized pMOR and is often done using the concept of oversampling~\cite{Hou1997Multiscale}.
 In this approach, the target subdomain $\varOmega_{\mathrm{in}}$, \ie{} that part of the domain for which one would like to construct basis functions, is extended and boundary conditions are prescribed on the boundary of the larger so-called oversampling domain $\varOmega$ to explore possible solutions.
-In the literature, this oversampling problem is also expressed in terms of a \textit{transfer operator} $\bm{T}$ that maps the values on the boundary $\partial\varOmega$ to the unknown solution restricted to the target subdomain $\bm{u}(\bm{\mu})\vert_{\varOmega_{\mathrm{in}}}$.
+In the literature, this oversampling problem is also expressed in terms of a \textit{transfer operator} $\bm{T}$ that maps the values on the boundary $\partial\varOmega$ to the unknown solution restricted to the target subdomain $\bm{u}\vert_{\varOmega_{\mathrm{in}}}$.
 The construction of (optimal) local approximation spaces then comprises the calculation of the left singular vectors of this transfer operator~\cite{Babuska2011Optimal,Smetana2016Optimal}.
@@
 Second, a framework for the shape optimization of lattice structures is proposed.
-It combines the aforementioned algorithm to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final reduced order model (ROM).
+It combines the aforementioned algorithm to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final \replaced{ROM}{reduced order model (ROM)}.
 Furthermore, the global approximation space is constructed from the local spaces using the partition of unity method or generalized finite element method (GFEM)~\cite{BABUSKA1997,Babuska2004Generalized}.
@@
 \Cref{sec:numerical experiments} comprises the numerical experiments.
-Based on the example of a graded concrete slab, the quality of the local spaces generated by the proposed method is analyzed, the resulting ROM is validated, and results for the solution of a shape optimization problem are presented.
+Based on the example of a graded concrete slab, the quality of the local spaces generated by the proposed method is analyzed, the resulting ROM is validated, and results for the solution of a \added{stress-constrained }shape optimization problem are presented.
 Finally, conclusions are given in~\cref{sec:conclusions}.
-
-\comment[id=pd]{TODO: add note on notation where appropriate.}
-
-\comment[id=pd]{TODO: shorten introduction.}
 
@@
 The approach adopted in this work to facilitate geometrical parameterizations belongs to the class of surface-based deformations~\cite{Botsch2010Polygon} and is based on transformations $\Phi_{\bm\mu}$ that map each material point $\bm{x}_{\mathrm{p}}$ of a parameter-independent reference or parent domain $\varOmega^{\mathrm{p}}$ to a point $\bm{x}_{\bm\mu}$ in the parameter-dependent current or physical domain $\varOmega^{\bm\mu}$.
+Throughout the paper, the parameter value (design variables) are denoted by $\m$ and dependence on the parameter is indicated by a sub- or superscript $(\cdot)_{\m}, (\cdot)^{\m}$.
+The sub- or superscript $(\cdot)_{\p}, (\cdot)^{\p}$ indicates that the quantity does not depend on the parameter.
 In the context of shape optimization of lattice structures, our objective is to determine the mapping $\Phi_{\bm\mu}$ for a single unit cell, see~\cref{fig:transformationmap}, and use it to describe the change in geometry of each unit cell throughout the structure.
@@
 With $\bm{x}_{\bm\mu}$ known for all points on the parent interface $\partial\varOmega^{\mathrm{p}}_{\mathrm{int}}$, the desired transformation (\eg{} enlarging/shrinking of the voids radius) is enforced by prescribing~\cref{eq:aux_2,eq:aux_3}.
-
 Given the transformation displacement $\bm{d}(\bm{x}_{\mathrm{p}};\bm\mu)$, the variational formulation is formulated over the parameter-independent parent domain $\varOmega^{\mathrm{p}}$ instead of the parameter-dependent physical domain $\varOmega^{\bm\mu}$.
@@
 
+%\begin{figure}
+%    \centering
+%    \include{./img/transformationmap.tex}
+%    \caption{Transformation map $\Phi_{\bm\mu}$ from parent to physical domain on the example of the unit square domain with a circular void placed in the center.
+The parameter $\bm\mu$ controls the radius of the void.}\label{fig:transformationmap}
+%\end{figure}
 \begin{figure}
 	\centering
-    \include{./img/transformationmap.tex}
+	\includegraphics[width=0.65\textwidth]{transformationmap.pdf}
 	\caption{Transformation map $\Phi_{\bm\mu}$ from parent to physical domain on the example of the unit square domain with a circular void placed in the center.
@@
 Also, for construction of a sufficiently rich local basis it has to be solved for many parameter values $\m$, but a sampling of the boundary values on $\varGamma^{\m}_{\out}$ is not required.
-% - setting, global domain = union of non overlapping subdomains = unit cells
-% - introduction of coarse grid where each cell corresponds to unit cell subdomain
-% - goal of constructing local spaces for only parts of the domain
-% - definition of oversampling problem and transfer operator
-% these are details of the specific problem we solve later
-% - the oversampling problem should be formulated for union of parent unit cells
-% - weak formulation is not required here? Should be first introduced in section 2.3?
-% - comment on number of archetypes/configurations
 
-\begin{figure}
+%\begin{figure}[h]
+%	\centering
+%    \include{./img/oversampling_domain.tex}
+%    \caption{Exemplary coarse grid discretization of the global domain $\varOmega^{\m}_{\gl}$, with target subdomain $\varOmega^{\m}_i$ within oversampling domain $\varOmega^{\m}$ in grey (a) and fine grid discretization of the unit cell (b).}%
+%    \label{fig:oversampling_domain}
+%\end{figure}
+\begin{figure}[h]
 	\centering
-    \include{./img/oversampling_domain.tex}
-    \caption{Exemplary coarse grid discretization of the global domain $\varOmega^{\m}_{\gl}$, with target subdomain $\varOmega^{\m}_i$ within oversampling domain $\varOmega^{\m}$ in grey (a) and fine grid discretization of the unit cell (b).}\label{fig:oversampling_domain}
+	\includegraphics[width=0.65\textwidth]{oversampling_domain.pdf}
+	\caption{Exemplary coarse grid discretization of the global domain $\varOmega^{\m}_{\gl}$, with target subdomain $\varOmega^{\m}_i$ within oversampling domain $\varOmega^{\m}$ in grey (a) and fine grid discretization of the unit cell (b).}%
+	\label{fig:oversampling_domain}
 \end{figure}
@@
 However, given a suitably\footnote{According to Definition 1 of~\cite{BABUSKA1997}.} defined partition of unity $\{{\varphi}_i\}$, the local basis functions can be included in the global FE space by multiplication with the standard FE shape functions ${\varphi}_i$, such that a conforming approximation is obtained.
-The global GFEM space is defined as
-\begin{equation}
-	\label{eq:Xgfem}
-    X_{\mathrm{GFEM}}:= \sum_{i} \varphi_iX^n_i = \left\lbrace\sum_i\varphi_i\bm\xi_i \;\big|\; \bm\xi_i\in X^n_i	\right\rbrace\,.
-\end{equation}
-Furthermore, the GFEM function denoted as $\bm{\psi}(\bm{x}) = \varphi(\bm{x})\bm{\xi}(\bm{x}), \bm{x}\in\omega_{\alpha}$ for a patch $\omega_{\alpha}$ is shown in~\cref{fig:gfem} to illustrate the procedure of constructing the GFEM functions.
+%The global GFEM space is defined as
+%\begin{equation}
+%	\label{eq:Xgfem}
+%    X_{\mathrm{GFEM}}:= \sum_{i} \varphi_iX^n_i = \left\lbrace\sum_i\varphi_i\bm\xi_i \;\big|\; \bm\xi_i\in X^n_i	\right\rbrace\,.
+%\end{equation}
+The GFEM function denoted as $\bm{\psi}(\bm{x}) = \varphi(\bm{x})\bm{\xi}(\bm{x}), \bm{x}\in\omega_{\alpha}$ for a patch $\omega_{\alpha}$ is shown in~\cref{fig:gfem} to illustrate the procedure of constructing the GFEM functions.
 \begin{figure}[h]
@@
        }
-       \caption{Construction of GFEM functions.}
+       \caption{Construction of GFEM functions.}%
        \label{fig:gfem}
@@
 For more details on the overall procedure the reader is referred to~\cite{Negri2015Efficient}.
-
 Finally, with a fast method to compute $\bm{K}_Q(\m)$, the calculation of the FE matrices $a^i_{kl}$ of each coarse grid cell are independent of the underlying fine grid discretization of the unit cell.
@@
 The compliance is included, since we do not actively constrain the second principal component and, therefore, designs leading to high compressive stresses would not be detected.
-The reference values $m(\bar\m)$ and $C(\bm{u}(\bar\m))$ are thus used to normalize.
+The reference values $m(\bar\m)$ and $C(\bm{u}(\bar\m))$ are used to normalize.
 The definition of the stress constraint follows a risk factor approach~\cite{Pastore2019Topology} and is written as
@@
 Note that $\vb{s}$ depends on $\bm{u}$ and thus also $\m$ although not explicitly written.
-
+The load $F$ is chosen so that the parameter value with $\mu_i=R_{+}, i=1, \ldots, 10$ yields $g_q>0$ (infeasible design), such that this trivial minimum of $m(\m)$ is excluded from the set of possible solutions.
 The shape optimization problem~\cref{eq:shape_opt} is then solved using the FOM and the ROM.
@@
 A thorough comparison of the computational efficiency including runtime measurements is planned for future work, as this is out of the scope of the current investigation.
-Other directions of future research include the exploration of offline training stategies in which the variation in the parameter value $\m$ and the boundary data $\bm{g}$ are addressed simultaneously, instead of sequentially as this may lead to redundant data in the snapshot set $\bm{S}$.
+Other directions of future research include the exploration of offline training strategies in which the variation in the parameter value $\m$ and the boundary data $\bm{g}$ are addressed simultaneously and suitable sampling distributions of the boundary data $\bm{g}$.
 

### 2024-09-02 src/parageom/auxiliary_problem.py
install parageom-0.1.0; remove relative imports

@@ -15,7 +15,7 @@ from multi.problems import LinearElasticityProblem
 
 from pymor.parameters.base import Mu, Parameters
 
-from .definitions import BeamData
+from parageom.definitions import BeamData
 
 
 class GlobalAuxiliaryProblem:
@@ -293,7 +293,7 @@ def discretize_auxiliary_problem(example: BeamData, fine_grid: str, facet_tags: 
 
 
 def main():
-    from .tasks import example
+    from parageom.tasks import example
     from dolfinx.io.utils import XDMFFile
 
     # transformation displacement is used to construct

### 2024-09-02 src/parageom/gfem.py

@@ -26,8 +26,8 @@ def enrichment_from_kernel(kernel):
 
 
 def main(args):
-    from .tasks import example
-    from .osp_v2 import oversampling_config_factory
+    from parageom.tasks import example
+    from parageom.hapod import oversampling_config_factory
 
     stem = pathlib.Path(__file__).stem  # gfem
     logfilename = example.log_gfem(args.nreal, args.cell).as_posix()

### 2024-09-02 src/parageom/hapod.py

@@ -1,20 +1,131 @@
-import sys
-
-from pathlib import Path
-
+import typing
+from dataclasses import dataclass
+import dolfinx as df
+import dolfinx.fem.petsc
 import numpy as np
+import numpy.typing as npt
+import ufl
+from mpi4py import MPI
+from petsc4py import PETSc
+from multi.io import read_mesh
+from multi.domain import StructuredQuadGrid, RectangularDomain
+from multi.preprocessing import create_meshtags
+from multi.boundary import within_range, plane_at, point_at
+from multi.materials import LinearElasticMaterial
+from multi.problems import LinearElasticityProblem
+from multi.solver import build_nullspace
+from multi.projection import orthogonal_part
+
+from pymor.algorithms.gram_schmidt import gram_schmidt
+from pymor.algorithms.pod import pod
+from pymor.bindings.fenicsx import FenicsxVectorSpace, FenicsxMatrixOperator
+from pymor.parameters.base import Parameters, ParameterSpace
+from pymor.operators.numpy import NumpyMatrixOperator
+from pymor.operators.interface import Operator
+from pymor.core.defaults import set_defaults
+from pymor.core.logger import getLogger
+from pymor.tools.random import new_rng
+
+from scipy.sparse import csr_array
 from scipy.sparse.linalg import LinearOperator, eigsh
 from scipy.special import erfinv
 
-from pymor.algorithms.pod import pod
-from pymor.algorithms.gram_schmidt import gram_schmidt
-from pymor.core.defaults import set_defaults
-from pymor.core.logger import getLogger
-from pymor.operators.interface import Operator
-from pymor.tools.random import new_rng
-from pymor.parameters.base import ParameterSpace
 
-from multi.projection import orthogonal_part
+@dataclass
+class OversamplingConfig:
+    cells_omega: npt.NDArray[np.int32]
+    cells_omega_in: npt.NDArray[np.int32]
+    kernel: tuple[int]
+    gamma_out: typing.Callable
+    gamma_d: typing.Optional[typing.Callable] = None
+    gamma_n: typing.Optional[typing.Callable] = None
+
+
+def oversampling_config_factory(k):
+    """Creates instance of `OversamplingConfig`"""
+
+    cells_omega = {
+            0: np.array([0, 1], dtype=np.int32),
[…]

### 2024-09-02 src/parageom/locmor.py

@@ -34,9 +34,9 @@ from multi.io import read_mesh
 from multi.interpolation import make_mapping
 from multi.sampling import create_random_values
 from multi.utils import LogMixin
-from .definitions import BeamData
-from .matrix_based_operator import FenicsxMatrixBasedOperator
-from .dofmap_gfem import GFEMDofMap
+from parageom.definitions import BeamData
+from parageom.matrix_based_operator import FenicsxMatrixBasedOperator
+from parageom.dofmap_gfem import GFEMDofMap
 
 
 EISubdomainOperatorWrapper = namedtuple(
@@ -285,7 +285,7 @@ def assemble_gfem_system(
 
     """
 
-    from .dofmap_gfem import select_modes
+    from parageom.dofmap_gfem import select_modes
 
     # no need to apply bcs as basis functions should
     # satisfy these automatically
@@ -408,7 +408,7 @@ def assemble_gfem_system_with_ei(
 
     """
 
-    from .dofmap_gfem import select_modes
+    from parageom.dofmap_gfem import select_modes
 
     # TODO: try to set BCs to better condition system matrix
 
@@ -681,9 +681,9 @@ def discretize_transfer_problem(example: BeamData, configuration: str) -> tuple[
         configuration: The configuration/archetype.
     """
 
-    from .fom import ParaGeomLinEla
-    from .auxiliary_problem import GlobalAuxiliaryProblem
-    from .matrix_based_operator import FenicsxMatrixBasedOperator, BCGeom, BCTopo, _create_dirichlet_bcs
+    from parageom.fom import ParaGeomLinEla
+    from parageom.auxiliary_problem import GlobalAuxiliaryProblem
+    from parageom.matrix_based_operator import FenicsxMatrixBasedOperator, BCGeom, BCTopo, _create_dirichlet_bcs
     from multi.preprocessing import create_meshtags
 
     global_coarse_domain, _, _ = read_mesh(example.coarse_grid("global"), MPI.COMM_WORLD,
@@ -973,7 +973,5 @@ def discretize_transfer_problem(example: BeamData, configuration: str) -> tuple[
 
 
 if __name__ == "__main__":
-    from .tasks import example
-    # discretize_transfer_problem(example, "left")
+    from parageom.tasks import example
     discretize_transfer_problem(example, "right")
-    # discretize_transfer_problem(example, "inner")

### 2024-09-02 src/parageom/optimization.py

@@ -32,10 +32,10 @@ ModelWrapper = namedtuple(
 
 def main(args):
     """solve optimization problem for different models"""
-    from .tasks import example
-    from .validate_rom import build_fom, build_rom
-    from .dofmap_gfem import GFEMDofMap
-    from .locmor import reconstruct
+    from parageom.tasks import example
+    from parageom.validate_rom import build_fom, build_rom
+    from parageom.dofmap_gfem import GFEMDofMap
+    from parageom.locmor import reconstruct
 
     stem = pathlib.Path(__file__).stem
     # always use first realization for optimization, nreal=0

### 2024-09-02 src/parageom/osp_v2.py

@@ -1,522 +1,0 @@
-"""oversampling version 2"""
-
-import typing
-from dataclasses import dataclass
-import dolfinx as df
-import dolfinx.fem.petsc
-import numpy as np
-import numpy.typing as npt
-import ufl
-from mpi4py import MPI
-from petsc4py import PETSc
-from multi.io import read_mesh
-from multi.domain import StructuredQuadGrid, RectangularDomain
-from multi.preprocessing import create_meshtags
-from multi.boundary import within_range, plane_at, point_at
-from multi.materials import LinearElasticMaterial
-from multi.problems import LinearElasticityProblem
-from multi.solver import build_nullspace
-from multi.projection import orthogonal_part
-
-from pymor.algorithms.gram_schmidt import gram_schmidt
-from pymor.algorithms.pod import pod
-from pymor.bindings.fenicsx import FenicsxVectorSpace, FenicsxMatrixOperator
-from pymor.parameters.base import Parameters, ParameterSpace
-from pymor.operators.numpy import NumpyMatrixOperator
-from pymor.core.defaults import set_defaults
-from pymor.core.logger import getLogger
-from pymor.tools.random import new_rng
-
-from scipy.sparse import csr_array
-
-
-@dataclass
-class OversamplingConfig:
-    cells_omega: npt.NDArray[np.int32]
-    cells_omega_in: npt.NDArray[np.int32]
-    kernel: tuple[int]
-    gamma_out: typing.Callable
-    gamma_d: typing.Optional[typing.Callable] = None
-    gamma_n: typing.Optional[typing.Callable] = None
-
-
-def oversampling_config_factory(k):
-    """Creates instance of `OversamplingConfig`"""
-
-    cells_omega = {
-            0: np.array([0, 1], dtype=np.int32),
-            1: np.array([0, 1, 2], dtype=np.int32),
-            2: np.array([0, 1, 2, 3], dtype=np.int32),
-            3: np.array([1, 2, 3, 4], dtype=np.int32),
-            4: np.array([2, 3, 4, 5], dtype=np.int32),
-            5: np.array([3, 4, 5, 6], dtype=np.int32),
-            6: np.array([4, 5, 6, 7], dtype=np.int32),
-            7: np.array([5, 6, 7, 8], dtype=np.int32),
-            8: np.array([6, 7, 8, 9], dtype=np.int32),
-            9: np.array([7, 8, 9], dtype=np.int32),
[…]

### 2024-09-02 src/parageom/plot_projerr.py

@@ -4,7 +4,7 @@ from multi.plotting_context import PlottingContext
 
 
 def postproc(nreal: int, method: str, distr: str, config: str, eps: float):
-    from .tasks import example
+    from parageom.tasks import example
 
     instream = example.projerr(nreal, method, distr, config)
     data = np.load(instream)
@@ -24,7 +24,7 @@ def postproc(nreal: int, method: str, distr: str, config: str, eps: float):
 
 
 def main(cli):
-    from .tasks import example
+    from parageom.tasks import example
 
     distr = example.distributions[0]
     bamcd = read_bam_colors()

### 2024-09-02 src/parageom/plot_svals_osp.py

@@ -3,7 +3,7 @@ from multi.plotting_context import PlottingContext
 
 
 def main(cli):
-    from .tasks import example
+    from parageom.tasks import example
 
     pod = {
     0     :    "58/420",

### 2024-09-02 src/parageom/pp_stress.py

@@ -13,9 +13,9 @@ from pymor.models.basic import StationaryModel
 
 def main(args):
     """compute principal stress using FOM and ROM"""
-    from .tasks import example
-    from .auxiliary_problem import discretize_auxiliary_problem
-    from .fom import discretize_fom, ParaGeomLinEla
+    from parageom.tasks import example
+    from parageom.auxiliary_problem import discretize_auxiliary_problem
+    from parageom.fom import discretize_fom, ParaGeomLinEla
 
     # ### Build FOM
     coarse_grid_path = example.coarse_grid("global").as_posix()
@@ -93,7 +93,7 @@ def main(args):
 
 
 def compute_principal_stress(model, mu, u, stress, parageom, rec_data=None, xdmf_filename=None):
-    from .stress_analysis import principal_stress_2d, project
+    from parageom.stress_analysis import principal_stress_2d, project
 
     V = u.function_space
     domain = V.mesh
@@ -143,10 +143,10 @@ def compute_principal_stress(model, mu, u, stress, parageom, rec_data=None, xdmf
 
 
 def build_localized_rom(cli, example, global_auxp, trafo_disp, parameters, ω=0.5):
-    from .fom import discretize_subdomain_operators, ParaGeomLinEla
-    from .ei import interpolate_subdomain_operator
-    from .locmor import EISubdomainOperatorWrapper, assemble_gfem_system_with_ei
-    from .dofmap_gfem import GFEMDofMap
+    from parageom.fom import discretize_subdomain_operators, ParaGeomLinEla
+    from parageom.ei import interpolate_subdomain_operator
+    from parageom.locmor import EISubdomainOperatorWrapper, assemble_gfem_system_with_ei
+    from parageom.dofmap_gfem import GFEMDofMap
 
     nreal = 0
 

### 2024-09-02 src/parageom/preprocessing.py

@@ -194,7 +194,7 @@ def create_fine_scale_grid(example, typus: str, output: str):
 if __name__ == "__main__":
     import sys
     import argparse
-    from .tasks import example
+    from parageom.tasks import example
 
     parser = argparse.ArgumentParser(
         description="Preprocessing for the parageom example."

### 2024-09-02 src/parageom/tasks.py

@@ -4,7 +4,7 @@ import os
 import shutil
 from pathlib import Path
 from doit.tools import run_once
-from .definitions import BeamData, ROOT
+from parageom.definitions import BeamData, ROOT
 
 os.environ["PYMOR_COLORS_DISABLE"] = "1"
 example = BeamData(name="parageom", run_mode="DEBUG")
@@ -44,7 +44,7 @@ def with_h5(xdmf: Path) -> list[Path]:
 
 def task_parent_unit_cell():
     """ParaGeom: Create mesh for parent unit cell"""
-    from .preprocessing import discretize_unit_cell
+    from parageom.preprocessing import discretize_unit_cell
 
     def create_parent_unit_cell(targets):
         unit_length = example.unit_length
@@ -63,15 +63,13 @@ def task_parent_unit_cell():
 
 def task_coarse_grid():
     """ParaGeom: Create structured coarse grids"""
-    module = "src.parageom.preprocessing"
-
     for config in ["global"]:
         yield {
             "name": config,
-            "file_dep": [],
+            "file_dep": [SRC / "preprocessing.py"],
             "actions": [
-                "python3 -m {} {} {} --output %(targets)s".format(
-                    module, config, "coarse"
+                "python3 %(dependencies)s {} {} --output %(targets)s".format(
+                    config, "coarse"
                 )
             ],
             "targets": [example.coarse_grid(config)],
@@ -82,16 +80,11 @@ def task_coarse_grid():
 
 def task_fine_grid():
     """ParaGeom: Create parent domain mesh"""
-    module = "src.parageom.preprocessing"
     for config in ["global"]:
         yield {
             "name": config,
-            "file_dep": [example.coarse_grid(config), example.parent_unit_cell],
-            "actions": [
-                "python3 -m {} {} {} --output %(targets)s".format(
-                    module, config, "fine"
-                )
-            ],
+            "file_dep": [example.coarse_grid(config), example.parent_unit_cell, SRC / "preprocessing.py"],
+            "actions": ["python3 src/parageom/preprocessing.py {} {} --output %(targets)s".format(config, "fine")],
             "targets": [example.parent_domain(config)],
             "clean": True,
         }
@@ -107,11 +100,10 @@ def task_preproc():
 
 def task_hapod():
     """ParaGeom: Construct basis via HAPOD"""
-    module = "src.parageom.osp_v2"
 
     for nreal in range(example.num_real):
         for k in range(11):
-            deps = [SRC / "osp_v2.py"]
[…]

### 2024-09-02 src/parageom/run_locrom.py
rm unused files; mv

@@ -1,322 +1,0 @@
-"""Build and run localized ROM."""
-
-from pathlib import Path
-from collections import defaultdict
-
-import numpy as np
-
-import dolfinx as df
-from dolfinx.common import Timer
-
-from pymor.core.defaults import set_defaults
-from pymor.core.logger import getLogger
-from pymor.operators.constructions import VectorOperator
-from pymor.models.basic import StationaryModel
-from pymor.tools.random import new_rng
-
-
-def main(args):
-    from .tasks import example
-    from .auxiliary_problem import discretize_auxiliary_problem
-    from .fom import discretize_fom, discretize_subdomain_operators
-
-    from .ei import interpolate_subdomain_operator
-    from .locmor import reconstruct, assemble_gfem_system, assemble_gfem_system_with_ei, EISubdomainOperatorWrapper
-    from .dofmap_gfem import GFEMDofMap
-
-    # ### logger
-    set_defaults(
-        {
-            "pymor.core.logger.getLogger.filename": example.log_run_locrom(
-                args.nreal, args.method, args.distr, ei=args.ei
-            )
-        }
-    )
-    if args.debug:
-        loglevel = "DEBUG"
-    else:
-        loglevel = "INFO"
-    logger = getLogger(Path(__file__).stem, level=loglevel)
-
-    # ### Discretize global FOM
-    global_parent_domain_path = example.parent_domain("global")
-    coarse_grid_path = example.coarse_grid("global")
-    # do not change interface tags; see src/parageom/preprocessing.py::create_parent_domain
-    interface_tags = [i for i in range(15, 25)]
-    global_auxp = discretize_auxiliary_problem(
-        example,
-        global_parent_domain_path.as_posix(),
-        interface_tags,
-        example.parameters["global"],
-        coarse_grid=coarse_grid_path.as_posix(),
-    )
-    trafo_d_gl = df.fem.Function(global_auxp.problem.V, name="d_trafo")
-    fom, _ = discretize_fom(example, global_auxp, trafo_d_gl)
-    # h1_product = fom.products["h1_0_semi"]
-    energy_product = fom.products["energy"]
-
-    # ### Discretize subdomain operators
-    # NOTE
-    # rhs_local is non-zero although this should only be the case for cell 0
-    # this is handled during assembly
-    # see `assemble_gfem_system` and `assemble_gfem_system_with_ei`
[…]
